```{r,data,generation,echo=FALSE,results="hide"}
while(T){
  type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
  n = sample(7:15,1)
  beta1 = c(1,3,0.2,0.5,2)[type]
  beta2 = c(1,3,0.2,2,0.5)[type]
  bottom = sample(seq(20,70,5),1)
  width = sample(seq(5,30,5),1)
  i = 1:n
  x = round(rbeta(n,beta1,beta2)*width+bottom,2)
  xbar = mean(x)
  devs = x-xbar
  sdevs = devs^2
  var = sum(sdevs)/(n-1)
  if(round(xbar,2)==xbar){
    break
  }
}

mydf = data.frame(i,x,devs,sdevs)
colnames(mydf) = c("$$i$$","$$\\mathbf{x}$$","deviations = $\\mathbf{x}-\\bar{x}$","SqrDev = $(\\mathbf{x}-\\bar{x})^2$")
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r paste(lapply(x,function(x) sprintf("%.2f",x)),collapse=", ")` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_VAR.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_VAR.csv).

Determine the unbiased sample [variance](https://en.wikipedia.org/wiki/Variance#Sample_variance){target="blank"} ([with Bessel correction](https://en.wikipedia.org/wiki/Bessel's_correction#Formula){target="blank"}).  You can round your answer to the [hundredths place](https://en.wikipedia.org/wiki/Orders_of_magnitude_(numbers)#10%E2%88%922){target="blank"}.

You probably wonder why you would make the Bessel correction. The reason is important. We will see that the main goal of statistics is to infer the underlying probability distribution of a collection of empirical observations. In other words, we have a [lottery machine](https://en.wikipedia.org/wiki/Lottery_machine){target="blank"} filled with many balls (population), but we only see a [small sample](https://en.wikipedia.org/wiki/Simple_random_sample){target="blank"} of those balls, and our goal is to guess what the population looks like based on a small sample (see [statistical inference](https://en.wikipedia.org/wiki/Statistical_inference){target="blank"}).

It turns out that when guessing the population's variance from a small sample, your guess is better after making the Bessel correction.



Solution
========

(ANS=`r var`)

First, determine the mean of the sample.

$$\bar{x} = \frac{\sum \mathbf{x}}{n} =`r mean(x)` $$

Determine the squared deviations (squared distances from $\bar{x}$).

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>

```{r tab1, results="asis", echo=F}
knitr::kable(mydf,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```


Determine the unbiased sample variance by summing the squared deviations and dividing the sum by $n-1$.

$$s^2 = \frac{\sum (\mathbf{x}-\bar{x})^2}{n-1} = \frac{`r sum(sdevs)`}{`r n`-1} = `r var` $$

### spreadsheet

```{r echo=F}
xbar = mean(x)
sol = data.frame(i,x)

sol = rbind(sol,
            c("",""),
            c("","xbar"),
            c("",paste0("=average(B2:B",n+1,")",collapse="")),
            c("","")
            )

deviations = c(paste0("=B",2:(n+1),"-B$",n+4),rep("",4))
sqrdev = c(paste0("=C",2:(n+1),"^2"),
           "",
           "VAR",
           paste0("=sum(D2:D",n+1,")/",n-1),
           paste0("=VAR(B2:B",n+1,")"))
sol = cbind(sol,deviations,sqrdev)
write.csv(sol,file="get_VAR_sol.csv", row.names=F)

```

You can do this with [a spreadsheet](get_VAR_sol.csv).

```{r fig.height=6,fig.width=8,echo=F}
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol)
```
\

And, actually, you can skip a lot of work by using the [`VAR` function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_VAR_function){target="blank"}.

### R

You can do this with R

```{r}
x = read.csv("get_VAR.csv")$x
var(x)
```

You could also do it the long way.

```{r}
x = read.csv("get_VAR.csv")$x
n = length(x)
xbar = mean(x)
deviations = x-xbar
sqrdev = deviations^2
VAR = sum(sqrdev)/(n-1)
VAR
```


Meta-information
================
extype: num
exsolution: `r var(x)`
exname: get_VAR
extol: 0.01
exextra[numwidth]: 999.999