```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}

type = sample(1:3,3) #bell, uniform, bimodal
n = 10000
beta1 = c(2,1,0.6)[type]
beta2 = c(2,1,0.6)[type]
mid = sample(seq(30,70,5),3)
width = sample(seq(20,60,10),1)
bottom = mid-width/2

x1 = rbeta(n,beta1[1],beta2[1])*(width-0.02)+bottom[1]+0.01
x2 = rbeta(n,beta1[2],beta2[2])*(width-0.02)+bottom[2]+0.01
x3 = rbeta(n,beta1[3],beta2[3])*(width-0.02)+bottom[3]+0.01

h1 = hist(x1,xlim=c(0,100),breaks=seq(0,100,2.5))
h2 = hist(x2,xlim=c(0,100),breaks=seq(0,100,2.5))
h3 = hist(x3,xlim=c(0,100),breaks=seq(0,100,2.5))

counts1 = h1$counts
counts2 = h2$counts
counts3 = h3$counts

ym = max(c(counts1,counts2,counts3))

ans1 = paste0((mid==min(mid))*1,collapse="")
ans2 = paste0((mid==max(mid))*1,collapse="")
ans3 = paste0((type==min(type))*1,collapse="")
ans4 = paste0((type==max(type))*1,collapse="")

```


Question
========

Three large samples were taken from three different populations. Their distributions are shown as histograms.

```{r echo=F,fig.width=6,fig.height=4}
par(mfrow=c(3,1),mar=c(0,0,0,0))
par(mar=c(2,2,2,2))
hist(x1,xlim=c(0,100),breaks=seq(0,100,2.5),ann=F,axes=F,ylim=c(0,ym),col=rgb(0.9,0.7,0.7))
axis(1,line=0,at=seq(0,100,5))
mtext("sample X",2,0,srt=90)

par(mar=c(2,2,2,2))
hist(x2,xlim=c(0,100),breaks=seq(0,100,2.5),ann=F,axes=F,ylim=c(0,ym),col=rgb(0.9,0.7,0.7))
axis(1,line=0,at=seq(0,100,5))
mtext("sample Y",2,0,srt=90)

par(mar=c(2,2,2,2))
hist(x3,xlim=c(0,100),breaks=seq(0,100,2.5),ann=F,axes=F,ylim=c(0,ym),col=rgb(0.9,0.7,0.7))
axis(1,line=0,at=seq(0,100,5))
mtext("sample Z",2,0,srt=90)
```
\

a.  Which sample has the smallest mean? \#\#ANSWER1\#\#
b.  Which sample has the largest mean? \#\#ANSWER2\#\#
c.  Which sample has the smallest standard deviation? \#\#ANSWER3\#\#
d.  Which sample has the largest standard deviation? \#\#ANSWER4\#\#


Answerlist
--------
* X
* Y
* Z
* X
* Y
* Z
* X
* Y
* Z
* X
* Y
* Z


Solution
========

All three distributions look like they have similar ranges (widths), but different shapes. So, we will use the fact that for a given range, bell shape has the smallest standard deviation and bimodal has the largest standard deviation. This is because a bell shape has many measurements near the middle, whereas the bimodal shape has many measurements near the edges of its interval. 


a. (`r c("X","Y","Z")[mid==min(mid)]`)
b. (`r c("X","Y","Z")[mid==max(mid)]`)
c. (`r c("X","Y","Z")[type==min(type)]`)
d. (`r c("X","Y","Z")[type==max(type)]`)

Answerlist
--------
* The distribution with its center sitting furthest left has the smallest mean.
* .
* .
* The distribution with its center sitting furthest right has the largest mean.
* .
* .
* The bell shape has the smallest standard deviation. It is this simple because all the distributions have the same range, but the bell has a higher fraction near its mean.
* .
* .
* The bimodal shape has the largest standard deviation. It is this simple because all the distributions have the same range, but the bimodal shape has a higher fraction far from its mean.
* .
* .


Meta-information
================
extype: cloze
exclozetype: schoice|schoice|schoice|schoice
exsolution: `r ans1`|`r ans2`|`r ans3`|`r ans4`
exname: compare_SD_same_width
extol: 0
exextra[numwidth]: 999.999