```{r,data,generation,echo=FALSE,results="hide"}
while(T){
  n = sample(8:20,1)
  p = runif(1,0.05,0.95)
  x = sample(c(0,1),n,T)
  phat = mean(x)
  devs = x-phat
  adevs = abs(devs)
  MAD = mean(adevs)
  i = 1:n
  mydf = data.frame(i,x,devs,adevs)
  colnames(mydf) = c("$$i$$","$$\\mathbf{x}$$","deviations = $\\mathbf{x}-\\hat{p}$","AbsDev = $\\vert\\mathbf{x}-\\hat{p}\\vert$")
  if(sum(x)>1 & sum(x)<(n-1) & round(phat,3)==phat){break}
}
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r paste(lapply(x,function(x) sprintf("%d",x)),collapse=", ")` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_MAD.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_MAD.csv).

Determine the sample's [mean absolute deviation (MAD) around the sample proportion](https://en.wikipedia.org/wiki/Average_absolute_deviation#Mean_absolute_deviation_around_a_central_point){target="blank"}.  You can round your answer to the [hundredths place](https://en.wikipedia.org/wiki/Orders_of_magnitude_(numbers)#10%E2%88%922){target="blank"}.


Solution
========

(ANS=`r MAD`)

First, determine the sample proportion (mean of 0s and 1s).

$$\hat{p} = \frac{`r sum(x)`}{`r n`} =`r mean(x)` $$

Determine the absolute deviations (distances from $\hat{p}$).

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>

```{r tab1, results="asis", echo=F}
knitr::kable(mydf,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```

Now, take the mean of the absolute deviations.

$$\text{MAD}_\text{prop} = \frac{\sum \vert\mathbf{x}-\hat{p}\vert}{n} =  `r MAD` $$




### spreadsheet

```{r echo=F}
xbar = mean(x)
sol = data.frame(i,x)

sol = rbind(sol,
            c("",""),
            c("","phat"),
            c("",paste0("=average(B2:B",n+1,")",collapse=""))
            )

deviations = c(paste0("=B",2:(n+1),"-B$",n+4),rep("",3))
absdev = c(paste0("=abs(C",2:(n+1),")"),
           "",
           "MAD_prop",
           paste0("=average(D2:D",n+1,")"))
sol = cbind(sol,deviations,absdev)
write.csv(sol,file="get_MAD_sol.csv", row.names=F)

```

You can do this with [a spreadsheet](get_MAD_sol.csv).

```{r fig.height=6,fig.width=8,echo=F}
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol)
```
\


### R

You can do this with R

```{r}
x = read.csv("get_MAD.csv")$x
phat = mean(x)
deviations = x-phat
AbsDev = abs(deviations)
MAD = mean(AbsDev)
x
phat
deviations
AbsDev
MAD
```


### Algebra

With a little bit of algebra, we can simplify the formula for $\text{MAD}_\text{prop}$. (I've not shown the algebra...)

$$\text{MAD}_\text{prop} = 2(1-\hat{p})\hat{p}$$

In other words, if $n_0$ is the number of 0s and $n_1$ is the number of 1s, then
$$\text{MAD}_\text{prop} = 2 \cdot \frac{n_0}{n} \cdot \frac{n_1}{n}$$
So, in this case:
$$\text{MAD}_\text{prop} = 2 \cdot \frac{`r sum(x==0)`}{`r n`}\cdot \frac{`r sum(x)`}{`r n`} = `r MAD`$$


Meta-information
================
extype: num
exsolution: `r MAD`
exname: get_MAD
extol: 0.01
exextra[numwidth]: 999.999