```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
n = sample(c(60,70,75,80,90,100,110,120,125,130,150),1)
beta1 = c(1,5,0.2,0.5,2)[type]
beta2 = c(1,5,0.2,2,0.5)[type]
bottom = sample(seq(20,70,5),1)
width = sample(seq(5,30,5),1)
i = 1:n
x = rbeta(n,beta1,beta2)*width+bottom
myhist = hist(x,breaks=7)
brks = myhist$breaks
counts = myhist$counts

b1 = sample(brks[3:(length(brks)-1)],1)
n1 = sum(x<b1)
b2 = sample(brks[2:(length(brks)-2)],1)
n2 = sum(x>b2)
b3 = sort(sample(brks[2:(length(brks)-1)],2))
c3 = mean(b3)
r3 = (b3[2]-b3[1])/2
n3 = sum(abs(x-c3)<r3)
b4 = sort(sample(brks[2:(length(brks)-1)],2))
c4 = mean(b4)
r4 = (b4[2]-b4[1])/2
n4 = sum(abs(x-c4)>r4)

b5 = sample(brks[3:(length(brks)-1)],1)
n5 = sum(x<b5)
b6 = sample(brks[2:(length(brks)-2)],1)
n6 = sum(x>b6)
b7 = sort(sample(brks[2:(length(brks)-1)],2))
c7 = mean(b7)
r7 = (b7[2]-b7[1])/2
n7 = sum(abs(x-c7)<r7)
b8 = sort(sample(brks[2:(length(brks)-1)],2))
c8 = mean(b8)
r8 = (b8[2]-b8[1])/2
n8 = sum(abs(x-c8)>r8)

```


Question
========

A random procedure generated `r n` measurements, which were organized into the [histogram](https://en.wikipedia.org/wiki/Histogram){target="blank"} shown below. You can assume the measurements are of a [continuous variable](https://en.wikipedia.org/wiki/Continuous_or_discrete_variable){target="blank"}, such that every measurement is in one of the intervals (and not on a break).

```{r echo=F, fig.width=5, fig.height=4}
ylim = max(counts)*1.1
hist(x,labels = TRUE,ylim=c(0,ylim),breaks=brks)
```
\


Answerlist
--------
* Evaluate $\#[\mathbf{x}<`r b1`]$.
* Evaluate $\#[\mathbf{x}>`r b2`]$.
* Evaluate $\#[|\mathbf{x}-`r c3`|<`r r3`]$.
* Evaluate $\#[|\mathbf{x}-`r c4`|>`r r4`]$.
* Find boundary $b$ such that $\#[\mathbf{x}<b] = `r n5`$.
* Find boundary $b$ such that $\#[\mathbf{x}>b] = `r n6`$.
* Find radius $r$ such that $\#[|\mathbf{x}-`r c7`|<r] = `r n7`$.
* Find radius $r$ such that $\#[|\mathbf{x}-`r c8`|>r] = `r n8`$.


Solution
========

[video](https://vimeo.com/531556037){target="_blank"}

The first 4 questions involve adding up frequencies of the indicated intervals. The last 4 questions can be done by guessing and checking until something works.



Answerlist
--------
* (ANS=`r n1`) There are `r n1` measurements less than `r b1`. So, $\#[\mathbf{x}<`r b1`]=`r n1`$.
* (ANS=`r n2`) There are `r n2` measurements more than `r b2`. So, $\#[\mathbf{x}>`r b2`]=`r n2`$.
* (ANS=`r n3`) There are `r n3` measurements closer than `r r3` units from `r c3`. So, $\#[|\mathbf{x}-`r c3`|<`r r3`]=`r n3`$.
* (ANS=`r n4`) There are `r n4` measurements further than `r r4` units from `r c4`. So, $\#[|\mathbf{x}-`r c4`|>`r r4`]=`r n4`$.
* (ANS=`r b5`) A boundary at $b=`r b5`$ has `r n5` measurements less than it. So, $\#[\mathbf{x}<`r b5`]=`r n5`$.
* (ANS=`r b6`) A boundary at $b=`r b6`$ has `r n6` measurements more than it. So, $\#[\mathbf{x}>`r b6`]=`r n6`$.
* (ANS=`r r7`) An interval with radius $r=`r r7`$ around center `r c7` contains `r n7` measurements. So, $\#[|\mathbf{x}-`r c7`|<`r r7`]=`r n7`$.
* (ANS=`r r8`) An interval with radius $r=`r r8`$ around center `r c8` excludes `r n8` measurements.  So, $\#[|\mathbf{x}-`r c8`|>`r r8`]=`r n8`$.



Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num
exsolution: `r n1`|`r n2`|`r n3`|`r n4`|`r b5`|`r b6`|`r r7`|`r r8`
exname: read_histogram
extol: 0
exextra[numwidth]: 999
