```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
while(T){
  n = 25
  beta1 = c(3,1,0.5,0.7,1.4)
  beta2 = c(3,1,0.5,1.4,0.7)
  type = sample(1:5,1)
  bot = sample(c(10,20,30,40,50,60,70),1)
  ran = sample(c(5,10,15,20,25,30),1)
  x = round(rbeta(n,beta1[type],beta2[type])*ran+bot,1)
  bp = boxplot(x,horizontal=T)
  if(length(bp$out)==0 & sum(duplicated(bp$stats))==0){
    break
  }
}

xs = bp$stats

s1 = sample(1:5,1)
x1 = xs[s1]
p1 = c(0,0.25,0.5,0.75,1)[s1] 

s2 = sample(1:5,1)
x2 = xs[s2]
p2 = c(1,0.75,0.5,0.25,0)[s2] 

s3 = sort(sample(1:5,2))
x3 = xs[s3]
p3 = c(0,0.25,0.5,0.75,1)[s3[2]]-c(0,0.25,0.5,0.75,1)[s3[1]]

s4 = sort(sample(1:5,2))
x4 = xs[s4]
p4 = c(0,0.25,0.5,0.75,1)[s4[2]]-c(0,0.25,0.5,0.75,1)[s4[1]]
c4 = mean(x4)
r4 = diff(x4)/2

s5 = sample(2:4,1)
x5 = xs[s5]
p5 = c(0,0.25,0.5,0.75,1)[s5] 

s6 = sample(2:4,1)
x6 = xs[s6]
p6 = c(1,0.75,0.5,0.25,0)[s6] 

```


Question
========

A sample was taken from an unknown population. The values were organized into a [boxplot](https://en.wikipedia.org/wiki/Box_plot){target="blank"}.

```{r echo=F,fig.width=6,fig.height=4}
bp = boxplot(x,horizontal=T,range=0)
for(stat in bp$stats){
  text(stat,1.3,stat,srt=90)
}
```
\

For simplicity, assume no measurements lie on the hinges, median, or whisker tips (so we do not worry about inclusive vs. exclusive boundaries). This assumption is approximately true with a very large sample from a continuous distribution.
 
Answerlist
--------
* What proportion of measurements are below `r x1`? $$\text{prop}[\mathbf{x}< `r x1`] = \text{?}$$
* What proportion of measurements are above `r x2`? $$\text{prop}[\mathbf{x}> `r x2`] = \text{?}$$
* What proportion of measurements are between `r x3[1]` and `r x3[2]`? $$\text{prop}[`r x3[1]`< \mathbf{x} < `r x3[2]`] = \text{?}$$
* What proportion of measurements are closer than `r r4` units from `r c4`? $$\text{prop}[|\mathbf{x}-`r c4`|<`r r4`] = \text{?}$$
* Determine boundary $b$ such that $\text{prop}[\mathbf{x}<b]=`r p5`$.
* Determine boundary $b$ such that $\text{prop}[\mathbf{x}>b]=`r p6`$.
* What is the range?
* What is the interquartile range (IQR)?
* What is the median?


Solution
========

[video](https://vimeo.com/531912276){target="_blank"}

You need to know that each region (whisker or half-box) contains 25% of the measurements.

Answerlist
--------
* `r p1`
* `r p2`
* `r p3`
* `r p4`
* `r x5`
* `r x6`
* `r round(xs[5]-xs[1],1)`. Subtract min from max. $$`r round(xs[5],1)`-`r round(xs[1],1)` = `r round(xs[5]-xs[1],1)`$$
* `r round(xs[4]-xs[2],1)`. Subtract $Q_1$ from $Q_3$. $$`r round(xs[4],1)`-`r round(xs[2],1)` = `r round(xs[4]-xs[2],1)`$$
* `r xs[3]`

Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num|num
exsolution: `r p1`|`r p2`|`r p3`|`r p4`|`r x5`|`r x6`|`r round(xs[5]-xs[1],1)`|`r round(xs[4]-xs[2],1)`|`r xs[3]`
exname: read_boxplot
extol: 0
exextra[numwidth]: 999.999
