```{r,data,generation,echo=FALSE,results="hide"}
while(T){
  type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
  n = sample(7:15,1)
  beta1 = c(1,3,0.2,0.5,2)[type]
  beta2 = c(1,3,0.2,2,0.5)[type]
  bottom = sample(seq(20,70,5),1)
  width = sample(seq(5,30,5),1)
  i = 1:n
  x = round(rbeta(n,beta1,beta2)*width+bottom,2)
  xbar = mean(x)
  devs = x-xbar
  sdevs = devs^2
  var = mean(sdevs)
  if(round(xbar,2)==xbar){
    break
  }
}

mydf = data.frame(i,x,devs,sdevs)
colnames(mydf) = c("$$i$$","$$\\mathbf{x}$$","deviations = $\\mathbf{x}-\\bar{x}$","SqrDev = $(\\mathbf{x}-\\bar{x})^2$")
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r paste(lapply(x,function(x) sprintf("%.2f",x)),collapse=", ")` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_VAR.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_VAR.csv).

Determine the biased sample [variance](https://en.wikipedia.org/wiki/Variance#Sample_variance){target="blank"} ([without Bessel correction](https://en.wikipedia.org/wiki/Bessel's_correction#Formula){target="blank"}).  You can round your answer to the [hundredths place](https://en.wikipedia.org/wiki/Orders_of_magnitude_(numbers)#10%E2%88%922){target="blank"}.


Solution
========

(ANS=`r var`)

[video](https://vimeo.com/531907796){target="_blank"}

First, determine the mean of the sample.

$$\bar{x} = \frac{\sum \mathbf{x}}{n} =`r mean(x)` $$

Determine the squared deviations (squared distances from $\bar{x}$).

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>

```{r tab1, results="asis", echo=F}
knitr::kable(mydf,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```

Now, take the mean of the squared deviations.

$$\text{VAR} = \frac{\sum (\mathbf{x}-\bar{x})^2}{n} = `r var` $$

### spreadsheet

```{r echo=F}
xbar = mean(x)
sol = data.frame(i,x)

sol = rbind(sol,
            c("",""),
            c("","xbar"),
            c("",paste0("=average(B2:B",n+1,")",collapse="")),
            c("","")
            )

deviations = c(paste0("=B",2:(n+1),"-B$",n+4),rep("",4))
sqrdev = c(paste0("=C",2:(n+1),"^2"),
           "",
           "VAR",
           paste0("=average(D2:D",n+1,")"),
           paste0("=VAR.P(B2:B",n+1,")"))
sol = cbind(sol,deviations,sqrdev)
write.csv(sol,file="get_VAR_sol.csv", row.names=F)

```

You can do this with [a spreadsheet](get_VAR_sol.csv).

```{r fig.height=6,fig.width=8,echo=F}
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol)
```
\

And, actually, you can skip a lot of work by using the [`VAR.P` function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_VARP_function){target="blank"}. Using the population variance formula is equivalent to using the biased sample variance formula.


### R

You can do this with R

```{r}
x = read.csv("get_VAR.csv")$x
xbar = mean(x)
deviations = x-xbar
sqrdev = deviations^2
VAR = mean(sqrdev)
x
xbar
deviations
sqrdev
VAR
# The built-in var() function almost works, but it is too fancy, and makes a Bessel correction. To use it, we need to undo the Bessel correction.
n = length(x)
var(x)*(n-1)/(n)
```


Meta-information
================
extype: num
exsolution: `r var`
exname: get_VAR
extol: 0.01
exextra[numwidth]: 999.999
