```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
while(T){
  mu = sample(seq(30,70,10),1)
  sig = sample(2:7,1)
  Z = round(rnorm(2),2)
  x = round(mu+sig*Z,2)
  Z2 = (x-mu)/sig
  if(sum(Z==Z2)==2 & Z[1]%%1!=0 & Z[2]%%1!=0){break}
}
```


Question
========

A [standard score](https://en.wikipedia.org/wiki/Standard_score){target="blank"} ($z$) can be calculated from a measurement ($x$), the population mean ($\mu$), and the population standard deviation ($\sigma$).
$$z = \frac{x-\mu}{\sigma} $$

The following (normal) spinner has population mean $\mu=`r mu`$ and population standard deviation $\sigma=`r sig`$.

```{r echo=F,fig.width=5,fig.height=5}
par(mar=c(0,0,0,0))
theta = seq(0,2*pi,0.01)
plot(sin(theta),cos(theta),type="l",lwd=4,axes=F,ann=F)
for(z in -3:3){
  p = pnorm(z)
  lines(c(0.8,1)*sin(p*2*pi),c(0.8,1)*cos(p*2*pi),lwd=2.5)
}
for(z in -2:2){
  p = pnorm(z)
  text(0.73*sin(p*2*pi),0.73*cos(p*2*pi),mu+z*sig)
}
for(z in (-3*sig):(3*sig)/sig){
  p = pnorm(z)
  lines(c(0.85,1)*sin(p*2*pi),c(0.85,1)*cos(p*2*pi))
}
for(z in (-30*sig):(30*sig)/10/sig){
  p = pnorm(z)
  lines(c(0.95,1)*sin(p*2*pi),c(0.95,1)*cos(p*2*pi))
}
# for(z in -3:3/2){
#   p = pnorm(z)
#   lines(c(0.85,1)*sin(p*2*pi),c(0.85,1)*cos(p*2*pi))
# }
# for(z in -30:30/10){
#   p = pnorm(z)
#   lines(c(0.90,1)*sin(p*2*pi),c(0.90,1)*cos(p*2*pi))
# }
# for(z in -300:300/100){
#   p = pnorm(z)
#   lines(c(0.97,1)*sin(p*2*pi),c(0.97,1)*cos(p*2*pi))
# }

polygon(0.05*cos(theta),0.05*sin(theta),col="red")
arrows(-0.5,-0.1,0.5,0.1,lwd=8)
arrows(-0.5,-0.1,0.5,0.1,lwd=7,col="red")
polygon(0.01*cos(theta),0.01*sin(theta),col="red")
```
\

Answerlist
--------
* If a spin has a measurement $x=`r x[1]`$, what is the standard score ($z$)?
* If a spin has a standard score $z=`r Z[2]`$, what is the measurement ($x$)?


Solution
========

Answerlist
--------
* Use the formula.
$$z = \frac{x-\mu}{\sigma} = \frac{`r x[1]`-`r mu`}{`r sig`} = `r Z[1]`$$
* Use the formula.
$$x = \mu+z\sigma = `r mu`+(`r Z[2]`)(`r sig`) = `r x[2]` $$


Meta-information
================
extype: cloze
exclozetype: num|num
exsolution: `r Z[1]`|`r x[2]`
exname: standard_score
extol: 0.01|1
exextra[numwidth,character]: 999.99