```{r data generation, echo = FALSE, results = "hide"}
ell = runif(1,0.65,0.97)
z = round(qnorm(ell),2)
ell = round(pnorm(z),4)
la = round(pnorm(-z),4)
sol = 2*la
```

Question
========
Determine the probability that the **absolute** standard normal variable is more than `r z`. In other words, evaluate $P\left(|Z| > `r z`\right)$.


Solution
========

(ANS=`r sol`)

[video](https://vimeo.com/531931017){target="_blank"}

First, you need to identify that we are looking for a two-tail area (the sum of the left and right tails). This is because $z$-scores far from 0 satisfy $|Z|>`r z`$, while $z$-scores near 0 do not satisfy the inequality.

Start with a sketch.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),c(dnorm(zs[zs<=-z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

Remember, the entire area is always 1. Most z-tables only provides left areas, so three methods are shown to get the two-tail area from a left-area (cumulative probability) table.  

**Method 1:** First find the area of the left tail.
$$P(Z<`r -z`) =  `r la`$$
Recognize normal distributions are symmetric, so we also know the area of the right tail.
$$P(Z>`r z`) = `r la` $$
The two areas add to our desired two-tail area.
$$\begin{aligned}
P\left(|Z|>`r z`\right) &= P(Z<`r -z`)+P(Z>`r z`) \\
&= `r la` + `r la` \\
&= `r sol`
\end{aligned}$$

This technique can be summarized with the following formula:
$$P\left(|Z|>z\right) = 2\cdot P(Z<-z)$$
assuming $z>0$.

So, in our case (when $z=`r z`$),
$$\begin{aligned}
P\left(|Z|<`r z`\right) &= 2\cdot P(Z<`r -z`) \\
&= 2\cdot`r la` \\
&= `r sol`
\end{aligned}$$

This method is shown graphically:

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),c(dnorm(zs[zs<=-z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/2,0.01),la)
text(xr,max(dnorm(xr)/2,0.01),la)
```
\

Notice we need to add both tails.

**Method 2:** You can achieve the same result by using the following formula:
$$P\left(|Z|>z\right) = 2\cdot\left[1-P(Z<z) \right] $$
So,
$$\begin{aligned}
P\left(|Z|>`r z`\right) &= 2\cdot\left[1-P(Z<`r z`) \right] \\
&= 2\cdot\left[1-`r ell`\right] \\
&= 2\cdot`r la`\\
&= `r sol`
\end{aligned}$$
```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),c(dnorm(zs[zs<=-z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),c(dnorm(zs[zs>=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,-z,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.42,0.42,0.4))
text((-3+z)/2,0.45,ell)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/2,0.01),la)
text(xr,max(dnorm(xr)/2,0.01),la)
```
\

It might be helpful to visualize with a spinner.

```{r echo=F,fig.width=6,fig.height=6}
par(mar=c(0,0,0,0))
theta=seq(0,2*pi,length.out=1000)
plot(sin(theta),cos(theta),type="l",ann=F,axes=F,xlim=c(-1.3,1.3),ylim=c(-1.3,1.3))
zs = -3:3
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.8,1),cos(p*2*pi)*c(0.8,1),lwd=2)
  pt = min(max(p,0.007),0.993)
  if(abs(zs[i])<3){
    text(0.75*sin(pt*2*pi),0.75*cos(pt*2*pi),zs[i])
  }
}
zs = seq(-2.5,2.5,1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.85,1),cos(p*2*pi)*c(0.85,1),lwd=2)
  if(abs(zs[i])<3){
    text(sin(p*2*pi)*0.8,cos(p*2*pi)*0.8,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.9,1),cos(p*2*pi)*c(0.9,1),lwd=1)
}
zs = seq(-2,2,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  if(zs[i]%%0.5 != 0){
    text(sin(p*2*pi)*0.85,cos(p*2*pi)*0.85,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.05)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.95,1),cos(p*2*pi)*c(0.95,1))
}
zs = seq(-3,3,0.01)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.98,1),cos(p*2*pi)*c(0.98,1))
}

for(p in seq(0.1,0.9,0.1)){
  lines(sin(p*2*pi)*c(1,1.05),cos(p*2*pi)*c(1,1.05))
  text(sin(p*2*pi)*1.1,cos(p*2*pi)*1.1,p,cex=0.8,srt=-(p*360-90)%%180-90)
}
for(p in seq(0.01,0.99,0.01)){
  lines(sin(p*2*pi)*c(1,1.03),cos(p*2*pi)*c(1,1.03))
}
for(p in seq(0.002,0.998,0.002)){
  lines(sin(p*2*pi)*c(1,1.01),cos(p*2*pi)*c(1,1.01))
}
lines(c(0,0),c(1,1.05),lwd=2)
text(0,1.09,"1 or 0",cex=0.8)
text(0,0,cex=1.5,"z")


ps = c(seq(pnorm(z),1,0.001),seq(0,pnorm(-z),0.001))
polygon(c(0,sin(ps*2*pi),0),c(0,cos(ps*2*pi),0),col=rgb(0.9,0.3,0.3,0.4),border=rgb(0.9,0.3,0.3))
pe = pnorm(c(-z,z))
pm = 0

lines(sin(ps*2*pi)*1.2,cos(ps*2*pi)*1.2,lwd=2)
lines(sin(pe[1]*2*pi)*c(1.15,1.2),cos(pe[1]*2*pi)*c(1.15,1.2),lwd=2)
lines(sin(pe[2]*2*pi)*c(1.15,1.2),cos(pe[2]*2*pi)*c(1.15,1.2),lwd=2)
text(0.5*sin(2*pi*pe[1]),0.5*cos(2*pi*pe[1]),-z,srt=(180-360*pe[1])%%180-90)
text(0.5*sin(2*pi*pe[2]),0.5*cos(2*pi*pe[2]),z,srt=(180-360*pe[2])%%180-90)
text(1.25*sin(2*pi*pe[1]),1.25*cos(2*pi*pe[1]),round(pe[1],4),srt=-(pe[1]*360-90)%%180-90,cex=0.8)
text(1.25*sin(2*pi*pe[2]),1.25*cos(2*pi*pe[2]),round(pe[2],4),srt=-(pe[2]*360-90)%%180-90,cex=0.8)
text(sin(pm*2*pi)*1.25,
     cos(pm*2*pi)*1.25,
     sprintf("%.4f",1-(round(pe[2],4)-round(pe[1],4))),
     srt=-(pm*360-90)%%180-90)
```
\

In a spreadsheet, you could use the `NORM.DIST()` function.

```
=2*NORM.DIST(`r -z`,0,1,TRUE)
```

In R, you could use the `pnorm` function.

```
2*pnorm(`r -z`)
```


Meta-information
================
extype: num
exsolution: `r sol`
exname: Get two-tail area from Z
extol: 0.002
