```{r data generation, echo = FALSE, results = "hide"}
while(T)
{
  ell = round(runif(1,0.01,0.99),2)
  ra = 1-ell
  z = qnorm(ell)
  zlow = floor(z*100)/100
  zhigh = ceiling(z*100)/100
  if(round(z,4) != round(z,2)) {break}
}
```

Question
========
Determine $z$ such that $P(Z>z)=`r ra`$. In other words, what $z$-score is less than $`r ra*100`$% of standard normal values? (Answers within 0.01 from the correct value will be marked correct.)


Solution
========

(ANS=`r z`)

[video](https://vimeo.com/531931611){target="_blank"}

Start with a sketch. Rightward numbers (toward $+\infty$) will be more than our boundary $z$, so we shade a rightward region with area `r ra`.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text((3+z)/2,max(dnorm((3+z)/2)/3,0.02),ra)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

You should first find the left area.
$$\begin{aligned}
P(Z<z) &= 1-P(Z>z) \\
&= 1-`r ra` \\
&= `r ell`
\end{aligned}$$

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text((-3+z)/2,max(dnorm((-3+z)/2)/3,0.02),ell)
text((3+z)/2,max(dnorm((3+z)/2)/3,0.02),ra)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\


```{r, echo=F}
z2 = seq(zlow-0.02,zhigh+0.02,0.01)
```

You should go to your $z$-table and find the $z$-score with the left area closest to `r ell`.

| $z$ | $P(Z<z)$ |
|:---:|:---:|
| `r z2[1]` | `r format(round(pnorm(z2[1]),4),4)` |
| `r z2[2]` | `r format(round(pnorm(z2[2]),4),4)` |
| **`r z2[3]`** | **`r format(round(pnorm(z2[3]),4),4)`** |
| **`r z2[4]`** | **`r format(round(pnorm(z2[4]),4),4)`** |
| `r z2[5]` | `r format(round(pnorm(z2[5]),4),4)` |
| `r z2[6]` | `r format(round(pnorm(z2[6]),4),4)` |

It turns out the exact answer is $z=`r z`$, which could be found by using an inverse normal function. On a spreadsheet:
```
=Norm.Inv(`r ell`,0,1)
```
Using R:
```
rightarea = `r ra`
leftarea = 1-rightarea
qnorm( leftarea )
```

But, the $z$-table is accurate enough, so I will accept either `r zlow` or `r zhigh` (anything within 0.01 of `r z`).

You might find it helpful to visualize with a spinner.

```{r echo=F,fig.width=6,fig.height=6}
par(mar=c(0,0,0,0))
theta=seq(0,2*pi,length.out=1000)
plot(sin(theta),cos(theta),type="l",ann=F,axes=F,xlim=c(-1.3,1.3),ylim=c(-1.3,1.3))
zs = -3:3
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.8,1),cos(p*2*pi)*c(0.8,1),lwd=2)
  pt = min(max(p,0.007),0.993)
  if(abs(zs[i])<3){
    text(0.75*sin(pt*2*pi),0.75*cos(pt*2*pi),zs[i])
  }
}
zs = seq(-2.5,2.5,1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.85,1),cos(p*2*pi)*c(0.85,1),lwd=2)
  if(abs(zs[i])<3){
    text(sin(p*2*pi)*0.8,cos(p*2*pi)*0.8,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.9,1),cos(p*2*pi)*c(0.9,1),lwd=1)
}
zs = seq(-2,2,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  if(zs[i]%%0.5 != 0){
    text(sin(p*2*pi)*0.85,cos(p*2*pi)*0.85,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.05)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.95,1),cos(p*2*pi)*c(0.95,1))
}
zs = seq(-3,3,0.01)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.98,1),cos(p*2*pi)*c(0.98,1))
}

for(p in seq(0.1,0.9,0.1)){
  lines(sin(p*2*pi)*c(1,1.05),cos(p*2*pi)*c(1,1.05))
  text(sin(p*2*pi)*1.1,cos(p*2*pi)*1.1,p,cex=0.8,srt=-(p*360-90)%%180-90)
}
for(p in seq(0.01,0.99,0.01)){
  lines(sin(p*2*pi)*c(1,1.03),cos(p*2*pi)*c(1,1.03))
}
for(p in seq(0.002,0.998,0.002)){
  lines(sin(p*2*pi)*c(1,1.01),cos(p*2*pi)*c(1,1.01))
}
lines(c(0,0),c(1,1.05),lwd=2)
text(0,1.09,"1 or 0",cex=0.8)
text(0,0,cex=1.5,"z")

pp = pnorm(z)
ps = seq(pp,1,0.001)
ps2 = seq(0,pp,0.001)
polygon(c(0,sin(ps*2*pi),0),c(0,cos(ps*2*pi),0),col=rgb(0.9,0.3,0.3,0.4),border=rgb(0.9,0.3,0.3,0))
pe = c(pp,1)
pe2 = c(0,pp)
pm = mean(pe)
pm2 = mean(pe2)

lines(sin(ps*2*pi)*1.2,cos(ps*2*pi)*1.2,lwd=2)
lines(sin(pe[1]*2*pi)*c(1.15,1.2),cos(pe[1]*2*pi)*c(1.15,1.2),lwd=2)
lines(sin(pe[2]*2*pi)*c(1.15,1.2),cos(pe[2]*2*pi)*c(1.15,1.2),lwd=2)
lines(sin(ps2*2*pi)*1.2,cos(ps2*2*pi)*1.2,lty=2,lwd=1)
text(0.4*sin(2*pi*pe[1]),0.4*cos(2*pi*pe[1]),round(z,7),srt=(180-360*pe[1])%%180-90)
text(0.4*sin(2*pi*pe[2]),0.4*cos(2*pi*pe[2]),"positve infinity",srt=(180-360*pe[2])%%180-90)
text(1.25*sin(2*pi*pe[1]),1.25*cos(2*pi*pe[1]),round(pe[1],4),srt=-(pe[1]*360-90)%%180-90,cex=0.8)
text(1.25*sin(2*pi*pe[2]),1.25*cos(2*pi*pe[2]),round(pe[2],4),srt=-(pe[2]*360-90)%%180-90,cex=0.8)
text(sin(pm*2*pi)*1.16,
     cos(pm*2*pi)*1.16,
     sprintf("%.4f",round(pe[2],4)-round(pe[1],4)),
     srt=-(pm*360-90)%%180-90)
text(sin(pm2*2*pi)*1.16,
     cos(pm2*2*pi)*1.16,
     sprintf("%.4f",1-(round(pe[2],4)-round(pe[1],4))),
     srt=-(pm2*360-90)%%180-90)
```
\


Meta-information
================
extype: num
exsolution: `r z`
exname: Get z from right area
extol: 0.011
