```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}

# reseed for exam question
set.seed(runif(1,3,7))

n = 100
i = 1:n
type = sample(1:2,1)
ns = c(8,10)[type]
x = sample(1:ns,n,T)
brks = 1:(ns+1)-0.5
myhist = hist(x,breaks=brks)
counts = myhist$counts
strdie = c("eight","ten")[type]

while(T){
  b1 = sample(brks[3:(length(brks)-1)],1)
  n1 = sum(x<b1)
  b2 = sample(brks[2:(length(brks)-2)],1)
  n2 = sum(x>b2)
  b3 = sort(sample(brks[2:(length(brks)-1)],2))
  c3 = mean(b3)
  r3 = (b3[2]-b3[1])/2
  n3 = sum(abs(x-c3)<r3)
  b4 = sort(sample(brks[2:(length(brks)-1)],2))
  c4 = mean(b4)
  r4 = (b4[2]-b4[1])/2
  n4 = sum(abs(x-c4)>r4)
  
  b5 = sample(brks[3:(length(brks)-1)],1)
  n5 = sum(x<b5)
  b6 = sample(brks[2:(length(brks)-2)],1)
  n6 = sum(x>b6)
  b7 = sort(sample(brks[2:(length(brks)-1)],2))
  c7 = mean(b7)
  r7 = (b7[2]-b7[1])/2
  if(r7==round(r7)){
    wrd7 = "integer"
  } else {
    wrd7 = "half-integer"
  }
  n7 = sum(abs(x-c7)<r7)
  b8 = sort(sample(brks[2:(length(brks)-1)],2))
  c8 = mean(b8)
  r8 = (b8[2]-b8[1])/2
  if(r8==round(r8)){
    wrd8 = "integer"
  } else {
    wrd8 = "half-integer"
  }
  n8 = sum(abs(x-c8)>r8)
  p1 = n1/n
  p2 = n2/n
  p3=n3/n
  p4=n4/n
  p5=n5/n
  p6=n6/n
  p7=n7/n
  p8=n8/n
  if(r3>1 & r4>1 & r7>1 & r8>1 & r7!=r8 & r7!=r3 & r8!=r4){break}
}
```

Question
========

A standard `r strdie`-sided [die](https://en.wikipedia.org/wiki/Dice) was rolled many times, and the results were organized into the [histogram](https://en.wikipedia.org/wiki/Histogram){target="blank"} shown below. 

```{r echo=F,fig.width=6,fig.height=5}
ylim = max(counts/n)*1.1
hist(x,breaks=brks,labels = TRUE,ylim=c(0,ylim),xlim=c(0,ns+1),freq=F,ylab="Relative Frequency")
```
\


Answerlist
--------
* Evaluate $\text{prop}[\mathbf{x}<`r b1`]$.
* Evaluate $\text{prop}[\mathbf{x}>`r b2`]$.
* Evaluate $\text{prop}[|\mathbf{x}-`r c3`|<`r r3`]$.
* Evaluate $\text{prop}[|\mathbf{x}-`r c4`|>`r r4`]$.
* Evaluate half-integer boundary $b$ such that $\text{prop}[\mathbf{x}<b] = `r n5/n`$.
* Evaluate half-integer boundary $b$ such that $\text{prop}[\mathbf{x}>b] = `r n6/n`$.
* Evaluate `r wrd7` radius $r$ such that $\text{prop}[|\mathbf{x}-`r c7`|<r] = `r n7/n`$.
* Evaluate `r wrd8` radius $r$ such that $\text{prop}[|\mathbf{x}-`r c8`|>r] = `r n8/n`$.


Solution
========


Answerlist
--------
* The answer is `r p1` because $\text{prop}[\mathbf{x}<`r b1`]=`r p1`$.
* The answer is `r p2` because $\text{prop}[\mathbf{x}>`r b2`]=`r p2`$.
* The answer is `r p3` because $\text{prop}[|\mathbf{x}-`r c3`|<`r r3`]=`r p3`$.
* The answer is `r p4` because $\text{prop}[|\mathbf{x}-`r c4`|>`r r4`]=`r p4`$.
* The answer is `r b5` because $\text{prop}[\mathbf{x}<`r b5`]=`r p5`$.
* The answer is `r b6` because $\text{prop}[\mathbf{x}>`r b6`]=`r p6`$.
* The answer is `r r7` because $\text{prop}[|\mathbf{x}-`r c7`|<`r r7`]=`r p7`$.
* The answer is `r r8` because $\text{prop}[|\mathbf{x}-`r c8`|>`r r8`]=`r p8`$.


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num
exsolution: `r n1/n`|`r n2/n`|`r n3/n`|`r n4/n`|`r b5`|`r b6`|`r r7`|`r r8`
exname: read_histogram_discrete_prop
extol: 0.001
exextra[numwidth]: 999.99
