```{r,data,generation,echo=FALSE,results="hide"}
type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
n = sample(7:15,1)
beta1 = c(1,5,0.2,0.5,2)[type]
beta2 = c(1,5,0.2,2,0.5)[type]
bottom = sample(seq(20,70,5),1)
width = sample(seq(5,30,5),1)
i = 1:n
x = round(rbeta(n,beta1,beta2)*width+bottom,2)
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r x` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_mean.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_mean.csv).


Determine $\bar{x}$, the [sample mean](https://en.wikipedia.org/wiki/Mean){target="blank"}. Your answer can be rounded to the nearest tenth.



Solution
========

(ANS=`r mean(x)`)

[video](https://vimeo.com/531905674){target="_blank"}

You need to sum the values ($\sum \mathbf{x}$, see [summation](https://en.wikipedia.org/wiki/Summation){target="blank"}) and divide by the sample size ($n$).

$$\bar{x} = \frac{\sum \mathbf{x}}{n} = \frac{`r sum(x)`}{`r length(x)`} = `r round(mean(x),4)`$$
You can round $\bar{x}$ to the nearest tenth: `r round(mean(x),1)`.

```{r echo=F,fig.width=6,fig.height=7}
sol = data.frame(i,x)
sol = rbind(sol,c("",""),c("","average(x)"),c("",paste0("=average(B2:B",n+1,")",collapse="")))
write.csv(sol,"mean_solution.csv",row.names=F)
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol)
```
\

In a spreadsheet, you can use the [`AVERAGE` function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_AVERAGE_function){target="blank"}. You can see a [solution spreadsheet here](mean_solution.csv).

In R, you can use the [`MEAN` function](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/mean){target="blank"}.

```{r}
# First, get the csv into the working directory... then...
data = read.csv("get_mean.csv")
x = data$x
xbar = mean(x)
round(xbar,1)
```



Meta-information
================
extype: num
exsolution: `r mean(x)`
exname: get_mean
extol: 0.1
exextra[numwidth]: 999.999
