
```{r data generation, echo = FALSE, results = "hide"}
library(exams)
options(scipen=100)
types = c("A","B","C","D")
fruit = c("apple","orange","banana","mango","plum","grapefruit","kiwi","lemon","papaya","apricot","persimmon","watermelon","pineapple","coconut")
mass = c(180,180,180,250,90,240,80,100,450,50,170,9000,2000,700)
is = sample(1:length(mass),4)
while(T){
  is = sample(1:length(mass),4)
  f = fruit[is]
  fs = paste(f,"s",sep="")
  mus = round(mass[is]*runif(4,0.7,1.3))
  sigs = round(mus/runif(4,6,20))
  ps = runif(4,0.02,0.98)
  zs = round(qnorm(ps),2)
  xs = signif(mus+zs*sigs,4)
  zs = round((xs-mus)/sigs,2)
  ps = round(pnorm(zs),4)
  if(sum(zs>0)>=1 & sum(zs<0)>=1 & length(unique(abs(zs)))==4 & min(c(dist(abs(zs),method = "manhattan"))) > 0.1){
    break
  }
}

scs1 = mchoice2string(zs==max(zs))
scs2 = mchoice2string(zs==min(zs))
scs3 = mchoice2string(abs(zs)==min(abs(zs)))
scs4 = mchoice2string(abs(zs)==max(abs(zs)))


data = data.frame(f,mus,sigs,xs)
colnames(data) = c("Fruit type",
                    "Population mean",
                    "Population standard deviation",
                    "Measurement")

write.csv(data,"fruit.csv",row.names=F)

tab1 = data.frame(fs,mus,sigs)
colnames(tab1) = c("Type of fruit","$\\mu$, population mean mass in grams","$\\sigma$, population standard deviation in grams")
tab2 = data.frame(f,xs)
colnames(tab2) = c("Specimen type","$x$, mass of specimen in grams")
```

Question
========

A farm produces 4 types of fruit: `r fs[1]`, `r fs[2]`, `r fs[3]`, and `r fs[4]`. The fruits' masses follow normal distributions, with population parameters dependent on the type of fruit.

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>

```{r tab1, results="asis",echo=F}
knitr::kable(tab1,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```


One specimen of each type is weighed. The results are shown below.

```{r tab2, results="asis",echo=F}
knitr::kable(tab2,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```

The population parameters and specimen masses can be [downloaded as a csv](fruit.csv){target="blank"}.

For each measurement, determine the [standard score](https://en.wikipedia.org/wiki/Standard_score) and the [cumulative probability](https://en.wikipedia.org/wiki/Cumulative_distribution_function){target="blank"}. Then determine which specimen is most unusually large, most unusually small, most typically sized, and most unusually sized.


a. ##ANSWER1##
b. ##ANSWER2##
c. ##ANSWER3##
d. ##ANSWER4##
e. ##ANSWER5##
f. ##ANSWER6##
g. ##ANSWER7##
h. ##ANSWER8##
i. Which specimen is the most unusually large? ##ANSWER9##
j. Which specimen is the most unusually small? ##ANSWER10##
k. Which specimen is most typically sized? ##ANSWER11##
l. Which specimen is most unusually sized? ##ANSWER12##


Answerlist
--------
* Calculate a $z$ score for the `r f[1]` specimen. (Round to the nearest hundredth.)
* What proportion of `r fs[1]` have less mass than the `r f[1]` specimen?
* Calculate a $z$ score for the `r f[2]` specimen. (Round to the nearest hundredth.)
* What proportion of `r fs[2]` have less mass than the `r f[2]` specimen?
* Calculate a $z$ score for the `r f[3]` specimen. (Round to the nearest hundredth.)
* What proportion of `r fs[3]` have less mass than the `r f[3]` specimen?
* Calculate a $z$ score for the `r f[4]` specimen. (Round to the nearest hundredth.)
* What proportion of `r fs[4]` have less mass than the `r f[4]` specimen?
* The `r f[1]` specimen is most unusually large.
* The `r f[2]` specimen is most unusually large.
* The `r f[3]` specimen is most unusually large.
* The `r f[4]` specimen is most unusually large.
* The `r f[1]` specimen is most unusually small.
* The `r f[2]` specimen is most unusually small.
* The `r f[3]` specimen is most unusually small.
* The `r f[4]` specimen is most unusually small.
* The `r f[1]` specimen is most typically sized.
* The `r f[2]` specimen is most typically sized.
* The `r f[3]` specimen is most typically sized.
* The `r f[4]` specimen is most typically sized.
* The `r f[1]` specimen is most unusually sized.
* The `r f[2]` specimen is most unusually sized.
* The `r f[3]` specimen is most unusually sized.
* The `r f[4]` specimen is most unusually sized.

Solution
========

[video](https://vimeo.com/531932946){target="_blank"}

The formula to determine the $z$-score of a **measurement** is the ratio with numerator the difference between measurement and population mean and denominator the population standard deviation.

$$z = \frac{x-\mu}{\sigma} $$

The highest $z$-score (furthest right on [number line](https://en.wikipedia.org/wiki/Number_line){target="blank"}) corresponds to the most unusually large measurement. 

The smallest $z$-score (furthest left on [number line](https://en.wikipedia.org/wiki/Number_line){target="blank"}) corresponds to the most unusually small measurement.

The smallest [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$-score corresponds to the most usually sized measurement.

The largest [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$-score corresponds to the most unusually sized measurement.


Answerlist
--------
* We use the formula for calculating the `r f[1]` specimen's $z$ score.
$$z = \frac{`r xs[1]`-`r mus[1]`}{`r sigs[1]`} = `r zs[1]` $$
* Use the $z$ table to find the cumulative probability
$$P(Z<`r zs[1]`) = `r ps[1]` $$
* We use the formula for calculating the `r f[2]` specimen's $z$ score.
$$z = \frac{`r xs[2]`-`r mus[2]`}{`r sigs[2]`} = `r zs[2]` $$
* Use the $z$ table to find the cumulative probability
$$P(Z<`r zs[2]`) = `r ps[2]` $$
* We use the formula for calculating the `r f[3]` specimen's $z$ score.
$$z = \frac{`r xs[3]`-`r mus[3]`}{`r sigs[3]`} = `r zs[3]` $$
* Use the $z$ table to find the cumulative probability
$$P(Z<`r zs[3]`) = `r ps[3]` $$
* We use the formula for calculating the `r f[4]` specimen's $z$ score.
$$z = \frac{`r xs[4]`-`r mus[4]`}{`r sigs[4]`} = `r zs[4]` $$
* Use the $z$ table to find the cumulative probability
$$P(Z<`r zs[4]`) = `r ps[4]` $$
* We determine the maximum $z$ score is $`r max(zs)`$, which belongs to the `r f[zs==max(zs)]`.
* .
* .
* .
* We determine the minimum $z$ score is $`r min(zs)`$, which belongs to the `r f[zs==min(zs)]`.
* .
* .
* .
* We determine the minimum [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$ score is $`r min(abs(zs))`$, which belongs to the `r f[abs(zs)==min(abs(zs))]`. 
* .
* .
* .
* We determine the maximum [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$ score is $`r max(abs(zs))`$, which belongs to the `r f[abs(zs)==max(abs(zs))]`.
* .
* .
* .


Meta-information
============
extype: cloze
exclozetype: num|num|num|num|num|num|num|num|schoice|schoice|schoice|schoice
exsolution: `r zs[1]`|`r ps[1]`|`r zs[2]`|`r ps[2]`|`r zs[3]`|`r ps[3]`|`r zs[4]`|`r ps[4]`|`r scs1`|`r scs2`|`r scs3`|`r scs4`
exname: compare zs
extol: 0.011|0.004|0.011|0.004|0.011|0.004|0.011|0.004|0|0|0|0
