```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)

while(T){
  n = sample(10:40,2)
  n1 = n[1]
  n2 = n[2]
  sigma1 = runif(1,10,50)
  sigma2 = sigma1*runif(1,0.6,1/0.6)
  se = sqrt(sigma1^2/n[1]+sigma2^2/n[2])
  me = 2*se
  mu1 = runif(1,200,600)
  mu2 = mu1+sample(c(-1,1),1)*me
  x1 = round(rnorm(n1,mu1,sigma1),1)
  x2 = round(rnorm(n2,mu2,sigma2),1)
  xbar1 = mean(x1)
  xbar2 = mean(x2)
  xbar1r = round(xbar1,1)
  xbar2r = round(xbar2,1)
  s1 = sd(x1)
  s2 = sd(x2)
  s1r = round(s1,1)
  s2r = round(s2,1)
  df = (s1^2/n1+s2^2/n2)^2/(s1^4/n1^2/(n1-1)+s2^4/n2^2/(n2-1))
  t = abs(xbar1-xbar2)/sqrt(s1^2/n1+s2^2/n2)
  cump = pt(t,df)
  pval = 2*(1-cump)
  if(pval<0.2 & pval>0.005 & abs(pval-0.05)>0.01){break}
}

x1rep = x1
x2rep = x2

if(pval<0.05){
  ans2="10"
  s2="The $p$-value is less than 0.05, so the result is significant."
} else {
  ans2="01"
  s2="The $p$-value is more than 0.05, so the result is NOT significant."
}

default_source_hook <- knitr::knit_hooks$get('source')
knitr::knit_hooks$set(source = function(xx, options) {
  xx <- gsub('x1rep',paste(x1,collapse=","),xx)
  xx <- gsub('x2rep',paste(x2,collapse=","),xx)
  default_source_hook(xx, options)
})

```

Question
========

A doctor runs a [controlled experiment](https://en.wikipedia.org/wiki/Randomized_controlled_trial){target="blank"}. The participants are randomly assigned to two groups: control and treatment. The participants in the control group are given a [placebo](https://en.wikipedia.org/wiki/Placebo-controlled_study){target="blank"}. The participants in the treatment group are given a drug.

```{r}
len = max(n1,n2)
if(n1>n2){
  x1d = x1
  x2d = c(x2,rep(NA,n1-n2))
} else {
  x2d = x2
  x1d = c(x1,rep(NA,n2-n1))
}
data = data.frame(x1d,x2d)
colnames(data) = c("x1","x2")
write.csv(data,file="triglyceride.csv",row.names=F,na = "")
```

After a month, each participant's triglyceride level (in mg/dL) is measured. Those measurements are shown below. They can also be [downloaded as a csv](triglyceride.csv){target="blank"}.

```{r}
cat(
sprintf("Control: %s\n\nTreatment: %s",
        paste0(x1,collapse = ", "),
        paste0(x2,collapse = ", ")
        )
)
```

You are asked to perform a two-tail [two-sample Welch's $t$ test](https://en.wikipedia.org/wiki/Welch%27s_t-test){target="blank"} to determine whether there is a significant difference of means in the two samples.

To do this by hand, you would first determine the absolute $t$ score as defined here.
$$|t| = \frac{|\overline{x}_1-\overline{x}_2|}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}}$$
You'd also need to calculate the degrees of freedom.

$$\text{df} = \frac{\left(\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}\right)^2}{\frac{s_1^4}{n_1^2(n_1-1)}+\frac{s_2^4}{n_2^2(n_2-1)}}$$
And then, the $p$-value:

$$p\text{-value} = P\left(|T|>|t|\right) $$

However, this problem is easy when using a spreadsheet or R, so I would recommend using one of those tools. 

In a spreadsheet, you can use [`T.TEST`](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_TTEST_function){target="blank"} with mode=2 for a two-tail test and type=3 for Welch's test.

In R, you can use [`t.test`](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/t.test){target="blank"} with the default settings. 

##ANSWER1##

Is the difference of means significant (using a significance level of 0.05)?

##ANSWER2##

Answerlist
--------
* Determine the $p$-value.
* Yes, the drug causes a difference in average triglyceride level.
* No, we don't know whether the drug causes a difference.


Solution
========

[video](https://vimeo.com/531954367){target="_blank"}

To do this by hand, you first calculate the sample statistics.
$$n_1 = `r n1`$$
$$n_2 = `r n2`$$
$$\bar{x}_1 = `r xbar1r`$$
$$\bar{x}_2 = `r xbar2r`$$
$$s_1 = `r s1r`$$
$$s_2 = `r s2r`$$
$$|t| = `r round(t,2)`$$
$$\text{df} = `r round(df,3)`$$
You will get quite close if you round df down ([floor](https://en.wikipedia.org/wiki/Floor_and_ceiling_functions){target="blank"}).

Then, using a computer application or a $t$ table, you should be able to determine the following probabilities (using interpolation to estimate if using $t$ table).
$$P(T<`r round(t,2)`) = `r round(cump,4)` $$
$$P(|T|>`r round(t,2)`) = `r round(pval,4)` $$
$$p\text{-value} = `r round(pval,4)` $$

### Spreadsheet

```{r}
spacer=rep("",length(x1d))
eq = c("","=T.TEST(A2:A1000,B2:B1000,2,3)",rep("",length(x1d)-2))
sol = data.frame(x1d,x2d,spacer,eq)
colnames(sol)=c("x1","x2","","")
write.csv(sol,"triglyceride_solution.csv",row.names=F)
```

You just need to use `T.TEST` with the proper settings. You can [download the solution as a csv](triglyceride_solution.csv){target="blank"}.

```{r fig.height=5,fig.width=6}
draw_ss = function(df,widths=rep(1,length(df[1,])+1)){
  xline = c(0,cumsum(widths)/sum(widths))
  nrow = length(df[,1])
  ncol = length(df[1,])
  xtex = (xline[2:(ncol+2)]+xline[1:(ncol+1)])/2
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  #xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=xline)
  abline(v=xline[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}
draw_ss(sol[1:9,],c(0.4,1,1,0.5,3))
```
\

### R

You just use `t.test` with the default settings. The hardest part is getting the data imported. You can do this in 2 ways: copy/paste or `read.csv`.

#### copy/paste

```{r copypaste, echo=T}
x1 = c(x1rep)
x2 = c(x2rep)
t.test(x1,x2)
```

#### read.csv

Download `triglyceride.csv` and move it to a directory (folder). Make a script, `welchttest.r`, and put it in the same directory. Set the working directory to this directory. Then, run the script. 

```{r csv, echo=T}
### welchttest.r
data = read.csv("triglyceride.csv")
x1 = data$x1
x2 = data$x2
t.test(x1,x2)
```


Meta-information
================
extype: cloze
exsolution: `r pval`|`r ans2`
exclozetype: num|schoice
exname: hypothesis test 2 means
extol: 0.005|0
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE
