```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
n = sample(c(60,70,75,80,90,100,110,120,125,130,150),1)
beta1 = c(1,10,0.2,0.5,2)[type]
beta2 = c(1,10,0.2,2,0.5)[type]
bottom = sample(seq(20,70,5),1)
width = sample(seq(5,30,5),1)
i = 1:n
x = rbeta(n,beta1,beta2)*width+bottom
myhist = hist(x,breaks=5)
brks = myhist$breaks
counts = myhist$counts

library(knitr)
library(stringr)
default_source_hook <- knit_hooks$get('source')

knit_hooks$set(source = function(x, options) {
  x <- str_replace_all(x, pattern = 'brks', paste0("c(",paste0(brks,collapse=","),")"))
  default_source_hook(x, options)
})
```


Question
========

```{r echo=F}
write.csv(data.frame(i,x),file="make_freq_dist.csv",row.names = FALSE)
```

A random procedure generated many measurements: [download data](make_freq_dist.csv)

Please complete the [frequency distribution](https://en.wikipedia.org/wiki/Frequency_distribution){target="blank"} using breaks `r brks`:

```{r echo=F}
s = ''
for(i in 1:length(counts)){
  s = paste0(s,"|",brks[i]," to ",brks[i+1],"|\\#\\#ANSWER",i,"\\#\\#|\n")
}
```

| Interval |  Frequency  |
|:---:|:---:|
`r s`

Solution
========

You will want to use a computer to answer these questions. 

In a spreadsheet, open the data, add the breaks as a column; then, use the [``FREQUENCY function``](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_FREQUENCY_function){target="blank"}.

In R, open the data and use the [``hist``](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/hist) function. You supply the breaks and read the counts:

```{r fig.show='hide'}
mydata = read.csv("make_freq_dist.csv")
x = mydata$x
myhist = hist(x,breaks=brks)
myhist$counts
```

```{r echo=F}
s = ''
for(i in 1:length(counts)){
  s = paste0(s,"|",brks[i]," to ",brks[i+1],"|",counts[i],"|\n")
}
```

| interval |  frequency  |
|:---:|:---:|
`r s`


Meta-information
================
extype: cloze
exclozetype: `r paste0(rep("num",length(counts)),collapse="|")`
exsolution: `r paste0(counts,collapse="|")`
exname: make_frequency_distribution
extol: 0
exextra[numwidth]: 999
