```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)
n = sample(30:200,1)
la = runif(1,0.01,0.99)
mag = sample(-1:2,1)

shape1 = 3^runif(1,-0.5,1)
shape2 = 3^runif(1,-0.5,1)
while(T){
  ab = sort(sample(1:100,2))
  a = ab[1]
  b = ab[2]
  w = b-a
  if(w>3){break}
}

mu = round(w*(shape1)/(shape1+shape2)+a,2)
sig = round(w*sqrt(shape1*shape2/(shape1+shape2)^2/(shape1+shape2+1)),2)

mut = mu*n
sigt = sig*sqrt(n)
zapp = round(qnorm(la),2)
xt = signif(round(mut+zapp*sigt,2),3)
z = (xt-mut)/sigt
zlo = floor(z*100)/100
zhi = ceiling(z*100)/100
la = pnorm(z)
lalo = round(pnorm(zlo),4)
lahi = round(pnorm(zhi),4)

ra = 1-la
ralo = round(1-lalo,4)
rahi = round(1-lahi,4)

tol = max(abs(rahi-ralo)*1.1,0.002)

```

Question
========

The continuous random variable $X$ follows the distribution shown by the density curve and spinner below. It has a mean of $\mu_{X}=`r mu`$ and standard deviation of $\sigma_{X}=`r sig`$.

```{r, fig.width=6,fig.height=3}
x = seq(a-1,b+1,length.out=2000)
plt = plot(x,dbeta((x-a)/w,shape1,shape2)/w,type='l',ylab = "density",xlab="X")
polygon(c(x,a),c(dbeta((x-a)/w,shape1,shape2)/w,0),col=rgb(0.8,0.1,0.1,0.5))
```
\

```{r, fig.width=6,fig.height=6}
theta = seq(0,2*pi,length.out=1000)
x = cos(theta)
y = sin(theta)
par(mar=c(0,0,0,0))
plot(0,0,type="n",axes=F,ann=F,xlim=c(-1,1),ylim=c(-1,1))
polygon(x,y)
large_step = 10^floor(log(w)/log(10))
med_step = large_step/2
small_step = med_step/5

lines(c(0.75,1),c(0,0),lwd=4)
text(0.65,0,b)
lines(c(-0.75,-1),c(0,0),lwd=4)
text(-0.65,0,a)

for(i in seq(0,100,small_step)){
  if(a<i && i<b && i%%large_step==0){
    p = pbeta((i-a)/w,shape1,shape2)
    lines(c(-0.85*cos(pi*p),-cos(pi*p)),c(0.85*sin(pi*p),sin(pi*p)),lwd=2)
    lines(c(-0.85*cos(pi*p),-cos(pi*p)),c(-0.85*sin(pi*p),-sin(pi*p)),lwd=2)
    if(p>0.02 && p<0.98){
      text(-0.75*cos(pi*p),0.75*sin(pi*p),i)
      text(-0.75*cos(pi*p),-0.75*sin(pi*p),i)
    }
  } else if(a<i && i<b && i%%med_step==0){
    p = pbeta((i-a)/w,shape1,shape2)
    lines(c(-0.90*cos(pi*p),-cos(pi*p)),c(0.90*sin(pi*p),sin(pi*p)),lwd=1)
    lines(c(-0.90*cos(pi*p),-cos(pi*p)),c(-0.90*sin(pi*p),-sin(pi*p)),lwd=1)
  } else if(a<i && i<b && i%%small_step==0){
    p = pbeta((i-a)/w,shape1,shape2)
    lines(c(-0.95*cos(pi*p),-cos(pi*p)),c(0.95*sin(pi*p),sin(pi*p)),lwd=1)
    lines(c(-0.95*cos(pi*p),-cos(pi*p)),c(-0.95*sin(pi*p),-sin(pi*p)),lwd=1)
  }
}

arrows(-0.6,-0.1,0.6,0.1,lwd=3)
text(0,0.2,"Random Variable X")
```
\

That spinner ($X$) will be fairly spun `r n` times, and the **total** of the spins will be recorded. Determine the probability that the random total is more than `r xt`.
$$P\left(\sum{X}>`r xt`\right) ~=~ \,? $$

(Even though single spins are **not** normally distributed, the random totals will be normally distributed.
See [central limit theorem](https://en.wikipedia.org/wiki/Central_limit_theorem){target="_blank"}.)


Solution
========

[video](https://vimeo.com/531942180){target="_blank"}

We use the central limit formulas for a random total.
$$\mu_{\sum{X}} ~=~ n\cdot\mu_{X} ~=~ (`r n`)(`r mu`) ~=~ `r mut` $$
$$\sigma_{\sum{X}} = \sigma_{X}\sqrt{n} = (`r sig`)\sqrt{`r n`} = `r sigt`$$
So, we think the random total is normally distributed with a mean of `r mut` and a standard deviation of `r sigt`.
$$\sum{X} \sim N(`r mut`,\,`r sigt`) $$

Calculate a $z$-score for the boundary.
$$z = \frac{`r xt`-`r mut`}{`r sigt`} = `r z` $$

You can round that to $z=`r zlo`$ or $z=`r zhi`$. Any of the following probabilities will get credit.
$$P(Z > `r z`) = `r ra` $$
$$P(Z > `r zlo`) = `r ralo` $$
$$P(Z > `r zhi`) = `r rahi` $$


Let's draw a sketch.

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mut-3*sigt,mut+3*sigt,length.out=7),4),line=3)
mtext(TeX("$\\sum X$ axis"),1,3,at=-3.5)
abline(h=0,lwd=3)
abline(h=seq(0,0.4,0.05),lwd=0.5)
abline(v=seq(-3,3,0.2),lwd=0.5)

polygon(c(Z[Z>=z],3,z),c(dnorm(Z[Z>=z]),0,0),col=rgb(1,0.5,0.5,0.5))

text(2.01,0.32,TeX(sprintf("$P(%s>%s) = %.4f$","\\sum X",xt,round(1-pnorm(z),4))))
arrows(1.6,0.3,(2*z+3)/3,max(dnorm((2*z+3)/3)/2,0.03),lwd=2)
```
\



Meta-information
================
extype: num
exsolution: `r ra`
exname: prob tot more
extol: `r tol`
