```{r,data,generation,echo=FALSE,results="hide"}
# type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
# n = sample(7:15,1)
# beta1 = c(1,5,0.2,0.5,2)[type]
# beta2 = c(1,5,0.2,2,0.5)[type]
# bottom = sample(seq(20,70,5),1)
# width = sample(seq(5,30,5),1)
# i = 1:n
# x = round(rbeta(n,beta1,beta2)*width+bottom,2)
x = c(57.1,46.89,50.16,59.8,52.81,57.94,50.84,56.87,45.47,46.18,56.06,48.76,58.13,53.48,51.64)
n = length(x)
i = 1:n
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r x` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_median.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_median.csv).

Determine the sample [median](https://en.wikipedia.org/wiki/Median){target="blank"}. Please enter an exact answer.



Solution
========

[video](https://vimeo.com/531906339){target="_blank"}

(ANS=`r median(x)`)

To determine the median by hand, you first sort the sample.
$$\text{sort}(\mathbf{x})=`r sort(x)`$$
If the sample size is odd, just find the number in the middle number (the $\frac{n+1}{2}$th value). If the sample size is even, determine the mean of the middle two numbers (the $\frac{n}{2}$th and $\left(\frac{n}{2}+1\right)$th values).


```{r echo=F,fig.height=8,fig.width=6}
sol = data.frame(i,x)
sol = rbind(sol,c("",""),c("","median(x)"),c("",paste0("=median(B2:B",n+1,")",collapse="")))
write.csv(sol,"median_solution.csv",row.names=F)
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol)
```
\

In a spreadsheet, you can use the [`MEDIAN` function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_MEDIAN_function){target="blank"}. You can see a [solution spreadsheet here](median_solution.csv).

In R, you can use the [`MEDIAN` function](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/median){target="blank"}.

```{r}
# First, get the csv into the working directory... then...
data = read.csv("get_median.csv")
x = data$x
median(x)
```



Meta-information
================
extype: num
exsolution: `r median(x)`
exname: get_median
extol: 0
exextra[numwidth]: 999.999
