```{r,data,generation,echo=FALSE,results="hide"}
type = 1 # sample(1:5,1,prob=c(0.1,0.1,0.1,0.35,0.35)) #uniform, bell, bimodal, skewright, skewleft
n = 1000
beta1 = c(1,5,0.2,0.5,2)[type]
beta2 = c(1,5,0.2,2,0.5)[type]
bottom = sample(seq(20,70,5),1)
width = sample(seq(5,30,5),1)
i = 1:n
x = round(rbeta(n,beta1,beta2)*width+bottom,2)

if(type < 4){
  ans = "100"
  wrd = "symmetric"
  eq = "$\\text{mean} \\approx \\text{median}$"
} else if(type == 4) {
  ans = "010"
  wrd = "skew-right"
  eq = "$\\text{mean} > \\text{median}$"
} else {
  ans = "001"
  wrd = "skew-left"
  eq = "$\\text{mean} < \\text{median}$"
}

```


Question
========

Depending on the type of distribution, we can make a strong claim regarding the mean and median.

* Symmetric (bell, uniform, bimodal)
  - $\text{mean} \approx \text{median}$
* Skew-right
  - $\text{mean} > \text{median}$
* Skew-left
  - $\text{mean} < \text{median}$

A sample was gathered and visualized with a histogram.

```{r, echo=F,fig.width=5,fig.height=4}
hist(x)
```
\

What claim can you make regarding the mean and median?

Answerlist
--------
* $\text{mean} \approx \text{median}$
* $\text{mean} > \text{median}$
* $\text{mean} < \text{median}$


Solution
========

[video](https://vimeo.com/531906554){target="_blank"}

The distribution is `r wrd`, so `r eq`.

Answerlist
--------
* `r ans=="100"`
* `r ans=="010"`
* `r ans=="001"`

Meta-information
================
extype: schoice
exsolution: `r ans`
exname: compare_mean_median
extol: 0
exextra[stringwidth]: 9
