```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
while(T){
  type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
  n = sample(c(60,75,80,90,100,110,120,125,130,150),1)
  beta1 = c(1,10,0.2,0.5,2)[type]
  beta2 = c(1,10,0.2,2,0.5)[type]
  bottom = sample(seq(20,70,5),1)
  width = sample(seq(5,30,5),1)
  i = 1:n
  x = rbeta(n,beta1,beta2)*width+bottom
  myhist = hist(x,breaks=5)
  brks = myhist$breaks
  counts = myhist$counts
  relf = counts/n
  dens = relf/(brks[2]-brks[1])
  if(brks[2]-brks[1] != 1){
    break
  }
}

sol = signif(c(rbind(relf,dens)),6)
soln = paste0(sol,collapse="|")
tol = paste0(sol/10,collapse="|")

```


Question
========

```{r echo=F}
write.csv(data.frame(i,x),file="make_freq_dist.csv",row.names = FALSE)
```

A random procedure generated $n=`r n`$ measurements: [download data](make_freq_dist.csv)

I've already determined the frequencies. Please determine the relative frequencies and the densities. A brief description of relative frequency and density can be found [here](https://en.wikipedia.org/wiki/Histogram){target="blank"}.

```{r echo=F}
s = ''
for(i in 1:length(counts)){
  s = paste0(s,"|",brks[i]," to ",brks[i+1],"|",counts[i],"|\\#\\#ANSWER",i*2-1,"\\#\\# ","| \\#\\#ANSWER",i*2,"\\#\\#|\n")
}
```

| Interval |  Frequency  | Relative Frequency | Density |
|:---:|:---:|:---:|:---:|
`r s`


Solution
========

To determine the relative frequencies, just divide each frequency by `r n` (because $n=`r n`$). To determine the densities, divide the relative frequencies by the width of the interval, which in this case is the same for each interval ($`r brks[2]`-`r brks[1]`=`r brks[2]-brks[1]`$).

```{r echo=F}
s = ''
for(i in 1:length(counts)){
  s = paste0(s,"|",brks[i]," to ",brks[i+1],"|",counts[i],"|",signif(counts[i]/n,4),"|",signif(counts[i]/n/(brks[i+1]-brks[i]),4),"|\n")
}
```

| Interval |  Frequency  | Relative Frequency | Density |
|:---:|:---:|:---:|:---:|
`r s`



Meta-information
================
extype: cloze
exclozetype: `r paste0(rep("num",length(counts)*2),collapse="|")`
exsolution: `r soln`
exname: relfreq_and_dens
extol: `r tol`
exextra[numwidth,character]: 9999
