```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
while(T){
  la = runif(1,0.01,0.99)
  z = round(qnorm(la),2)
  mag = sample(-1:2,1)
  mu = signif(sample(10:99,1)/10 * 10^mag,2)
  sig = round(runif(1,mu/20,mu/3),-mag+1)
  lets = strsplit("ABCDEFGHIJKMQRVWXY",split=NULL)
  rvi = sample(1:length(lets[[1]]),size=1)
  rv = lets[[1]][rvi]
  xl = tolower(rv)
  x = round(z*sig+mu,-mag+1)
  z = (x-mu)/sig
  zr = round(z,2)
  la = round(pnorm(z),4)
  if(sig != 0 & z == zr & z != 0){break}
}
```

Question
========
Random variable $`r rv`$ is [normally distributed](https://en.wikipedia.org/wiki/Normal_distribution){target="blank"} with mean $\mu = `r mu`$ and standard deviation $\sigma = `r sig`$.
$$`r rv` = \text{N}(`r mu`,\,`r sig`) $$
Evaluate $P(`r rv` < `r x`)$.


Solution
========

[video (using z-table)](https://vimeo.com/531933369){target="_blank"}

[video (using app)](https://vimeo.com/531933725){target="_blank"}: the [app](http://18.191.167.248:3838/nprob/){target="_blank"}

First, draw a sketch. We can label the $`r rv`$ axis by adding integer multiples of `r sig` to `r mu`. We know to shade toward the left because small values of $`r rv`$ satisfy the condition $`r rv`<`r x`$.

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext(paste(rv,"axis",collapse=""),1,3,at=-3.5)
abline(h=seq(0,0.4,0.05))
for(zz in seq(-3,3,0.2)){
  lines(c(zz,zz),c(0,0.4))
}
abline(h=0,lwd=3)
if(!(zr %in% seq(-3,3,1))){
  axis(1,at=zr,labels="",line=3)
  mtext(x,1,4.9,at=zr)
}
polygon(c(Z[Z<=zr],zr,-3),c(dnorm(Z[Z<=zr]),0,0),col=rgb(1,0.5,0.5,0.5))
mtext("density",2,line=1)
```
\


We are given a specific $`r xl`$ value as a boundary. (Remember, for random variables we use uppercase letters, but for specific values we use lowercase.)
$$`r xl` = `r x` $$
We calculate the $z$ value of the boundary.
$$\begin{aligned}
z &= \frac{`r xl`-\mu}{\sigma} \\\\
&= \frac{`r x`-`r mu`}{`r sig`} \\\\
&= `r z`
\end{aligned}$$

We have rephrased our problem into a **standard normal** probability problem, because
$$P(`r rv`<`r x`) ~=~ P(Z<`r zr`) $$

So, we just need to evaluate $P(Z<`r zr`)$. To do this, you just need a $z$-table.
```{r}
zs = seq(z-0.02,z+0.02,0.01)
las = round(pnorm(zs),4)
```

| $z$ | $P(Z<z)$ |
|:---:|:---:|
| `r zs[1]` | `r las[1]` |
| `r zs[2]` | `r las[2]` |
| **`r zs[3]`** | **`r las[3]`** |
| `r zs[4]` | `r las[4]` |
| `r zs[5]` | `r las[5]` |


Thus, we find our answer.
$$P(`r rv`<`r x`) = `r la` $$


Meta-information
================
extype: num
exsolution: `r la`
exname: Get leftarea from normal
extol: 0.0002
