```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
while(T){
  la = runif(2,0.002,0.998)
  z = sort(round(qnorm(la),2))
  mag = sample(-1:2,1)
  mu = signif(sample(10:99,1)/10 * 10^mag,2)
  sig = round(runif(1,mu/20,mu/3),-mag+1)
  lets = strsplit("ABCDEFGHIJKMQRVWXY",split=NULL)
  rvi = sample(1:length(lets[[1]]),size=1)
  rv = lets[[1]][rvi]
  xl = tolower(rv)
  x = round(z*sig+mu,-mag+1)
  x1 = x[1]
  x2 = x[2]
  z = (x-mu)/sig
  z1 = z[1]
  z2 = z[2]
  la = round(pnorm(z),4)
  
  ba = la[2]-la[1]
  if(sig != 0 & z1 == round(z1,2) & z2 == round(z2,2) & ba > 0.01){break}
}
```

Question
========
Random variable $`r rv`$ is normally distributed with mean $\mu = `r mu`$ and standard deviation $\sigma = `r sig`$.
$$`r rv` = \text{N}(`r mu`,\,`r sig`) $$
Evaluate $P\left(`r x1` < X < `r x2`\right)$. In other words, what is the probability that $`r rv`$ is between $`r x1`$ and $`r x2`$?


Solution
========

[video (using z-table)](https://vimeo.com/531933369){target="_blank"}

[video (using app)](https://vimeo.com/531933725){target="_blank"}: the [app](http://18.191.167.248:3838/nprob/){target="_blank"}

First, draw a sketch. We can label the $`r rv`$ axis by adding integer multiples of `r sig` to `r mu`. We know to shade the between $`r x1`$ and $`r x2`$ because those values satisfy the condition $`r x1` < X < `r x2`$. 

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext(paste(rv,"axis",collapse=""),1,3,at=-3.5)
abline(h=seq(0,0.4,0.05))
for(zz in seq(-3,3,0.2)){
  lines(c(zz,zz),c(0,0.4))
}
abline(h=0,lwd=3)
axis(1,at=c(z1,z2),labels=c("",""),line=3)
mtext(x1,1,4.9,at=z1)
mtext(x2,1,4.9,at=z2)
polygon(c(Z[Z>=z1 & Z<=z2],z2,z1),c(dnorm(Z[Z>=z1 & Z<=z2]),0,0),col=rgb(1,0.5,0.5,0.5))
mtext("density",2,line=1)
```
\


We calculate the $z$ values of the boundaries. Left boundary:
$$\begin{aligned}
z_1 &= \frac{x_1-\mu}{\sigma} \\\\
&= \frac{`r x1`-`r mu`}{`r sig`} \\\\
&= `r z1`
\end{aligned}$$

Right boundary:

$$\begin{aligned}
z_2 &= \frac{x_2-\mu}{\sigma} \\\\
&= \frac{`r x2`-`r mu`}{`r sig`} \\\\
&= `r z2`
\end{aligned}$$

We rephrase our problem into a **standard normal** probability problem:

$$P\left(`r x1` < X < `r x2`\right) ~=~ P(`r z1`<Z<`r z2`)$$

So, we just need to evaluate $P(`r z1`<Z<`r z2`)$.

From here we have a formula that lets us use the $z$ table. (We practiced this part before.)
$$\begin{aligned}
P(z_1 < Z < z_2) &= P(Z<z_2) - P(Z<z1) \\
P(`r z1`<Z<`r z2`) &= P(Z<`r z2`)-P(Z<`r z1`)\\
&= `r la[2]`-`r la[1]` \\
&= `r ba`
\end{aligned}$$


Meta-information
================
extype: num
exsolution: `r ba`
exname: Get centralarea from normal
extol: 0.001


