```{r data generation, echo = FALSE, results = "hide"}
library(knitr)
opts_chunk$set(echo = FALSE)
options(scipen=100)

while(T){
  mu0 = sample(seq(200,800,100),1)
  n = sample(c(4,5,6,8,10,12,15,18,20,24,25,30,36,40),1)
  df = n-1
  sigma = runif(1,0.2,2)
  se = sigma/sqrt(n)
  gamma = 0.95
  tstar = qt(gamma+(1-gamma)/2,df)
  me = tstar*se
  mu = mu0+sample(c(-1,1),1)*me
  x = round(rnorm(n,mu,sigma),2)
  tt = t.test(x,mu=mu0)
  pval = tt$p.value
  CI = tt$conf.int
  t = tt$statistic
  if(pval>0.005 & pval<0.3 & abs(pval-0.05)>0.005){break}
}
write.csv(data.frame(x),"data_one_mean.csv",row.names=F)

if(pval<0.05){
  ans4="10"
  s4="The p-value is less than 0.05, so the difference is significant."
} else {
  ans4="01"
  s4="The p-value is more than 0.05, so the difference is NOT significant."
}

xrep = x
default_source_hook <- knit_hooks$get('source')
knit_hooks$set(source = function(y, options) {
  y <- gsub('xrep',paste0(x,collapse=", "),y)
  default_source_hook(y, options)
})
```

Question
========

An [automatic bottle filler](https://en.wikipedia.org/wiki/Filler_(packaging)){target="blank"} is supposed to average `r sprintf("%.2f",mu0)` ml of fluid in each bottle. You sampled some random bottles, recording their volumes:

```{r datadump,comment=NA}
xst = lapply(x,function(y){sprintf("%.2f",y)})
nfilledrows = length(xst)%/%10
nlast = length(xst)%%10
s = ""
if(nfilledrows>0){
  for(i in 1:nfilledrows){
    s = paste0(s,paste0(xst[((i-1)*10+1):(i*10)],collapse=", "))
    if(i==nfilledrows & nlast==0){
      s = paste0(s,"\n")
    } else {
      s = paste0(s,",\n")
    }
  }
}
if(nlast>0){
  s = paste0(s,
             paste0(xst[(nfilledrows*10+1):length(xst)],collapse=", "))
}
cat(s)
```

[Download data.](data_one_mean.csv){target="blank"}

You are asked to determine a 95% confidence interval, calculate an appropriate $p$-value (using two-tail $t$ test), and state whether the filler needs adjustment using a significance level of 0.05.

##ANSWER1##

##ANSWER2##

##ANSWER3##

Does the filler need adjustment?

##ANSWER4##

Answerlist
--------
* Determine the lower boundary of the confidence interval.
* Determine the upper boundary of the confidence interval.
* Determine the $p$-value.
* Yes
* No


Solution
========

$$n=`r n`$$
$$\text{df}=`r df` $$

$$t^\star = `r round(tstar,2)`$$
$$\bar{x} = `r round(mean(x),4)`$$
$$s = `r round(sd(x),4)`$$


$$\text{CI}_\text{bounds} = \bar{x}\pm \frac{t^\star s}{\sqrt{n}} $$

$$\text{CI}_\text{bounds} = \text{(`r round(CI[1],3)`, `r round(CI[2],3)`)}$$

$$t=\frac{\bar{x}-\mu_0}{s/\sqrt{n}}=\frac{`r round(mean(x),4)`-`r mu0`}{`r round(sd(x),4)`/ \sqrt{`r n`}}$$
$$t = `r round((mean(x)-mu0)/(sd(x)/sqrt(n)),3)`$$
$$p\text{-value}= P\left(|T|>`r round(abs(t),3)`\right) $$
$$p\text{-value} = `r round(pval,3)` $$

Answerlist
--------
* `r CI[1]`
* `r CI[2]`
* `r pval`
* `r s4`
* 


Meta-information
================
extype: cloze
exsolution: `r CI[1]`|`r CI[2]`|`r pval`|`r ans4`
exclozetype: num|num|num|schoice
exname: one mean
extol: 0.1|0.1|0.01|0
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE
