```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
while(T){
  la = runif(1,0.55,0.99)
  z = round(qnorm(la),2)
  mag = sample(-1:2,1)
  mu = signif(sample(10:99,1)/10 * 10^mag,2)
  sig = round(runif(1,mu/20,mu/3),-mag+1)
  lets = strsplit("ABCDEFGHIJKMQRVWXY",split=NULL)
  rvi = sample(1:length(lets[[1]]),size=1)
  rv = lets[[1]][rvi]
  xl = tolower(rv)
  x = round(z*sig+mu,-mag+1)
  d = x-mu
  x1 = mu-d
  x2 = mu+d
  z = (x-mu)/sig
  zr = round(z,2)
  la = round(pnorm(z),4)
  ra = 1-la
  ca = 1-2*ra
  if(sig != 0 & z == zr & z != 0){break}
}
```

Question
========
Random variable $`r rv`$ is normally distributed with mean $\mu = `r mu`$ and standard deviation $\sigma = `r sig`$.
$$`r rv` = \text{N}(`r mu`,\,`r sig`) $$
Evaluate $P\left(\big|`r rv`-`r mu`\big| < `r d`\right)$. In other words, what is the probability that $`r rv`$ is within $\pm`r x-mu`$ units from the mean?


Solution
========
First, draw a sketch. We can label the $`r rv`$ axis by adding integer multiples of `r sig` to `r mu`. We know to shade the center because values near `r mu` satisfy the condition $\big|`r rv`-`r mu`\big| < `r x-mu`$. We draw the boundaries at $`r mu`-`r d`=`r x1`$ and $`r mu`+`r d`=`r x2`$ because those are the solutions to $\big|`r rv`-`r mu`\big| = `r d`$. We can also rephrase the probability.

$$P\left(\big|`r rv`-`r mu`\big| < `r d`\right) ~=~ P(`r x1`<`r rv`<`r x2`)$$

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext(paste(rv,"axis",collapse=""),1,3,at=-3.5)
abline(h=seq(0,0.4,0.05))
for(zz in seq(-3,3,0.2)){
  lines(c(zz,zz),c(0,0.4))
}
abline(h=0,lwd=3)
if(!(zr %in% seq(-3,3,1))){
  axis(1,at=c(-zr,zr),labels=c("",""),line=3)
  mtext(x,1,4.9,at=zr)
  mtext(round(mu-z*sig,2),1,4.9,at=-zr)
}
polygon(c(Z[Z>=-zr & Z<=zr],zr,-zr),c(dnorm(Z[Z>=-zr & Z<=zr]),0,0),col=rgb(1,0.5,0.5,0.5))
mtext("density",2,line=1)
```
\


We calculate the $z$ values of the boundaries. Left boundary:
$$\begin{aligned}
z_1 &= \frac{`r x1`-`r mu`}{`r sig`} \\\\
&= \frac{`r -d`}{`r sig`} \\\\
&= `r -z`
\end{aligned}$$

Right boundary:

$$\begin{aligned}
z_2 &= \frac{`r x2`-`r mu`}{`r sig`} \\\\
&= \frac{`r d`}{`r sig`} \\\\
&= `r z`
\end{aligned}$$

We have rephrased our problem into a **standard normal** probability problem, because

$$P\left(\big|`r rv`-`r mu`\big| < `r d`\right) ~=~ P(`r x1`<`r rv`<`r x2`) ~=~ P(`r -z`<Z<`r z`) ~=~ P\left(\big|Z\big|<`r z`\right)$$

So, we just need to evaluate $P(|Z|<`r z`)$. I will also point out that $\frac{`r d`}{`r sig`} = `r z`$. In general, if $X$ is normally distributed, then:
$$P\left(\big|X-\mu\big|<d\right) ~=~ P\left(\big|Z\big| < \frac{d}{\sigma}\right)$$

From here we have a formula that lets us use the $z$ table. (We practiced this part before.)
$$\begin{aligned}
P(|Z|<`r z`) &= 2 \times P(Z<`r z`)-1 \\
&= 2 \times `r la`-1 \\
&= `r ca`
\end{aligned}$$


Meta-information
================
extype: num
exsolution: `r ca`
exname: Get centralarea from normal
extol: 0.0002


