```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)
p = round(runif(1,0.2,0.8),2)
mu = p
sigma = sqrt(p*(1-p))

n = c(25,100,400)
totlb = round(n*mu-2*sqrt(n)*sigma)
totub = round(n*mu+2*sqrt(n)*sigma)
avelb = mu-2*sigma/sqrt(n)
aveub = mu+2*sigma/sqrt(n)
```

Question
========

### Population

Random variable $X$ follows a [Bernoulli distribution](https://en.wikipedia.org/wiki/Bernoulli_distribution){target="blank"} with $p=`r p`$. In this context, each spin is often called a trial. A "0" is a "fail" and a "1" is a "success".

```{r, fig.width=4, fig.height=4}
par(mar=c(0,0,0,0))
theta=seq(0,2*pi,length.out=1000)
plot(cos(theta),sin(theta),type="l",ann=F,axes=F)
lines(c(0,sin(p*2*pi)),c(0,cos(p*2*pi)))
lines(c(0,0),c(0,1))
text(sin(p*pi)*0.8,cos(p*pi)*0.8,"1",cex=2)
text(sin(2*(p+(1-p)/2)*pi)*0.8,cos(2*(p+(1-p)/2)*pi)*0.8,"0",cex=2)
polygon(0.05*cos(theta),0.05*sin(theta),col="red")
arrows(-0.5,-0.1,0.5,0.1,lwd=6)
arrows(-0.5,-0.1,0.5,0.1,lwd=4,col="red")
polygon(0.01*cos(theta),0.01*sin(theta),col="red")
```
\

The average of a Bernoulli random variable is equal to $p$.
$$\mu=p$$
Determine $\mu$. ##ANSWER1##

The standard deviation of a Bernoulli random variable is equal to $\sqrt{p(1-p)}$.
$$\sigma = \sqrt{p(1-p)} $$
Determine $\sigma$. ##ANSWER2##
(Round to thousandths place.)


### Interval of typical success totals

The [binomial distribution](https://en.wikipedia.org/wiki/Binomial_distribution){target="blank"} predicts how many successes there will be for a given number of trials. Let the interval of typical successes be defined as having lower bound $np-2\sqrt{np(1-p)}$ and upper bound $np+2\sqrt{np(1-p)}$.
$$\text{interval of typical successes} = \left(np-2\sqrt{np(1-p)}, \, np+2\sqrt{np(1-p)}\right)$$

For the listed values of $n$, determine the bounds. (Round to nearest integer.)

| $n$ | _ lower bound of typical successes _ | _ upper bound of typical successes _ |
|:---:|:---:|:---:|
| 25  | ##ANSWER3## | ##ANSWER4## |
| 100  | ##ANSWER5## | ##ANSWER6## |
| 400  | ##ANSWER7## | ##ANSWER8## |



### Interval of typical proportions

Let the interval of typical proportions be defined as having lower bound $p-2\sqrt{\frac{p(1-p)}{n}}$ and upper bound $p+2\sqrt{\frac{p(1-p)}{n}}$.

$$\text{interval of typical proportions} = \left(p-2\sqrt{\frac{p(1-p)}{n}}, \, p+2\sqrt{\frac{p(1-p)}{n}}\right)$$
For the listed values of $n$, determine the bounds. (Round to nearest hundredth.)

| $n$ | _ lower bound of typical proportions _ | _ upper bound of typical proportions _ |
|:---:|:---:|:---:|
| 25  | ##ANSWER9## | ##ANSWER10## |
| 100  | ##ANSWER11## | ##ANSWER12## |
| 400  | ##ANSWER13## | ##ANSWER14## |



Solution
========

### Bernoulli

$$\mu = `r mu` $$
$$\sigma = `r sigma` $$

### Interval of typical successes

| $n$ | _ lower bound of typical successes _ | _ upper bound of typical successes _ |
|:---:|:---:|:---:|
| 25  | `r totlb[1]` | `r totub[1]` |
| 100  | `r totlb[2]` | `r totub[2]` |
| 400  | `r totlb[3]` | `r totub[3]` |

### Interval of typical proportions

| $n$ | _ lower bound of typical proportions _ | _ upper bound of typical proportions _ |
|:---:|:---:|:---:|
| 25  | `r round(avelb[1],2)` | `r round(aveub[1],2)` |
| 100  | `r round(avelb[2],2)` | `r round(aveub[2],2)` |
| 400  | `r round(avelb[3],2)` | `r round(aveub[3],2)` |


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num|num|num|num|num|num|num
exsolution: `r mu`|`r sigma`|`r totlb[1]`|`r totub[1]`|`r totlb[2]`|`r totub[2]`|`r totlb[3]`|`r totub[3]`|`r avelb[1]`|`r aveub[1]`|`r avelb[2]`|`r aveub[2]`|`r avelb[3]`|`r aveub[3]`
exname: intervals typical successes, proportions
extol: 0.01|0.001|1|1|1|1|1|1|0.01|0.01|0.01|0.01|0.01|0.01
exextra[numwidth,numeric]: 7
exextra[stringwidth,numeric]: 7
