```{r,data,generation,echo=FALSE,results="hide"}
types = c("Bell","Uniform","Bimodal","Right-skew","Left-skew")

n = 200
beta1 = c(3,1,0.5,0.6,1.5)
beta2 = c(3,1,0.5,1.5,0.6)

set.seed(200) #added line 

bell = round(rbeta(n,beta1[1],beta2[1])*10,2)
uniform = round(rbeta(n,beta1[2],beta2[2])*10,2)
bimodal = round(rbeta(n,beta1[3],beta2[3])*10,2)
right_skew = round(rbeta(n,beta1[4],beta2[4])*10,2)
left_skew = round(rbeta(n,beta1[5],beta2[5])*10,2)

data = data.frame(bell,uniform,bimodal,right_skew,left_skew)

ord = c(3,5,2,1,4) #sample(1:5)
type = 4 #sample(1:5,1)

i = 1:n
x = c(0.07, 4.25, 7.25, 2.22, 0.47, 0.00, 0.01, 0.55, 2.25, 3.41,
 0.67, 0.02, 1.40, 5.89, 0.17, 6.95, 0.05, 1.25, 3.79, 6.90,
 0.07, 3.89, 6.09, 2.17, 2.16, 3.23, 7.95, 7.56, 5.22, 0.04,
 5.96, 0.76, 2.02, 2.78, 0.83, 0.54, 0.14, 1.27, 3.69, 0.44,
 0.02, 0.03, 1.30, 4.74, 1.04, 0.82, 5.82, 2.48, 5.96, 3.42,
 4.69, 5.78, 2.39, 0.01, 4.53, 0.92, 3.56, 8.80, 5.34, 0.01,
 0.92, 1.36, 2.34, 8.02, 0.36, 0.02, 4.88, 0.00, 2.25, 3.71,
 2.66, 4.83, 1.38, 2.84, 3.59, 5.97, 1.65, 1.14, 4.70, 2.28,
 1.00, 1.12, 2.49, 0.07, 0.23, 0.52, 1.09, 1.84, 1.21, 6.63,
 0.81, 0.02, 0.28, 1.67, 5.47, 3.46, 8.89, 4.67, 3.92, 0.39,
 2.99, 3.21, 4.34, 2.05, 0.94, 0.83, 0.15, 3.42, 2.84, 0.92,
 0.31, 6.33, 0.00, 0.31, 0.94, 1.06, 0.08, 0.69, 6.52, 0.36,
 8.22, 0.17, 1.33, 3.05, 2.01, 9.81, 5.36, 2.64, 0.28, 0.12,
 3.80, 4.27, 5.00, 4.11, 1.38, 2.66, 0.36, 1.29, 0.41, 0.59,
 2.71, 5.33, 3.50, 4.50, 1.94, 1.11, 2.97, 0.45, 0.60, 8.96,
 0.47, 0.02, 9.35, 3.74, 0.65, 2.07, 3.40, 3.25, 8.67, 7.69,
 5.91, 0.07, 0.73, 5.24, 1.27, 7.05, 1.57, 0.62, 3.95, 1.84,
 0.76, 4.65, 3.23, 6.83, 0.47, 3.09, 0.32, 1.47, 0.41, 3.87,
 0.35, 1.81, 1.28, 0.70, 4.05, 0.80, 1.58, 0.06, 1.54, 3.44,
 4.21, 1.40, 1.32, 2.07, 0.34, 1.78, 0.80, 1.64, 4.56, 0.58) #data[,type]

write.csv(data.frame(i,x),"make_hist.csv",row.names=F)
ans1 = paste0((type==ord)*1,collapse="")
ans2 = paste0((1:5==type)*1,collapse="")
```


Question
========

A sample of size $n=`r n`$ was taken from an unknown population. 

```{r echo=F, comment=NA}
y = lapply(x, function(xx) sprintf("%5.2f",xx))
data2 = matrix(y,ncol=10)
s = paste( apply(data2,1,function(xx) paste(xx,collapse=",")), collapse=",\n")
cat(s)
```

You can [download the data as a CSV](make_hist.csv). Determine which histogram visualizes the data, and describe the shape of the data.

```{r echo=F,fig.width=8,fig.height=5}
par(mfrow=c(2,3))
for(j in 1:5){
  par(mar=c(2,1,3,1))
  i = ord[j]
  hist(data[,i],axes=F,xlim=c(0,10),main=paste0("histogram ",j),col=rgb(0.9,0.7,0.7))
  box("figure")
}
```
\

* Which histogram is appropriate? \#\#ANSWER1\#\#
* Which shape description is appropriate?  \#\#ANSWER2\#\#


Answerlist
--------
* histogram 1
* histogram 2
* histogram 3
* histogram 4
* histogram 5
* Bell
* Uniform
* Bimodal
* Right-skew
* Left-skew


Solution
========

[video](https://vimeo.com/531905389){target="_blank"}

You should make a histogram. This is easy in R.

```{r fig.height=4,fig.width=5}
x = read.csv("make_hist.csv")$x
hist(x)
```
\

Using a spreadsheet is [way more work](https://wikieducator.org/OpenOffice/Calc_3/Histogram){target="blank"}. But, you could use the [`FREQUENCY` function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_FREQUENCY_function){target="_blank"}.

The correct histogram is number `r (1:5)[type==ord]`.
The correct description is `r types[type]`.

Answerlist
--------
* `r (type==ord)[1]`
* `r (type==ord)[2]`
* `r (type==ord)[3]`
* `r (type==ord)[4]`
* `r (type==ord)[5]`
* `r (1:5==type)[1]`
* `r (1:5==type)[2]`
* `r (1:5==type)[3]`
* `r (1:5==type)[4]`
* `r (1:5==type)[5]`


Meta-information
================
extype: cloze
exclozetype: schoice|schoice
exsolution: `r ans1`|`r ans2`
exname: identify_histogram
extol: 0
exextra[numwidth]: 999.999
