```{r data generation, echo = FALSE, results = "hide"}
library(latex2exp)
while(T)
{
  ell = round(runif(1,0.60,0.98),2)
  ra = round(1-ell,2)
  ca = round(1-2*ra,2)
  ta = round(ra*2,2)
  z = qnorm(ell)
  zlow = floor(z*100)/100
  zhigh = ceiling(z*100)/100
  if(round(z,4) != round(z,2)) {break}
}
```

Question
========
Determine $z$ such that $P(|Z|>z)=`r ta`$. In other words, how far from 0 should boundaries be set such that `r ta*100`% of standard normal values are outside those boundaries? (Answers within 0.01 from the correct value will be marked correct.)


Solution
========
Start with a sketch. The total two-tail area is `r ta`, so each tail has half that area.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),
        c(dnorm(zs[zs<=-z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text(min(z+0.4,2.6),max(dnorm(z+0.4)/3,0.04),TeX(sprintf("$\\frac{%.2f}{2}$",ta)))
text(max(-z-0.4,-2.6),max(dnorm(z+0.4)/3,0.04),TeX(sprintf("$\\frac{%.2f}{2}$",ta)))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","z=?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
```
\

**Method 1:** Determine the area of each tail and the center. Both tails have the same area, and all three areas add to 1. Thus,
$$P(Z>z) = \frac{`r ta`}{2} = `r ra`$$

$$P(Z<-z) = \frac{`r ta`}{2} = `r ra`$$

$$\begin{aligned}
P(|Z|<z) &= 1-P(|Z|>z) \\ 
&= 1-`r ta` \\
&= `r ca`
\end{aligned}$$
```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),
        c(dnorm(zs[zs<=-z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text(0,dnorm(0)/3,ca)
xl = (-3-z)/2
xr = (3+z)/2
text(xl,max(dnorm(xl)/3,0.02),ra)
text(xr,max(dnorm(xr)/3,0.02),ra)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","z=?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.42,0.42,0.4))
text((-3+z)/2,0.45,ell)
```
\

Find the left area.

$$\begin{aligned}
P(Z<z) &= `r ra`+`r ca` \\
&= `r ell`
\end{aligned}$$


```{r, echo=F}
z2 = seq(zlow-0.02,zhigh+0.02,0.01)
```

You should go to your $z$-table and find the $z$-score with the left area closest to `r ell`.

| $z$ | $P(Z<z)$ |
|:---:|:---:|
| `r z2[1]` | `r format(round(pnorm(z2[1]),4),4)` |
| `r z2[2]` | `r format(round(pnorm(z2[2]),4),4)` |
| **`r z2[3]`** | **`r format(round(pnorm(z2[3]),4),4)`** |
| **`r z2[4]`** | **`r format(round(pnorm(z2[4]),4),4)`** |
| `r z2[5]` | `r format(round(pnorm(z2[5]),4),4)` |
| `r z2[6]` | `r format(round(pnorm(z2[6]),4),4)` |

It turns out the exact answer is $z=`r z`$, which could be found by using an inverse normal function. On a spreadsheet:
```
=Norm.Inv(`r ell`,0,1)
```
Using R:
```
twotailarea = `r ta`
onetailarea = twotailarea/2
centralarea = 1-twotailarea
leftarea = onetailarea + centralarea
qnorm( leftarea )
```

**Method 2:** Another way to get `r ell` is by subtracting half of `r ta` from 1.
$$P(Z<z) ~=~ 1-\frac{`r ta`}{2}~=~ 1-`r ra` ~=~ `r ell`$$

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),ylim=c(0,0.5),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=-z],-z,-3),
        c(dnorm(zs[zs<=-z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
polygon(c(zs[zs>=z],3,z),
        c(dnorm(zs[zs>=z]),0,0),
        col=rgb(0.7,0.7,0.7,0.7))
text(min(z+0.4,2.6),max(dnorm(z+0.4)/3,0.04),ra)
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0,labels=c("-3","z=?","3"))
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
lines(c(-3,z,z),c(0.46,0.46,0.45))
text((-3+z)/2,0.476,ell)
```
\

Then, use the table. Or, R:
```
twotailarea = `r ta`
leftarea = 1 - twotailarea/2
qnorm( leftarea )
```

You might find a spinner visualization useful.

```{r echo=F,fig.width=6,fig.height=6}
par(mar=c(0,0,0,0))
theta=seq(0,2*pi,length.out=1000)
plot(sin(theta),cos(theta),type="l",ann=F,axes=F,xlim=c(-1.3,1.3),ylim=c(-1.3,1.3))
zs = -3:3
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.8,1),cos(p*2*pi)*c(0.8,1),lwd=2)
  pt = min(max(p,0.007),0.993)
  if(abs(zs[i])<3){
    text(0.75*sin(pt*2*pi),0.75*cos(pt*2*pi),zs[i])
  }
}
zs = seq(-2.5,2.5,1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.85,1),cos(p*2*pi)*c(0.85,1),lwd=2)
  if(abs(zs[i])<3){
    text(sin(p*2*pi)*0.8,cos(p*2*pi)*0.8,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.9,1),cos(p*2*pi)*c(0.9,1),lwd=1)
}
zs = seq(-2,2,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  if(zs[i]%%0.5 != 0){
    text(sin(p*2*pi)*0.85,cos(p*2*pi)*0.85,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.05)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.95,1),cos(p*2*pi)*c(0.95,1))
}
zs = seq(-3,3,0.01)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.98,1),cos(p*2*pi)*c(0.98,1))
}

for(p in seq(0.1,0.9,0.1)){
  lines(sin(p*2*pi)*c(1,1.05),cos(p*2*pi)*c(1,1.05))
  text(sin(p*2*pi)*1.1,cos(p*2*pi)*1.1,p,cex=0.8,srt=-(p*360-90)%%180-90)
}
for(p in seq(0.01,0.99,0.01)){
  lines(sin(p*2*pi)*c(1,1.03),cos(p*2*pi)*c(1,1.03))
}
for(p in seq(0.002,0.998,0.002)){
  lines(sin(p*2*pi)*c(1,1.01),cos(p*2*pi)*c(1,1.01))
}
lines(c(0,0),c(1,1.05),lwd=2)
text(0,1.09,"1 or 0",cex=0.8)
text(0,0,cex=1.5,"z")


ps = c(seq(pnorm(z),1,0.001),seq(0,pnorm(-z),0.001))
polygon(c(0,sin(ps*2*pi),0),c(0,cos(ps*2*pi),0),col=rgb(0.9,0.3,0.3,0.4),border=rgb(0.9,0.3,0.3))
pe = pnorm(c(-z,z))
pm = 0

lines(sin(ps*2*pi)*1.2,cos(ps*2*pi)*1.2,lwd=2)
lines(sin(pe[1]*2*pi)*c(1.15,1.2),cos(pe[1]*2*pi)*c(1.15,1.2),lwd=2)
lines(sin(pe[2]*2*pi)*c(1.15,1.2),cos(pe[2]*2*pi)*c(1.15,1.2),lwd=2)
text(0.4*sin(2*pi*pe[1]),0.4*cos(2*pi*pe[1]),round(-z,7),srt=(180-360*pe[1])%%180-90)
text(0.4*sin(2*pi*pe[2]),0.4*cos(2*pi*pe[2]),round(z,7),srt=(180-360*pe[2])%%180-90)
text(1.25*sin(2*pi*pe[1]),1.25*cos(2*pi*pe[1]),round(pe[1],4),srt=-(pe[1]*360-90)%%180-90,cex=0.8)
text(1.25*sin(2*pi*pe[2]),1.25*cos(2*pi*pe[2]),round(pe[2],4),srt=-(pe[2]*360-90)%%180-90,cex=0.8)
text(sin(pm*2*pi)*1.25,
     cos(pm*2*pi)*1.25,
     sprintf("%.4f",1-(round(pe[2],4)-round(pe[1],4))),
     srt=-(pm*360-90)%%180-90)
```
\


Meta-information
================
extype: num
exsolution: `r z`
exname: Get z from twotail area
extol: 0.011
