---
title: "Normal Distribution"
author: "Chad Worley"
date: "4/8/2020"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## General Normal Distribution

Let $X$ be normally distributed with mean $\mu$ and standard deviation $\sigma$.
$$X \sim \mathcal{N}(\mu,\,\sigma) $$
```{r, fig.width=8, fig.height=8}
par(mar=c(0,0,0,0))
plot(0,0,type="n",xlim=c(-1.2,1.2),ylim=c(-1.2,1.2),ann=F)
theta = seq(0,2*pi,0.01)
polygon(cos(theta),sin(theta))
z1 = round(seq(-3,3,1),3)
z2 = round(seq(-3,3,0.5),2)
z3 = round(seq(-3,3,0.1),2)
z4 = round(seq(-2,2,0.05),2)
z5 = round(seq(-2,2,0.01),2)
markz = function(z,L,lwd=1){
    p = pnorm(z)
    for(pp in p){
        lines(c((1-L)*cos(2*pi*pp),cos(2*pi*pp)),c((1-L)*sin(2*pi*pp),sin(2*pi*pp)),lwd=lwd)
    }
}
markz(z1,0.15,3)
markz(z2,0.1)
markz(z3,0.05)
# markz(z4,0.05)
# markz(z5,0.025)
z = -3:3
p = pnorm(z)
pc = sapply(p,function(x){min(max(x,0.007),0.993)})
r1 = 0.75
text(r1*cos(2*pi*pc[1]),r1*sin(2*pi*pc[1]),expression(paste(mu,"-3",sigma)),cex=1.5)
text(r1*cos(2*pi*pc[2]),r1*sin(2*pi*pc[2]),expression(paste(mu,"-2",sigma)),cex=1.5)
text(r1*cos(2*pi*pc[3]),r1*sin(2*pi*pc[3]),expression(paste(mu,"-",sigma)),cex=1.5)
text(r1*cos(2*pi*pc[4]),r1*sin(2*pi*pc[4]),expression(paste(mu)),cex=1.5)
text(r1*cos(2*pi*pc[5]),r1*sin(2*pi*pc[5]),expression(paste(mu,"+",sigma)),cex=1.5)
text(r1*cos(2*pi*pc[6]),r1*sin(2*pi*pc[6]),expression(paste(mu,"+2",sigma)),cex=1.5)
text(r1*cos(2*pi*pc[7]),r1*sin(2*pi*pc[7]),expression(paste(mu,"+3",sigma)),cex=1.5)

z = c(-1.5,-0.5,0.5,1.5)
pc = pnorm(z)
r1 = 0.8
text(r1*cos(2*pi*pc[1]),r1*sin(2*pi*pc[1]),expression(paste(mu,"-1.5",sigma)))
text(r1*cos(2*pi*pc[2]),r1*sin(2*pi*pc[2]),expression(paste(mu,"-0.5",sigma)))
text(r1*cos(2*pi*pc[3]),r1*sin(2*pi*pc[3]),expression(paste(mu,"+0.5",sigma)))
text(r1*cos(2*pi*pc[4]),r1*sin(2*pi*pc[4]),expression(paste(mu,"+1.5",sigma)))

# for(z in z2[!(z2 %in% z1)]){
#     p = pnorm(z)
#     pc = min(max(p,0.005),0.995)
#     text((1-0.26)*cos(2*pi*pc),(1-0.26)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),cex=0.9)
# }
# for(z in z3[!(z3 %in% z2) & abs(z3)<2]){
#     p = pnorm(z)
#     pc = min(max(p,0.005),0.995)
#     text((1-0.16)*cos(2*pi*pc),(1-0.16)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.9)
# }
# for(z in z4[!(z4 %in% z3) & abs(z4)<1.5]){
#     p = pnorm(z)
#     pc = min(max(p,0.005),0.995)
#     text((1-0.1)*cos(2*pi*pc),(1-0.1)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.6)
# }
p1 = round(seq(0,1,0.1),2)
p2 = round(seq(0,1,0.05),2)
p3 = round(seq(0,1,0.01),2)
lines(c(1,1.2),c(0,0),lwd=3)
for(p in p1){
    lines(c(cos(p*2*pi),1.15*cos(p*2*pi)),c(sin(p*2*pi),1.15*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.2*cos(pc*2*pi),1.2*sin(pc*2*pi),round(p,2),cex=1)
}
for(p in p2[!(p2 %in% p1)]){
    lines(c(cos(p*2*pi),1.1*cos(p*2*pi)),c(sin(p*2*pi),1.1*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.16*cos(pc*2*pi),1.16*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}
for(p in p3[!(p3 %in% p2)]){
    lines(c(cos(p*2*pi),1.05*cos(p*2*pi)),c(sin(p*2*pi),1.05*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.12*cos(pc*2*pi),1.12*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}

text(1,1,"U\nStandard Uniform\nDistribution")
text(0,0,"X\nA Normal Distribution\nwith")
text(0,-0.15,expression(paste("mean = ",mu)))
text(0,-0.2,expression(paste("stdev = ",sigma)))
```
\

For any specific value $x$, like a boundary or single result, we can find a corresponding $z$-score.
$$z = \frac{x-\mu}{\sigma} $$

The probability that $X$ lands below $x$ can be found by using the corresponding $z$-score.
$$P(X<x) ~=~  P\left(Z<\frac{x-\mu}{\sigma}\right)$$
To determine the probability that $X$ lands over $x$:
$$P(X>x) ~=~  P\left(Z>\frac{x-\mu}{\sigma}\right)$$
When discussing central and two-tail probabilities of general normal distributions, it is helpful to introduce a new variable. I will use $d$ as the signed distance from mean to boundary. This is formally called the **deviation** from mean. https://en.wikipedia.org/wiki/Deviation_(statistics)
$$d = x-\mu $$

$$z = \frac{d}{\sigma} $$

To determine the probability that $X$ lands within $\mu-d$ and $\mu+d$.
$$P\left(\big|X-\mu\big|<d\right)  ~=~  P\left(\big|Z\big|<\frac{d}{\sigma}\right)  $$
To determine the probability that $X$ lands outside $\mu-d$ and $\mu+d$.
$$P\left(\big|X-\mu\big|>d\right)  ~=~ P\left(\big|Z\big|>\frac{d}{\sigma}\right)$$
To determine probability that $X$ lands inside two arbitrary asymmetric boundaries $x_1$ and $x_2$.
$$P(x_1 < X < x_2) ~=~ P\left(\frac{x_1-\mu}{\sigma} < Z < \frac{x_2-\mu}{\sigma} \right) $$

Once you have translated the probability statement into a *standard* normal probability statement, you can use the techniques you have already practiced.

When determining boundaries from the probability, a little algebra is helpful.
$$x = \mu+z\sigma $$
$$d = z\sigma $$


## Example Normal Distribution

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

The spinner:

```{r, fig.width=8, fig.height=8}
mu = 100
sig = 5
par(mar=c(0,0,0,0))
plot(0,0,type="n",xlim=c(-1.2,1.2),ylim=c(-1.2,1.2),ann=F)
theta = seq(0,2*pi,0.01)
polygon(cos(theta),sin(theta))
z1 = round(seq(-3,3,1),3)
z2 = round(seq(-3,3,0.5),2)
z3 = round(seq(-3,3,0.1),2)
z4 = round(seq(-2,2,0.05),2)
z5 = round(seq(-2,2,0.01),2)
markz = function(z,L,lwd=1){
    p = pnorm(z)
    for(pp in p){
        lines(c((1-L)*cos(2*pi*pp),cos(2*pi*pp)),c((1-L)*sin(2*pi*pp),sin(2*pi*pp)),lwd=lwd)
    }
}
markz(z1,0.3,3)
markz(z2,0.18)
markz(z3,0.1)
markz(z4,0.05)
markz(z5,0.025)
for(z in z1){
    p = pnorm(z)
    pc = min(max(p,0.005),0.995)
    text((1-0.37)*cos(2*pi*pc),(1-0.37)*sin(2*pi*pc),round(signif(mu+sig*qnorm(p),4),6),cex=0.9)
}
for(z in z2[!(z2 %in% z1)]){
    p = pnorm(z)
    pc = min(max(p,0.005),0.995)
    text((1-0.26)*cos(2*pi*pc),(1-0.26)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),cex=0.9)
}
for(z in z3[!(z3 %in% z2) & abs(z3)<2]){
    p = pnorm(z)
    pc = min(max(p,0.005),0.995)
    text((1-0.16)*cos(2*pi*pc),(1-0.16)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.9)
}
for(z in z4[!(z4 %in% z3) & abs(z4)<1.5]){
    p = pnorm(z)
    pc = min(max(p,0.005),0.995)
    text((1-0.1)*cos(2*pi*pc),(1-0.1)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.6)
}
p1 = round(seq(0,1,0.1),2)
p2 = round(seq(0,1,0.05),2)
p3 = round(seq(0,1,0.01),2)
lines(c(1,1.2),c(0,0),lwd=3)
for(p in p1){
    lines(c(cos(p*2*pi),1.15*cos(p*2*pi)),c(sin(p*2*pi),1.15*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.2*cos(pc*2*pi),1.2*sin(pc*2*pi),round(p,2),cex=1)
}
for(p in p2[!(p2 %in% p1)]){
    lines(c(cos(p*2*pi),1.1*cos(p*2*pi)),c(sin(p*2*pi),1.1*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.16*cos(pc*2*pi),1.16*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}
for(p in p3[!(p3 %in% p2)]){
    lines(c(cos(p*2*pi),1.05*cos(p*2*pi)),c(sin(p*2*pi),1.05*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.12*cos(pc*2*pi),1.12*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}

text(1,1,"U\nStandard Uniform\nDistribution")
text(0,0,sprintf("X\nA Normal Distribution\nwith\nmean = %.4g \n and \nstdev = %.4g",mu,sig))
```
\

The density curve:

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext("X axis",1,3,at=-3.5)
abline(h=0,lwd=3)
abline(h=seq(0,0.4,0.05),lwd=0.5)
abline(v=seq(-3,3,0.2),lwd=0.5)
```
\

### Probability that $X$ lands UNDER $x$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
x = 95
z = (x-mu)/sig
```
Let's pick an arbitrary boundary $x=`r x`$. How would we calculate the probability that $X$ lands on a value less than `r x`?
$$P(X<`r x`) = \,?$$

Well, we convert the boundary into a $z$ score.
$$\begin{aligned}
z &= \frac{x-\mu}{\sigma} \\ \\
&= \frac{`r x`-`r mu`}{`r sig`} \\ \\
&= `r z`
\end{aligned}$$

So, we have rephrased the probability expression.
$$P(X<`r x`) ~=~ P(Z<`r z`)$$

$$P(Z<`r z`) ~=~ \,?$$

Luckily for us, we know how to determine standard normal probabilities. This one just needs to be looked up on a $z$ table.
$$P(Z<`r z`) ~=~ `r round(pnorm(z),4)`$$
Thus,
$$P(X<`r x`) ~=~ `r round(pnorm(z),4)`$$

Notice that on the spinner, if we shade the inner markings up to `r x`, the corresponding outer marking is approximately `r round(pnorm(z),3)`.

```{r, fig.width=8, fig.height=8}
mu = 100
sig = 5
par(mar=c(0,0,0,0))
plot(0,0,type="n",xlim=c(-1.2,1.2),ylim=c(-1.2,1.2),ann=F)
theta = seq(0,2*pi,0.01)
polygon(cos(theta),sin(theta))
z1 = round(seq(-3,3,1),3)
z2 = round(seq(-3,3,0.5),2)
z3 = round(seq(-3,3,0.1),2)
z4 = round(seq(-2,2,0.05),2)
z5 = round(seq(-2,2,0.01),2)
markz = function(z,L,lwd=1){
    p = pnorm(z)
    for(pp in p){
        lines(c((1-L)*cos(2*pi*pp),cos(2*pi*pp)),c((1-L)*sin(2*pi*pp),sin(2*pi*pp)),lwd=lwd)
    }
}
markz(z1,0.3,3)
markz(z2,0.18)
markz(z3,0.1)
markz(z4,0.05)
markz(z5,0.025)
for(z0 in z1){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.37)*cos(2*pi*pc),(1-0.37)*sin(2*pi*pc),round(signif(mu+sig*qnorm(p),4),6),cex=0.9)
}
for(z0 in z2[!(z2 %in% z1)]){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.26)*cos(2*pi*pc),(1-0.26)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),cex=0.9)
}
for(z0 in z3[!(z3 %in% z2) & abs(z3)<2]){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.16)*cos(2*pi*pc),(1-0.16)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.9)
}
for(z0 in z4[!(z4 %in% z3) & abs(z4)<1.5]){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.1)*cos(2*pi*pc),(1-0.1)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.6)
}
p1 = round(seq(0,1,0.1),2)
p2 = round(seq(0,1,0.05),2)
p3 = round(seq(0,1,0.01),2)
lines(c(1,1.2),c(0,0),lwd=3)
for(p in p1){
    lines(c(cos(p*2*pi),1.15*cos(p*2*pi)),c(sin(p*2*pi),1.15*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.2*cos(pc*2*pi),1.2*sin(pc*2*pi),round(p,2),cex=1)
}
for(p in p2[!(p2 %in% p1)]){
    lines(c(cos(p*2*pi),1.1*cos(p*2*pi)),c(sin(p*2*pi),1.1*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.16*cos(pc*2*pi),1.16*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}
for(p in p3[!(p3 %in% p2)]){
    lines(c(cos(p*2*pi),1.05*cos(p*2*pi)),c(sin(p*2*pi),1.05*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.12*cos(pc*2*pi),1.12*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}

P = seq(0,1,0.001)
col = rgb(1,0.5,0.5,0.5)
pr = pnorm((x-mu)/sig)
polygon(c(0,cos(2*pi*P[P<=pr])),c(0,sin(2*pi*P[P<=pr])),col=col)

text(1,1,"U\nStandard Uniform\nDistribution")
text(0,0,sprintf("X\nA Normal Distribution\nwith\nmean = %.4g \n and \nstdev = %.4g",mu,sig))
```
\


Notice if we shade under the density curve up to $x=`r x`$, we shade approximately `r round(pnorm(z)*100)` percentage squares.

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext("X axis",1,3,at=-3.5)
abline(h=0,lwd=3)
abline(h=seq(0,0.4,0.05),lwd=0.5)
abline(v=seq(-3,3,0.2),lwd=0.5)

polygon(c(Z[Z<=z],z,-3),c(dnorm(Z[Z<=z]),0,0),col=rgb(1,0.5,0.5,0.5))

text(-2.01,0.32,TeX(sprintf("$P(%s<%s) = %.4f$","X",x,round(pnorm(z),4))))
arrows(-1.6,0.3,(2*z-3)/3,max(dnorm((2*z-3)/3)/2,0.03),lwd=2)
```
\



### Probability that $X$ lands OVER $x$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
x = 102
z = (x-mu)/sig
```
Let's pick an arbitrary boundary $x=`r x`$. How would we calculate the probability that $X$ lands on a value more than `r x`?
$$P(X>`r x`) = \,?$$

Well, we convert the boundary into a $z$ score.
$$\begin{aligned}
z &= \frac{x-\mu}{\sigma} \\ \\
&= \frac{`r x`-`r mu`}{`r sig`} \\ \\
&= `r z`
\end{aligned}$$

So, we have rephrased the probability expression.
$$P(X>`r x`) ~=~ P(Z>`r z`)$$

$$P(Z>`r z`) ~=~ \,?$$

Luckily for us, we know how to determine standard normal probabilities. 
$$P(Z>`r z`) ~=~ `r round(1-pnorm(z),4)`$$
Thus,
$$P(X>`r x`) ~=~ `r round(1-pnorm(z),4)`$$

Notice that on the spinner, if we shade the inner markings over `r x`, the corresponding outer markings are approximately `r round(pnorm(z),3)` and 1. If we take a difference, we can estimate the probability. $1-`r round(pnorm(z),3)` = `r round(1-pnorm(z),3)`$. So, we see that about `r round(1-pnorm(z),3)*100`% of the circle is shaded.

```{r, fig.width=8, fig.height=8}
mu = 100
sig = 5
par(mar=c(0,0,0,0))
plot(0,0,type="n",xlim=c(-1.2,1.2),ylim=c(-1.2,1.2),ann=F)
theta = seq(0,2*pi,0.01)
polygon(cos(theta),sin(theta))
z1 = round(seq(-3,3,1),3)
z2 = round(seq(-3,3,0.5),2)
z3 = round(seq(-3,3,0.1),2)
z4 = round(seq(-2,2,0.05),2)
z5 = round(seq(-2,2,0.01),2)
markz = function(z,L,lwd=1){
    p = pnorm(z)
    for(pp in p){
        lines(c((1-L)*cos(2*pi*pp),cos(2*pi*pp)),c((1-L)*sin(2*pi*pp),sin(2*pi*pp)),lwd=lwd)
    }
}
markz(z1,0.3,3)
markz(z2,0.18)
markz(z3,0.1)
markz(z4,0.05)
markz(z5,0.025)
for(z0 in z1){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.37)*cos(2*pi*pc),(1-0.37)*sin(2*pi*pc),round(signif(mu+sig*qnorm(p),4),6),cex=0.9)
}
for(z0 in z2[!(z2 %in% z1)]){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.26)*cos(2*pi*pc),(1-0.26)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),cex=0.9)
}
for(z0 in z3[!(z3 %in% z2) & abs(z3)<2]){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.16)*cos(2*pi*pc),(1-0.16)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.9)
}
for(z0 in z4[!(z4 %in% z3) & abs(z4)<1.5]){
    p = pnorm(z0)
    pc = min(max(p,0.005),0.995)
    text((1-0.1)*cos(2*pi*pc),(1-0.1)*sin(2*pi*pc),signif(mu+sig*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.6)
}
p1 = round(seq(0,1,0.1),2)
p2 = round(seq(0,1,0.05),2)
p3 = round(seq(0,1,0.01),2)
lines(c(1,1.2),c(0,0),lwd=3)
for(p in p1){
    lines(c(cos(p*2*pi),1.15*cos(p*2*pi)),c(sin(p*2*pi),1.15*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.2*cos(pc*2*pi),1.2*sin(pc*2*pi),round(p,2),cex=1)
}
for(p in p2[!(p2 %in% p1)]){
    lines(c(cos(p*2*pi),1.1*cos(p*2*pi)),c(sin(p*2*pi),1.1*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.16*cos(pc*2*pi),1.16*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}
for(p in p3[!(p3 %in% p2)]){
    lines(c(cos(p*2*pi),1.05*cos(p*2*pi)),c(sin(p*2*pi),1.05*sin(p*2*pi)))
    pc = max(min(p,0.996),0.004)
    text(1.12*cos(pc*2*pi),1.12*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}

P = seq(0,1,0.001)
col = rgb(1,0.5,0.5,0.5)
pr = pnorm((x-mu)/sig)
polygon(c(0,cos(2*pi*P[P>=pr])),c(0,sin(2*pi*P[P>=pr])),col=col)

text(1,1,"U\nStandard Uniform\nDistribution")
text(0,0,sprintf("X\nA Normal Distribution\nwith\nmean = %.4g \n and \nstdev = %.4g",mu,sig))
```
\


Notice if we shade the density curve where $x>`r x`$, we shade approximately `r round(100-pnorm(z)*100)` percentage squares.

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext("X axis",1,3,at=-3.5)
abline(h=0,lwd=3)
abline(h=seq(0,0.4,0.05),lwd=0.5)
abline(v=seq(-3,3,0.2),lwd=0.5)

polygon(c(Z[Z>=z],3,z),c(dnorm(Z[Z>=z]),0,0),col=rgb(1,0.5,0.5,0.5))

text(2.01,0.32,TeX(sprintf("$P(%s>%s) = %.4f$","X",x,1-round(pnorm(z),4))))
arrows(1.6,0.3,(2*z+3)/3,max(dnorm((2*z+3)/3)/2,0.03),lwd=2)
```
\



### Probability that $X$ lands near $\mu$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$
```{r}
d = 7
x1 = mu-d
x2 = mu+d
z = d/sig
la = round(pnorm(z),4)
ra = 1-la
ta = 2*ra
ca = round(1-ta,4)
```

Let's pick an arbitrary deviation $d=`r d`$. Let's pretend that we are interesting in the probability that $X$ is within $\pm`r d`$ from $`r mu`$. In other words, what is the probability that $X$ is between $`r x1`$ and $`r x2`$?

$$P\left(\big|X-`r mu`\big|<`r d`\right) ~=~ \,? $$

We rephrase this as a standard normal probability.
$$\begin{aligned}
z &= \frac{d}{\sigma} \\\\
&= \frac{`r d`}{`r sig`} \\\\
&= `r z`
\end{aligned}$$


$$P\left(\big|X-`r mu`\big|<`r d`\right) ~=~ P\left(\big|Z\big|<`r z`\right) $$

$$ P\left(\big|Z\big|<`r z`\right) ~=~ ?$$


Luckily, we know how to find standard normal probabilities.
$$ P\left(\big|Z\big|<`r z`\right) ~=~ `r ca`$$

$$ P\left(\big|X-`r mu`\big|<`r d`\right) ~=~ `r ca`$$


### Probability that $X$ lands far from $\mu$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$
```{r}
d = 4
x1 = mu-d
x2 = mu+d
z = d/sig
la = round(pnorm(z),4)
ra = 1-la
ta = round(2*ra,4)
ca = round(1-ta,4)
```

Let's pick an arbitrary deviation $d=`r d`$. Let's pretend that we are interesting in the probability that $X$ is outside $\pm`r d`$ from $`r mu`$. In other words, what is the probability that $X$ is outside the boundaries $`r x1`$ and $`r x2`$?

$$P\left(\big|X-`r mu`\big|>`r d`\right) ~=~ \,? $$

We rephrase this as a standard normal probability.
$$\begin{aligned}
z &= \frac{d}{\sigma} \\\\
&= \frac{`r d`}{`r sig`} \\\\
&= `r z`
\end{aligned}$$


$$P\left(\big|X-`r mu`\big|>`r d`\right) ~=~ P\left(\big|Z\big|>`r z`\right) $$

$$ P\left(\big|Z\big|>`r z`\right) ~=~ ?$$


Luckily, we know how to find standard normal probabilities.
$$ P\left(\big|Z\big|>`r z`\right) ~=~ `r ta`$$

$$ P\left(\big|X-`r mu`\big|>`r d`\right) ~=~ `r ta`$$



### Probability that $X$ lands between $x_1$ and $x_2$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
rv = 'X'
x1 = 99
x2 = 108
z1 = (x1-mu)/sig
z2 = (x2-mu)/sig
ba = round(pnorm(z2),4)-round(pnorm(z1),4)
```

Let's pretend we want to know the probability that $X$ lands between $`r x1`$ and $`r x2`$.

First, draw a sketch. We can label the $`r rv`$ axis by adding integer multiples of `r sig` to `r mu`. We know to shade the between $`r x1`$ and $`r x2`$ because those values satisfy the condition $`r x1` < X < `r x2`$. 

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mu-3*sig,mu+3*sig,length.out=7),3),line=3)
mtext(paste(rv,"axis",collapse=""),1,3,at=-3.5)
abline(h=seq(0,0.4,0.05))
for(zz in seq(-3,3,0.2)){
  lines(c(zz,zz),c(0,0.4))
}
abline(h=0,lwd=3)
axis(1,at=c(z1,z2),labels=c("",""),line=3)
mtext(x1,1,4.9,at=z1)
mtext(x2,1,4.9,at=z2)
polygon(c(Z[Z>=z1 & Z<=z2],z2,z1),c(dnorm(Z[Z>=z1 & Z<=z2]),0,0),col=rgb(1,0.5,0.5,0.5))
mtext("density",2,line=1)
```
\


We calculate the $z$ values of the boundaries. Left boundary:
$$\begin{aligned}
z_1 &= \frac{x_1-\mu}{\sigma} \\\\
&= \frac{`r x1`-`r mu`}{`r sig`} \\\\
&= `r z1`
\end{aligned}$$

Right boundary:

$$\begin{aligned}
z_2 &= \frac{x_2-\mu}{\sigma} \\\\
&= \frac{`r x2`-`r mu`}{`r sig`} \\\\
&= `r z2`
\end{aligned}$$

We rephrase our problem into a **standard normal** probability problem:

$$P\left(`r x1` < X < `r x2`\right) ~=~ P(`r z1`<Z<`r z2`)$$

So, we just need to evaluate $P(`r z1`<Z<`r z2`)$.

From here we have a formula that lets us use the $z$ table. (We practiced this part before.)
$$\begin{aligned}
P(z_1 < Z < z_2) &= P(Z<z_2) - P(Z<z1) \\
P(`r z1`<Z<`r z2`) &= P(Z<`r z2`)-P(Z<`r z1`)\\
&= `r round(pnorm(z2),4)`-`r round(pnorm(z1),4)` \\
&= `r ba`
\end{aligned}$$


### Get $x$ from $P(X<x)$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
x = 103
z = (x-mu)/sig
la = round(pnorm(z),4)
```

Let's pretend we want to determine a boundary such that $X$ lands below that boundary `r la*100`% of the time. In other words, let's determine $x$ such that $P(X<x)=`r la`$.

We start by determining $z$ such that $P(Z<z)=`r la`$. We can do this by using the $z$ table.
$$P(Z<`r z`) = `r la` $$
So, we found the boundary as a $z$ score. 
$$z=`r z`$$
We calculate an associated $x$ boundary.
$$\begin{aligned}
x &= \mu+z\sigma \\
&= `r mu`+(`r z`)(`r sig`) \\
&= `r x`
\end{aligned}$$

So, we determined the boundary $x=`r x`$.
$$P(X<`r x`) ~=~ `r la` $$


### Get $x$ from $P(X>x)$

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
x = 91
z = (x-mu)/sig
la = round(pnorm(z),4)
ra = 1-la
```

Let's pretend we want to determine a boundary such that $X$ lands above that boundary `r ra*100`% of the time. In other words, let's determine $x$ such that $P(X>x)=`r ra`$.

We start by determining $z$ such that $P(Z>z)=`r ra`$. We can do this by using the $z$ table after using the complement rule to recognize that $P(Z<z)=1-`r ra`=`r la`$.
$$P(Z<`r z`) = `r la` $$
So, we found the boundary as a $z$ score. 
$$z=`r z`$$
We calculate an associated $x$ boundary.
$$\begin{aligned}
x &= \mu+z\sigma \\
&= `r mu`+(`r z`)(`r sig`) \\
&= `r x`
\end{aligned}$$

So, we determined the boundary $x=`r x`$.
$$P(X>`r x`) ~=~ `r ra` $$


### Get $d$ from $P(|X-\mu|<d)$. Also, report $\mu-d$ and $\mu+d$.

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
x = 102
z = (x-mu)/sig
la = round(pnorm(z),4)
ra = 1-la
ta = 2*ra
ca = round(1-ta,4)
```

As an example, let's find $d$, $x_1$, and $x_2$ such that:
$$P\left(\big|X-\mu\big|<d\right) ~=~ `r ca`$$
$$x_1 = \mu-d$$
$$x_2 = \mu+d$$

In other words, we want to find two boundaries equally far from the mean such that $X$ lands between them $`r ca*100`\%$ of the time.

We write an equivalent $Z$ probability.
$$P\left(\big|Z\big|<z\right) = `r ca` $$
We find $z$. To do that, we need to remember that if $P\left(\big|Z\big|<z\right) = `r ca` $, then
$$P(Z<z) ~=~ \frac{1+`rca`}{2} ~=~ `r la`$$
We determine the $z$ that works.
$$P(Z<`r z`) ~=~ `r la` $$
$$P(|Z|<`r z`) ~=~ `r ca` $$
We convert $z$ into $d$.
$$d ~=~ z\sigma ~=~ (`r z`)(`r sig`) ~=~ `r d` $$
Great, so $d=`r d`$. It is easy to now find the boundaries.
$$x_1 = `r mu`-`r d` = `r x1` $$
$$x_2 = `r mu`+`r d` = `r x2` $$


### Get $d$ from $P(|X-\mu|>d)$. Also, report $\mu-d$ and $\mu+d$.

Let's consider a normal distribution with a mean $\mu=100$ and a standard deviation $\sigma = 5$.
$$X \sim \mathcal{N}(100, \,5) $$

```{r}
x = 106
z = (x-mu)/sig
la = round(pnorm(z),4)
ra = 1-la
ta = round(2*ra,4)
ca = 1-ta
```

As an example, let's find $d$, $x_1$, and $x_2$ such that:
$$P\left(\big|X-\mu\big|>d\right) ~=~ `r ta`$$
$$x_1 = \mu-d$$
$$x_2 = \mu+d$$

In other words, we want to find two boundaries equally far from the mean such that $X$ lands between them $`r ca*100`\%$ of the time.

We write an equivalent $Z$ probability.
$$P\left(\big|Z\big|>z\right) = `r ta` $$
We find $z$. To do that, we need to remember that if $P\left(\big|Z\big|>z\right) = `r ta` $, then
$$P(Z<z) ~=~ \frac{2-`rta`}{2} ~=~ `r la`$$
We determine the $z$ that works.
$$P(Z<`r z`) ~=~ `r la` $$
$$P(|Z|>`r z`) ~=~ `r ta` $$
We convert $z$ into $d$.
$$d ~=~ z\sigma ~=~ (`r z`)(`r sig`) ~=~ `r d` $$
Great, so $d=`r d`$. It is easy to now find the boundaries.
$$x_1 = `r mu`-`r d` = `r x1` $$
$$x_2 = `r mu`+`r d` = `r x2` $$

