```{r,data,generation,echo=FALSE,results="hide"}
p = runif(1,0.07,0.4)
kmax = floor(log(0.0001/p,1-p))
kmaxmark = floor(log(0.001/p,1-p))
k = 0:kmax
klast = kmax+1
Pr = (1-p)^k*p
Prlast = 1-sum(Pr)
k = c(k,klast)
Pr = c(0,Pr,Prlast)

mu = round((1-p)/p,2)
pvar = (1-p)/p^2
psd = round(sqrt(pvar),2)

n = round(runif(1,6,9)^2)
x = rgeom(n,p)
i = 1:n

xbar = mean(x)
med = median(x)

data=data.frame(i,x)

write.csv(data,file="geometric_sample.csv",row.names=F)
```

Question
========

A sample's statistic typically approaches its corresponding population parameter, but with a small sample size there is regularly error. Let's explore this idea with an example.

A [geometric distribution](https://en.wikipedia.org/wiki/Geometric_distribution){target="blank"} is a well-studied discrete population. The spinner below represents a geometric distribution with the following population parameters.
$$\mu = `r mu`$$
$$\sigma = `r psd`$$

```{r echo=F,fig.width=5,fig.height=5}
par(mar=c(0,0,0,0))
theta = seq(0,2*pi,0.01)
plot(sin(theta),cos(theta),type="l",lwd=4,axes=F,ann=F)
cump = cumsum(Pr)
for(k in k){
  lines(c(0,sin(cump[k+1]*2*pi)),c(0,cos(cump[k+1]*2*pi)))
  if(k <= kmaxmark){
    ppp = (cump[k+1]+cump[k+2])/2
    text(0.6*sin(ppp*2*pi),0.6*cos(ppp*2*pi),k,cex=Pr[k+2]*30)
  }
}
polygon(0.05*cos(theta),0.05*sin(theta),col="red")
arrows(-0.5,-0.1,0.5,0.1,lwd=8)
arrows(-0.5,-0.1,0.5,0.1,lwd=7,col="red")
polygon(0.01*cos(theta),0.01*sin(theta),col="red")
```
\

That spinner was spun many times. The raw data is displayed below and [can be downloaded as a CSV](geometric_sample.csv){target="blank"}.

```{r echo=F,comment=""}
cat(paste(x,collapse=", "))
```

Please calculate the following sample statistics.

Answerlist
--------
* What is the sample size? (Determine $n$.)
* What is the sample mean? (Determine $\bar{x}$. You can round to nearest hundredth.)
* What is the **absolute difference** between the sample mean and population mean? (Calculate $|\bar{x}-\mu|$. You can round to nearest hundredth.)
* What is the sample median?
* What is the sample variance? (Determine $s^2$. Use Bessel's correction. You can round to nearest hundredth.)
* What is the sample standard deviation? (Determine $s$. Use Bessel's correction. You can round to nearest hundredth.)
* What is the **absolute difference** between the sample standard deviation and population standard deviation? (Calculate $|s-\sigma|$. You can round to nearest hundredth.)


Solution
========

Notice the sample mean and sample standard deviation do not match the population mean and population standard deviation exactly. 


Answerlist
--------
* `r n`
* `r xbar`
* `r abs(xbar-mu)`
* `r median(x)`
* `r var(x)`
* `r sd(x)`
* `r abs(sd(x)-psd)`


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num
exsolution: `r n`|`r xbar`|`r abs(xbar-mu)`|`r median(x)`|`r var(x)`|`r sd(x)`|`r abs(sd(x)-psd)`
exname: exam1q1
extol: 0|0.01|0.01|0|0.01|0.01|0.01
exextra[numwidth,logical]: TRUE