```{r,data,generation,echo=FALSE,results="hide"}
types = c("Bell","Uniform","Bimodal","Right-skew","Left-skew")

n = 200
beta1 = c(3,1,0.5,0.6,1.5)
beta2 = c(3,1,0.5,1.5,0.6)

bell = round(rbeta(n,beta1[1],beta2[1])*10,2)
uniform = round(rbeta(n,beta1[2],beta2[2])*10,2)
bimodal = round(rbeta(n,beta1[3],beta2[3])*10,2)
right_skew = round(rbeta(n,beta1[4],beta2[4])*10,2)
left_skew = round(rbeta(n,beta1[5],beta2[5])*10,2)

data = data.frame(bell,uniform,bimodal,right_skew,left_skew)

ord = sample(1:5)
type = sample(1:5,1)

i = 1:n
x = data[,type]

write.csv(data.frame(i,x),"make_hist.csv",row.names=F)
ans1 = paste0((type==ord)*1,collapse="")
ans2 = paste0((1:5==type)*1,collapse="")
```


Question
========

A sample of size $n=`r n`$ was taken from an unknown population. 

```{r echo=F, comment=NA}
y = lapply(x, function(xx) sprintf("%5.2f",xx))
data2 = matrix(y,ncol=10)
s = paste( apply(data2,1,function(xx) paste(xx,collapse=",")), collapse=",\n")
cat(s)
```

You can [download the data as a CSV](make_hist.csv). Determine which histogram visualizes the data, and describe the shape of the data.

```{r echo=F,fig.width=8,fig.height=5}
par(mfrow=c(2,3))
for(j in 1:5){
  par(mar=c(2,1,3,1))
  i = ord[j]
  hist(data[,i],axes=F,xlim=c(0,10),main=paste0("histogram ",j),col=rgb(0.9,0.7,0.7))
  box("figure")
}
```
\

* Which histogram is appropriate? \#\#ANSWER1\#\#
* Which shape description is appropriate?  \#\#ANSWER2\#\#


Answerlist
--------
* histogram 1
* histogram 2
* histogram 3
* histogram 4
* histogram 5
* Bell
* Uniform
* Bimodal
* Right-skew
* Left-skew


Solution
========

You should make a histogram. This is easy in R.

```{r}
x = read.csv("make_hist.csv")$x
hist(x)
```
\

Using a spreadsheet is [way more work](https://wikieducator.org/OpenOffice/Calc_3/Histogram){target="blank"}. 



Meta-information
================
extype: cloze
exclozetype: schoice|schoice
exsolution: `r ans1`|`r ans2`
exname: identify_histogram
extol: 0
exextra[numwidth]: 999.999
