---
title: "Proportion, histogram, mean, standard deviation, common distributions"
author: "Chad Worley"
date: "9/20/2021"
output:
  slidy_presentation:
    css: basic.css
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Proportion, binary data {.incremental}

When you have binary data, the proportion of success is the number of successes divided by the sample size.

$$\hat{p} = \frac{n_\text{s}}{n}$$

### Example

```{r binary_example}
set.seed(123)
x = sample(0:1,8,T,prob=c(0.6,0.4))
```
$$\mathbf{x} = `r x`$$
```{r showwork1,echo=T}
x = c(0,1,0,1,1,0,0,1)
n = length(x)
ns = sum(x)
phat = ns/n
data.frame(n,ns,phat)
```

$$n = `r n`$$
$$n_\text{s} = `r ns`$$
$$\hat{p} = \frac{`r ns`}{`r n`} = `r phat` $$

## Proportion, discrete or continuous data {.incremental}

If you have discrete or continuous data, you can determine the proportion of measurements within an interval.

$$\text{prop}[\cdots] = \frac{\#[\cdots]}{n}$$
where $\cdots$ will describe an interval.

### Example 

```{r example2}
set.seed(123)
x = rgeom(10,0.1)
```

Determine the proportion of measurements that are greater than 10.5.

$$\mathbf{x} = `r x`$$

```{r showwork2,echo=T}
x = c(10,16,29,1,3,30,25,4,4,5)
n = length(x)
ns = sum(x>10.5)
phat = ns/n
data.frame(n,ns,phat)
```

$$n = `r n`$$
$$\#[x>10.5] = n_s = `r ns` $$
$$\text{prop}[x>10.5] = \frac{\#[x>10.5]}{n} = \frac{`r ns`}{`r n`} = `r phat` $$

## Proportion (within interval) from histogram {.incremental}

```{r}
set.seed(1)
x = rexp(200,0.2)
myhist = hist(x,plot=F)
hist(x,labels=T,ylim=c(0,max(myhist$counts)*1.1))
```

- Let's assume $X$ is a continuous random variable, such that no measurements sit exactly on the breaks.
  - Note: it is helpful to determine the width of bars. In this case, we see there are 5 barwidths between 0 and 10. $10/5=2$, so the barwidth is equal to 2.
- How many measurements are less than 8?
  - We add the appropriate frequencies. $57+46+34+25 = `r 57+46+34+25`$
$$\#[x<8] = `r 57+46+34+25` $$
  - notice we do NOT count how many bars have frequencies less than 8!
- What proportion of the measurements are less than 8?
  - We need to know the sample size. $57+46+34+25+12+13+5+3+2+1+1+1=200$
  - We divide the count by the sample size.
  $$\text{prop}[x<8] = \frac{162}{200} = `r 162/200` $$
  

## Sample mean (average) {.incremental}

The sample mean is found by summing the measurements and dividing the sum by the sampe size.
$$\bar{x} = \frac{\sum x}{n} $$
In more technical situations, you may see more decorations to specify more precisely.

$$\bar{x} = \cfrac{\sum\limits_{i=1}^nx_i}{n} $$
Notice, with binary data, the sample mean is the sample proportion. (The sum of binary data equals the count of successes.)

## Example of finding mean {.incremental}

```{r findmeanexample}
set.seed(111)
x = rgeom(9,0.3)
```
### Problem

$$\mathbf{x} = `r x`$$
- Calculate the sample mean $\bar{x}$.

### Solution

$$\bar{x} = \frac{\sum x}{n} = \frac{`r paste(x,collapse="+")`}{`r length(x)`} = \frac{`r sum(x)`}{`r length(x)`} = `r round(sum(x)/length(x),4)` $$

```{r showworkmean,echo=T}
x = c(1,3,2,10,6,0,5,1,0)
xbar = mean(x)
print(xbar)
```

$$\bar{x}=`r round(mean(x),4)`$$

## The mean is the center of mass {.incremental}

```{r,fig.height=6,fig.width=8}
set.seed(99)
x = rgeom(30,0.4)
myh = hist(x,breaks=(-1):max(x)+0.5,plot=F)
plot(0,0,type="n",xlim=c(-0.5,max(x)+0.5),ylim=c(0,max(myh$counts)*1.2),axes=F,ann=F)
for(i in 1:length(myh$counts)){
  xx = i-1
  yy = myh$counts[i]
  if(yy>0){
    for(y in 1:yy){
      points(xx,y,pch=15,cex=5,col=rgb(0.8,0.9,0.9))
      points(xx,y,pch=0,cex=5,lwd=2)
    }
  }
}
points(mean(x),0,pch=17,col=rgb(0.7,0.3,0.3),cex=3)
abline(h=0.5)
mtext("center of mass",1,at=mean(x))
```

- More boxes on the left
- Boxes farther away on the right
- The platform would balance on the center of mass

## Median {.incremental}

- The median is another measure of center.
- To determine the median, first sort the data.
- Also, determine $n$, the sample size.
- If $n$ is odd (not a multiple of 2):
  - Find the sorted datum with index = $\frac{n+1}{2}$
- If $n$ is even (a multiple of 2):
  - Find the MEAN of sorted data with indices $\frac{n}{2}$ and $\frac{n+2}{2}$.

### Example 1

```{r}
set.seed(12)
x = round(rnorm(11,500,4))
n = length(x)
i = (n+1)/2
```

#### Problem

$$x=`r x`$$
Find the median of $x$. (Find $\text{median}(x)$)

#### Solution

- First, sort the data.
$$\text{sort}(x) = `r sort(x)` $$
- Determine $n$.
$$n = `r length(x)`$$
- Since $n$ is **odd**, we calculate the single index $\frac{n+1}{2}$.
$$i=\frac{n+1}{2}=\frac{`r length(x)`+1}{2}=`r (length(x)+1)/2`$$
We find which sorted datum in the `r i`th position.

$$\text{median}(x)=`r median(x)`$$
<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:1px solid #000000;
    padding: 4px;
}
</style>

```{r tab1, results="asis",echo=F}
i = 1:n
mydf = data.frame(i,sort(x))
colnames(mydf) = c("$i$","sort($x$)")
knitr::kable(mydf,format="html",table.attr='class="bordered-table"', align = "cc",row.names=F)
```

#### Solution using R

```{r,echo=T}
x = c(494,506,496,496,492,499,499,497,500,502,497)
median(x)
```

You could do it the long way.

```{r,echo=T}
x = c(494,506,496,496,492,499,499,497,500,502,497)
sortedx = sort(x)
n = length(x)
if(n%%2==1){ ## this checks if n is odd
  i = (n+1)/2
  median = sortedx[i]
} else { ## the only other option is n is even
  i1 = n/2
  i2 = (n+2)/2
  median = (sortedx[i1]+sortedx[i2])/2
}
median
```

## Median cont... {.incremental}

- The median is another measure of center.
- To determine the median, first sort the data.
- Also, determine $n$, the sample size.
- If $n$ is odd (not a multiple of 2):
  - Find the sorted datum with index = $\frac{n+1}{2}$
- If $n$ is even (a multiple of 2):
  - Find the MEAN of sorted data with indices $\frac{n}{2}$ and $\frac{n+2}{2}$.

### Example 2

```{r}
set.seed(13)
x = round(rnorm(14,750,7))
n = length(x)
i1 = n/2
i2 = (n+2)/2
```

#### Problem

$$x=`r x`$$
Find $\text{median}(x)$.

#### Solution

- First, sort the data.
$$\text{sort}(x) = `r sort(x)` $$
- Determine $n$.
$$n = `r length(x)`$$
- Since $n$ is **even**, we calculate two indices.
$$i_1=\frac{n}{2}=\frac{`r length(x)`}{2}=`r (length(x))/2`$$
$$i_2=\frac{n+2}{2}=\frac{`r length(x)`+2}{2}=`r (length(x)+2)/2`$$
We find the mean of the sorted data in the `r i1`th and `r i2`th positions.

$$\text{median}(x)=\frac{`r sort(x)[i1]`+`r sort(x)[i2]`}{`r n`}=`r median(x)`$$

```{r tab2, results="asis",echo=F}
i = 1:n
mydf = data.frame(i,sort(x))
colnames(mydf) = c("$i$","sort($x$)")
knitr::kable(mydf,format="html",table.attr='class="bordered-table"', align = "cc",row.names=F)
```

#### Solution using R

```{r,echo=T}
x = c(754,748,762,751,758,753,759,752,747,758,742,753,740,737)
median(x)
```

You could do it the long way.

```{r,echo=T}
x = c(754,748,762,751,758,753,759,752,747,758,742,753,740,737)
sortedx = sort(x)
n = length(x)
if(n%%2==1){ ## this checks if n is odd
  i = (n+1)/2
  median = sortedx[i]
} else { ## the only other option is n is even
  i1 = n/2
  i2 = (n+2)/2
  median = (sortedx[i1]+sortedx[i2])/2
}
median
```

## Common Distributions {.incremental}

There are 6 common distribution shapes I want you to name.

- Binary
- Uniform
- Bell (Normal)
- Bimodal
- Right-skewed
- Left-skewed

## Common Distribution: Binary {.incremental}

- Really, I just want to remind you that binary data exists. When the data are plotted in a histogram, there are just two bars.
```{r}
x = sample(0:1,100,T,c(0.4,0.6))
hist(x,breaks=c(-0.5,0.5,1.5),ann=F,axes=F)
axis(1,at=c(0,1))
```

## Common Distribution: Uniform {.incremental}

- There is [discrete uniform](https://en.wikipedia.org/wiki/Discrete_uniform_distribution){target="_blank"} or [continuous uniform](https://en.wikipedia.org/wiki/Continuous_uniform_distribution){target="_blank"}.
- A uniform histogram will have frequencies that are approximately equal.
- Symmetric
  - Standard dice rolled many times will produce a discrete uniform distribution.
```{r,echo=T}
x = sample(1:6,1000,T)
hist(x,breaks=seq(0.5,6.5,1),main="A 6-sided die is rolled 1000 times")
```

- The standard uniform distribution is an important continuous uniform distribution. The following spinner would produce it when spun many times.

```{r,fig.height=5,fig.width=5}
par(mar=c(0,0,0,0))
theta = seq(0,2*pi,length.out=360)
plot(sin(theta),cos(theta),type="l",axes=F,ann=F,lwd=3)
for(i in seq(1,99,1)){
  pr = i/100
  th = pr*2*pi
  if(i%%10==0){
    a = 0.9
    b = 0.82
    lwd = 2
    cex = 1.1
  } else if(i%%5==0) {
    a = 0.93
    b = 0.85
    lwd = 1.5
    cex = 0.8
  } else {
    a = 0.95
    b = 0.9
    lwd = 1
    cex = 0.5
  }
  if(pr<=0.5){
    srt = 90-pr*360
  } else {
    srt = 90-360*(pr-0.5)
  }
  lines(c(a*sin(th),sin(th)),c(a*cos(th),cos(th)),lwd=lwd)
  text(b*sin(th),b*cos(th),pr,cex=cex,srt=srt)
}
lines(c(0,0),c(0.7,1),lwd=4)
text(0,0.75,"1 | 0")
col = rgb(0.5,0.9,0.4)
polygon(0.08*cos(theta),0.08*sin(theta),lwd=2,col=col)
arrows(-0.8,-0.2,0.8,0.2,lwd=10,col=rgb(0,0,0))
arrows(-0.8,-0.2,0.8,0.2,lwd=5,col=col)
polygon(0.01*cos(theta),0.01*sin(theta),lwd=1,col=col)
```

```{r,echo=T}
x = runif(5000)
hist(x,breaks=seq(0,1,0.1))
```

## Common Distribution: Bell (Normal) {.incremental}

- Technically, there are multiple bell-shaped distributions. The normal distribution is just one of them.
- Symmetric
- However, we will use normal distributions A LOT.
- Sample averages (when repeatedly sampled) and sample totals (when repreatedly sampled) will tend to be approximately normal!

### Example: basketball practice

Let's say Jessica practices freethrows by shooting 100 attempts every day and Jessica has a 80% chance of making each shot. If we track 365 days...

```{r}
set.seed(123)
```

```{r,echo=T}
x = rbinom(365,100,0.8)
hist(x,
     main = "Histogram of totals from 365 days of practice",
     xlab = "Total points made in a day",
     breaks = seq(min(x)-0.5,
                  max(x)+0.5,
                  1)
     )
```



## Common Distribution: Bimodal {.incremental}

- A bimodal distribution has two (approximately equally tall) hills.
- Symmetric-ish
- Not many useful, realistic examples...
```{r}
x = rnorm(1000,10,2)*sample(c(-1,1),1000,T)+100
hist(x,breaks=20,main="Bimodal distribution")
```

## Common Distribution: Skewed right {.incremental}

- Data has a right-ward tail

```{r}
set.seed(123)
x = rbeta(1000,0.6,2)*20+rnorm(1000)+10
hist(x,main="Right-skew distribution")
```

- $\text{mean} > \text{median}$

## Common Distribution: Skewed left {.incremental}

- Data has a left-ward tail

```{r}
set.seed(1)
x = rbeta(1000,2,0.6)*20+rnorm(1000)+10
hist(x,main="Left-skew distribution")
```

- $\text{mean} < \text{median}$




## Measures of spread {.incremental}

We will discuss and use various measures of spread:

- [mean absolute deviation](https://en.wikipedia.org/wiki/Average_absolute_deviation)
- [range](https://en.wikipedia.org/wiki/Range_(statistics))
- [variance](https://en.wikipedia.org/wiki/Variance)
- [standard deviation](https://en.wikipedia.org/wiki/Standard_deviation)
- [interquartile range (IQR)](https://en.wikipedia.org/wiki/Interquartile_range)

## Measures of spread: mean of absolute deviations {.incremental}

A simple measure of spread is the average (mean) of absolute deviations from the sample mean.

$$\text{mean absolute deviation} = \frac{\sum |\mathbf{x}-\bar{x}|}{n} $$
If you prefer explicit limits...
$$\text{mean absolute deviation} = \frac{\sum\limits_{i=1}^n |x_i-\bar{x}|}{n} $$

### Example

```{r}
set.seed(111)
while(T){
  x = sample(1:9,8,T)+20
  n = length(x)
  xbar = mean(x)
  if((xbar*10)%%1==0){break}
}
```

$$x = `r x`$$

Determine the mean absolute deviation.

### Solution

- Determine the sample size by counting.
$$n=`r n`$$
- Add up the values to find the total.
$$\sum x = `r sum(x)`$$
- Find the sample mean.
$$\bar{x} = \frac{48}{8} = `r mean(x)`$$
- Determine the absolute deviations.
```{r}
xbar = mean(x)
deviation = x-xbar
absolute_deviation = abs(x-xbar)
mydf = data.frame(x,deviation,absolute_deviation)
knitr::kable(mydf,format="html",table.attr='class="bordered-table"', align = "ccc",row.names=F)
sad = sum(absolute_deviation)
mad = mean(absolute_deviation)
```
- Add up the absolute deviations.
$$\sum |\mathbf{x}-\bar{x}| = `r paste(absolute_deviation,collapse="+")` = `r sum(absolute_deviation)`$$
- Divide the sum (of absolute deviations) by the sample size.
$$\frac{\sum |\mathbf{x}-\bar{x}|}{n} = \frac{`r sum(absolute_deviation)`}{`r n`} = `r mad`$$

### Solution using R

```{r,echo=T}
x = c(28,29,28,21,27,25,21,29)
mean(abs(x-mean(x)))
```

### Visualization

```{r}
library(latex2exp)
```

```{r,fig.width=8,fig.height=8}
par(mar=c(0,0,1.2,0))
xbar = mean(x)
n = length(x)
i = 1:n
ran = max(x)-min(x)
tad = sum(abs(x-mean(x)))
MAD = tad/n
plot(x,n-i,axes=F,ann=F,ylim=c(-3,n),xlim=c(xbar-tad/2,xbar+tad/2),cex=3,lwd=3)
lines(c(xbar,xbar),c(-0.5,n-0.5),lty=2,col=rgb(0.9,0.5,0),lwd=2)
lines(c(xbar-MAD,xbar-MAD),c(-0.5,n-0.5),lty=2,col=rgb(0.2,0.2,0.2),lwd=2)
lines(c(xbar+MAD,xbar+MAD),c(-0.5,n-0.5),lty=2,col=rgb(0.2,0.2,0.2),lwd=2)
lines(c(rep(xbar,3),rep(xbar+MAD,3)),
      c(n,n+0.2,n+0.1,n+0.1,n+0.2,n),
      lwd=4,col=rgb(0.5,0,0.2))
lines(c(rep(xbar,3),rep(xbar-MAD,3)),
      c(n,n+0.2,n+0.1,n+0.1,n+0.2,n),
      lwd=4,col=rgb(0.5,0,0.2))
lines(c(min(x),max(x)),c(-0.5,-0.5))
for(ii in min(x):max(x)){
  lines(c(ii,ii),c(-0.5,-0.7))
  text(ii,-1,ii)
}
mtext(TeX("$\\bar{x}$"),3,at=xbar,line=0,col=rgb(0.9,0.5,0),cex=1.3)
mtext(TeX("$\\bar{x}+$MAD"),3,at=xbar+MAD,line=0,col=rgb(0.2,0.2,0.2),cex=1.3)
mtext(TeX("$\\bar{x}-$MAD"),3,at=xbar-MAD,line=0,col=rgb(0.2,0.2,0.2),cex=1.3)
cols = c(rgb(0,0,0),rgb(1,0,0),rgb(0,1,0),
         rgb(0,0,1),rgb(0.6,0.6,0),rgb(0,0.6,0.6),
         rgb(0.6,0,0.6),rgb(0.2,0.3,0.1))
for(i in 1:n){
  ln = n-i
  lines(c(rep(mean(x),3),rep(x[i],3)),
        c(ln-0.1,ln+0.1,ln,ln,ln+0.1,ln-0.1),
        col=cols[i],lwd=5)
}
xxx = xbar-tad/2
for(i in 1:n){
  x1 = xxx
  x2 = xxx+abs(x[i]-xbar)
  lines(c(x1+0.02,x1+0.02,x1+0.02,x2-0.02,x2-0.02,x2-0.02),c(-2.1,-1.9,-2,-2,-1.9,-2.1),
        col=cols[i],lwd=5)
  xxx = x2
}

lines(c(xbar-tad/2,xbar+tad/2),c(-3,-3),lwd=4,col=rgb(0.5,0.2,0.2))
for(i in 0:n){
  lines(rep(xbar-tad/2+i*MAD,2),c(-3.1,-2.9),lwd=4,col=rgb(0.5,0,0.2))
}

```

## Measures of spread: range {.incremental}

- The range is simply the difference between the maximum and minimum.
$$\text{range} = \max(\mathbf{x}) - \min(\mathbf{x})$$

### Example

```{r}
set.seed(111)
while(T){
  x = sample(1:9,8,T)+20
  n = length(x)
  xbar = mean(x)
  if((xbar*10)%%1==0){break}
}
```

$$\mathbf{x}=`r x`$$

- Find the range of $\mathbf{x}$.

### Solution

$$`r max(x)`-`r min(x)` = `r max(x)-min(x)` $$

### Notes

- The range of a sample is never larger than the range of the population.
- The range of a sample is often smaller than the range of the population.
- We say sample range is a biased estimator of population range.


## Measures of spread: variance {.incremental}

- variance without [Bessel correction](https://en.wikipedia.org/wiki/Bessel%27s_correction)
$$\text{var(x)} = \frac{\sum(\mathbf{x}-\bar{x})^2}{n}$$
- variance with [Bessel correction](https://en.wikipedia.org/wiki/Bessel%27s_correction)
$$\text{var(x)} = \frac{\sum(\mathbf{x}-\bar{x})^2}{n-1}$$
- The Bessel correction makes an unbiased estimator.

## Measures of spread: standard deviation {.incremental}

- sample standard deviation without [Bessel correction](https://en.wikipedia.org/wiki/Bessel%27s_correction)
$$s_n = \sqrt{\frac{\sum(\mathbf{x}-\bar{x})^2}{n}}$$
- sample standard deviation with [Bessel correction](https://en.wikipedia.org/wiki/Bessel%27s_correction)
$$s = \sqrt{\frac{\sum(\mathbf{x}-\bar{x})^2}{n-1}}$$

- almost always you will use Bessel correction when you have a sample