```{r,data,generation,echo=FALSE,results="hide"}
while(T){
  type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
  n = sample(7:15,1)
  beta1 = c(1,3,0.2,0.5,2)[type]
  beta2 = c(1,3,0.2,2,0.5)[type]
  bottom = sample(seq(20,70,5),1)
  width = sample(seq(5,30,5),1)
  i = 1:n
  x = round(rbeta(n,beta1,beta2)*width+bottom,2)
  xbar = mean(x)
  devs = x-xbar
  adevs = abs(devs)
  MAD = mean(adevs)
  if(round(xbar,2)==xbar){
    break
  }
}

mydf = data.frame(i,x,devs,adevs)
colnames(mydf) = c("$$i$$","$$\\mathbf{x}$$","deviations = $\\mathbf{x}-\\bar{x}$","AbsDev = $\\vert\\mathbf{x}-\\bar{x}\\vert$")
```


Question
========

A sample was gathered. 

$$\mathbf{x} = `r paste(lapply(x,function(x) sprintf("%.2f",x)),collapse=", ")` $$


```{r echo=F}
write.csv(data.frame(i,x),file="get_MAD.csv", quote = FALSE, row.names = FALSE)
```

You can download the data as a [CSV file](get_MAD.csv).

Determine the sample's [mean absolute deviation (MAD) around the mean](https://en.wikipedia.org/wiki/Average_absolute_deviation#Mean_absolute_deviation_around_a_central_point){target="blank"}.  You can round your answer to the [hundredths place](https://en.wikipedia.org/wiki/Orders_of_magnitude_(numbers)#10%E2%88%922){target="blank"}.


Solution
========

First, determine the mean of the sample.

$$\bar{x} = \frac{\sum \mathbf{x}}{n} =`r mean(x)` $$

Determine the absolute deviations (distances from $\bar{x}$).

`r knitr::kable(mydf)`


Now, take the mean of the absolute deviations.

$$\text{MAD} = \frac{\sum \vert\mathbf{x}-\bar{x}\vert}{n} = `r MAD` $$

### spreadsheet

```{r echo=F}
xbar = mean(x)
sol = data.frame(i,x)

sol = rbind(sol,
            c("",""),
            c("","xbar"),
            c("",paste0("=average(B2:B",n+1,")",collapse=""))
            )

deviations = c(paste0("=B",2:(n+1),"-B$",n+4),rep("",3))
absdev = c(paste0("=abs(C",2:(n+1),")"),
           "",
           "MAD",
           paste0("=average(D2:D",n+1,")"))
sol = cbind(sol,deviations,absdev)
write.csv(sol,file="get_MAD_sol.csv", row.names=F)

```

You can do this with [a spreadsheet](get_MAD_sol.csv).

```{r fig.height=6,fig.width=8,echo=F}
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol)
```
\


### R

You can do this with R

```{r}
x = read.csv("get_MAD.csv")$x
xbar = mean(x)
deviations = x-xbar
AbsDev = abs(deviations)
MAD = mean(AbsDev)
x
xbar
deviations
AbsDev
MAD
```


Meta-information
================
extype: num
exsolution: `r MAD`
exname: get_MAD
extol: 0.01
exextra[numwidth]: 999.999