```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)

while(T){
  n = sample(40:300,1)
  suc = sample(1:n,1)
  p = round(suc/n,4)
  gamma = sample(50:99/100,1)
  se = sqrt(p*(1-p)/n)
  zstar = qnorm(gamma+(1-gamma)/2)
  zstar_rnd = round(zstar,2)
  ME = zstar*se
  LB = p-ME
  UB = p+ME
  if(n*p>10 & n*(1-p)>10 & UB-LB>0.02){
    break
  }
}

LA = gamma+(1-gamma)/2
LA2 = LA

library(knitr)
default_source_hook <- knit_hooks$get('source')
knit_hooks$set(source = function(x, options) {
  x <- gsub('LA2',paste0(LA),x)
  default_source_hook(x, options)
})
```

Question
========

In $n=`r n`$ trials, there were $n_\text{s}=`r suc`$ successes, so the sample proportion is $\hat{p}=\frac{`r suc`}{`r n`}=`r p`$. You are tasked with determining the confidence interval (of the population proportion) with confidence level $\gamma=`r gamma`$.

To do this, you first determine $z^\star$ such that $P(|Z|<z^\star)=`r gamma`$. Then, evaluate $\hat{p}\pm z^\star\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$ to determine the boundaries of the confidence interval.

(You can round answers to the hundredths place.)

Answerlist
--------
* Determine the lower boundary of the confidence interval by using $\hat{p}- z^\star\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$.
* Determine the upper boundary of the confidence interval by using $\hat{p}+ z^\star\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$.


Solution
========

[video](https://vimeo.com/531948610){target="_blank"}

First, determine $z^\star$ such that $P(|Z|<z^\star)=`r gamma`$. It can help to draw a sketch.

```{r middle,fig.width=5,fig.height=3}
library(latex2exp)
zs = seq(-3,3,0.001)
par(mar=c(2,0,0,0))
plot(zs,dnorm(zs),type="l",lwd=2,axes=F,ann=F)
abline(h=0)
polygon(c(-zstar,zs[abs(zs)<=zstar],zstar),
        c(0,dnorm(zs[abs(zs)<=zstar]),0),
        col=rgb(1,0.5,0.5,0.5))
axis(1,at=c(-3,-zstar,0,zstar,3),labels=c("-3",TeX("$$-z^{*}$$"),"0",TeX("$$z^{*}$$"),"3"))
text(0,0.15,gamma,cex=2,col=rgb(0.5,0.2,0.2,1))
```
\

The entire area under the density curve is 1. Thus, we can determine the areas of the tails (using symmetry).
$$\text{area of each tail} = \frac{1-\gamma}{2} = \frac{1-`r gamma`}{2} = `r (1-gamma)/2`$$

```{r allthree,fig.width=5,fig.height=3}
library(latex2exp)
zs = seq(-3,3,0.001)
par(mar=c(2,0,0,0))
plot(zs,dnorm(zs),type="l",lwd=2,axes=F,ann=F)
abline(h=0)
polygon(c(-zstar,zs[abs(zs)<=zstar],zstar),
        c(0,dnorm(zs[abs(zs)<=zstar]),0),
        col=rgb(1,0.5,0.5,0.5))
axis(1,at=c(-3,-zstar,0,zstar,3),labels=c("-3",TeX("$$-z^{*}$$"),"0",TeX("$$z^{*}$$"),"3"))
text(0,0.15,gamma,cex=2,col=rgb(0.5,0.2,0.2,1))
polygon(c(-3,zs[zs<=-zstar],-zstar),
        c(0,dnorm(zs[zs<=-zstar]),0),
        col=rgb(0.5,1,0.5,0.5))
text(-2.5,0.15,(1-gamma)/2,cex=2,col=rgb(0.2,0.5,0.2,1))
polygon(c(zstar,zs[zs>=zstar],3),
        c(0,dnorm(zs[zs>=zstar]),0),
        col=rgb(0.5,0.5,1,0.5))
text(2.5,0.15,(1-gamma)/2,cex=2,col=rgb(0.2,0.2,0.5,1))
```
\

Determine a leftward area with boundary $z^\star$.
$$P(Z<z^\star)=\gamma + \frac{1-\gamma}{2}= `r gamma`+`r (1-gamma)/2`=`r gamma+(1-gamma)/2`$$

```{r left,fig.width=5,fig.height=3}
library(latex2exp)
zs = seq(-3,3,0.001)
par(mar=c(2,0,0,0))
plot(zs,dnorm(zs),type="l",lwd=2,axes=F,ann=F)
abline(h=0)
polygon(c(-3,zs[zs<=zstar],zstar),
        c(0,dnorm(zs[zs<=zstar]),0),
        col=rgb(1,1,0.5,0.5))
axis(1,at=c(-3,0,zstar,3),labels=c("-3","0",TeX("$$z^{*}$$"),"3"))
text(0,0.15,gamma+(1-gamma)/2,cex=2,col=rgb(0.5,0.5,0.2,1))
```
\

We rephrase the puzzle. We wish to determine $z^\star$ such that $P(Z<z^\star)=`r gamma+(1-gamma)/2`$. This is easy to determine using a $z$-table, `qnorm` in R, `norm.inv` in spreadsheets, or other methods for evaluating the [quantile function](https://en.wikipedia.org/wiki/Normal_distribution#Quantile_function){target="blank"}.

```{r, echo=T}
## Using R's qnorm:
qnorm(LA2)
```

$$P(Z<`r zstar_rnd`)=`r gamma+(1-gamma)/2`$$
In the original phrasing,

$$P(|Z|<`r zstar_rnd`)=`r gamma`$$

So, $z^\star = `r zstar_rnd`$. Now, use the given expressions to evaluate the boundaries.


Answerlist
--------
* Lower bound = $~`r p`-`r zstar_rnd`\sqrt{\frac{(`r p`)(1-`r p`)}{`r n`}} ~=~ `r round(LB,2)`$
* Upper bound = $~`r p`+`r zstar_rnd`\sqrt{\frac{(`r p`)(1-`r p`)}{`r n`}} ~=~ `r round(UB,2)`$


Meta-information
================
extype: cloze
exsolution: `r LB`|`r UB`
exclozetype: num|num
exname: confidence interval
extol: 0.01

