```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
n = sample(c(60,70,75,80,90,100,110,120,125,130,150),1)
beta1 = c(1,10,0.2,0.5,2)[type]
beta2 = c(1,10,0.2,2,0.5)[type]
bottom = sample(seq(20,70,5),1)
width = sample(seq(5,30,5),1)
i = 1:n
x = rbeta(n,beta1,beta2)*width+bottom
myhist = hist(x,breaks=5)
brks = myhist$breaks
counts = myhist$counts

library(knitr)
library(stringr)
default_source_hook <- knit_hooks$get('source')

knit_hooks$set(source = function(x, options) {
  x <- str_replace_all(x, pattern = 'brks', paste0("c(",paste0(brks,collapse=","),")"))
  default_source_hook(x, options)
})
```


Question
========

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:1px solid #000000;
    padding: 4px;
}
</style>

```{r echo=F}
write.csv(data.frame(i,x),file="make_freq_dist.csv",row.names = FALSE)
```

A random procedure generated many measurements: [download data](make_freq_dist.csv)

Please complete the [frequency distribution](https://en.wikipedia.org/wiki/Frequency_distribution){target="blank"} using breaks `r brks`:

```{r tab1, results="asis",echo=F}
interval = paste(brks[1:length(counts)]," to ", brks[2:(length(counts)+1)])
frequency = paste0("##ANSWER",1:length(counts),"##")
knitr::kable(data.frame(interval,frequency),format="html",table.attr='class="bordered-table"', align = "cc",row.names=F)
```


Solution
========

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:1px solid #000000;
    padding: 4px;
}
</style>

You will want to use a computer to answer these questions. 

### Spreadsheet

[video (using a spreadsheet)](https://vimeo.com/531556006){target="_blank"}

In a spreadsheet, open the data, add the breaks as a column; then, use the [``FREQUENCY function``](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_FREQUENCY_function){target="blank"}.



### R

[video (using R)](https://vimeo.com/531555988){target="blank"}

In R, open the data and use the [``hist``](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/hist) function. You supply the breaks and read the counts:

```{r}
mydata = read.csv("make_freq_dist.csv")
```

```{r}
x = mydata$x
```

```{r fig.show="hide"}
myhist = hist(x,breaks=brks)
```

```{r}
myhist$counts
```



```{r tab2, results="asis", echo=F}
interval = paste(brks[1:length(counts)]," to ", brks[2:(length(counts)+1)])
frequency = counts
knitr::kable(data.frame(interval,frequency),format="html",table.attr='class="bordered-table"', align = "cc",row.names=F)
```



Answerlist
--------
`r paste0("* ",counts,collapse="\n")`





Meta-information
================
extype: cloze
exclozetype: `r paste0(rep("num",length(counts)),collapse="|")`
exsolution: `r paste0(counts,collapse="|")`
exname: make_frequency_distribution
extol: 0
exextra[numwidth,numeric]: 10
