```{r data generation, echo = FALSE, results = "hide"}
while(T){
  n = sample(8:25,1)
  x = sample(2:5,1)
  ans = factorial(n)/factorial(n-x)
  if(ans < 1e8){
    break
  }
}

qn = sample(1:10,1)
# qn = 1

lto = function(l){
  s = ''
  if(length(l)==2){
    s = paste(l[1],"and",l[2])
  } else {
    for(i in 1:(length(l)-1)){
      s = paste(s,l[i],", ",sep="")
    }
    s = paste(s,"and",l[length(l)])
  }
  return(s)
}

poses = c("a president","a CFO","a vice president","a secretary","a manager")
ps = sample(poses,x)
q = paste("A company needs to select ",lto(ps),". Each position will be held by a different person. The company is considering the same pool of ",n," applicants for each position. How many possibilities exist?",sep="")

sss = ''
sss = paste(sss,n)
for(i in (n-1):(n-x+1)){
  sss = paste(sss,i,sep="\\cdot")
}

options(scipen=100)
```

Question
========

`r q`

Solution
========

This scenario describes a permutations problem (order matters). We are considering the nonrepeating sequences of size `r x` from a set of size `r n`.

$$\begin{aligned}
{_nP_r} &= \frac{n!}{(n-r)!} \\\\
n &= `r n` \\\\
r &= `r x` \\\\
{_{`r n`}P_{`r x`}} &= \frac{`r n`!}{(`r n`-`r x`)! } \\\\
&= \frac{`r n`!}{`r n-x`!} \\\\
&= `r sss` \\\\
&= `r ans`
\end{aligned}$$


```{r, echo=F}
samples <- vector(mode = "list", length = 10)
ss <- vector(mode = "list", length = 10)
for(i in 1:10){
  while(T){
    sam = sort(sample(1:n,x))
    if(!(all(sam %in% samples))){
      s = ''
      samples[[i]] = list(sam)
      daord = sample(1:x)
      cnt = 1
      for(j in 1:n){
        if(j %in% sam){
          s = paste(s,daord[cnt],collapse='')
          cnt = cnt+1
        } else {
          s = paste(s,'0',collapse='')
        }
      }
      ss[[i]] = s
      break
    }
  }
}
```

If you had a lot of time, you could list out all possibilities (using 1 for `r ps[1]`, 2 for `r ps[2]`...):

| Count | Possibility |
|:----:|:-----:|
| 1 | `r ss[[1]]` |
| 2 | `r ss[[2]]` |
| 3 | `r ss[[3]]` |
| 4 | `r ss[[4]]` |
| 5 | `r ss[[5]]` |
| $\vdots$ | $\vdots$ |
| `r ans-4` | `r ss[[6]]` |
| `r ans-3` | `r ss[[7]]` |
| `r ans-2` | `r ss[[8]]` |
| `r ans-1` | `r ss[[9]]` |
| `r ans` | `r ss[[10]]` |


Of course, you'd want to be more systematic than that.

Meta-information
============
extype: num
exsolution: `r ans`
exname: permutations
extol: 0.01
