
```{r data generation, echo = FALSE, results = "hide"}
library(exams)
options(scipen=100)
types = c("A","B","C","D")
fruit = c("apple","orange","banana","mango","plum","grapefruit","kiwi","lemon","papaya","apricot","persimmon","watermelon","pineapple","coconut")
mass = c(180,180,180,250,90,240,80,100,450,50,170,9000,2000,700)
is = sample(1:length(mass),4)
while(T){
  is = sample(1:length(mass),4)
  f = fruit[is]
  fs = paste(f,"s",sep="")
  mus = round(mass[is]*runif(4,0.7,1.3))
  sigs = round(mus/runif(4,6,20))
  ns = sample(5:12,4)^2
  ses = sigs/sqrt(ns)
  ps = runif(4,0.02,0.98)
  zs = round(qnorm(ps),2)
  xbars = signif(mus+zs*ses,4)
  zs = round((xbars-mus)/ses,2)
  ps = round(pnorm(zs),4)
  if(sum(zs>0)>=1 & sum(zs<0)>=1 & length(unique(abs(zs)))==4 & min(c(dist(abs(zs),method = "manhattan"))) > 0.1){
    break
  }
}

scs1 = mchoice2string(zs==max(zs))
scs2 = mchoice2string(zs==min(zs))
scs3 = mchoice2string(abs(zs)==min(abs(zs)))
scs4 = mchoice2string(abs(zs)==max(abs(zs)))


data = data.frame(f,mus,sigs,ns,xbars)
colnames(data) = c("fruit",
                   "mu",
                   "sigma",
                   "n",
                   "xbar")

write.csv(data,"fruit.csv",row.names=F)
```

Question
========

A farm produces 4 types of fruit: `r fs[1]`, `r fs[2]`, `r fs[3]`, and `r fs[4]`. The fruits' masses have population parameters dependent on the type of fruit. (All values are in grams.)

| _ Type of fruit _ | _ Population mean ($\mu$) _ | _ Population standard deviation ($\sigma$) _|
|:---:|:---:|:---:|
|`r fs[1]` | `r mus[1]` | `r sigs[1]` | 
|`r fs[2]` | `r mus[2]` | `r sigs[2]` |
|`r fs[3]` | `r mus[3]` | `r sigs[3]` |
|`r fs[4]` | `r mus[4]` | `r sigs[4]` |

A sample of each type is weighed. The results are shown below.

| _ Type of fruit _ | _ Sample size ($n$) _ | _ Sample mean ($\bar{x}$) _ |
|:---:|:---:|:---:|
|`r f[1]` | `r ns[1]` | `r xbars[1]` |
|`r f[2]` | `r ns[2]` | `r xbars[2]` |
|`r f[3]` | `r ns[3]` | `r xbars[3]` |
|`r f[4]` | `r ns[4]` | `r xbars[4]` |

The population parameters and sample statistics can be [downloaded as a csv](fruit.csv){target="blank"}.

For each sample, determine the mean's standard score and the mean's [cumulative probability](https://en.wikipedia.org/wiki/Cumulative_distribution_function){target="blank"} by assuming the requirements for the [central limit theorem](https://en.wikipedia.org/wiki/Central_limit_theorem){target="blank"} are met. Then determine which sample mean is most unusually large, most unusually small, most typically sized, and most unusually sized.

The standard score of a sample mean:
$$z = \frac{\bar{x}-\mu}{\sigma/\sqrt{n}}$$
Some people prefer to call this denominator [the standard error](https://en.wikipedia.org/wiki/Standard_error){target="blank"}.
$$\text{standard error of mean} = \text{SEM} = \frac{\sigma}{\sqrt{n}} $$
So the standard score ($z$-score) of a sample mean can also be expressed as
$$z = \frac{\bar{x}-\mu}{\text{SEM}} $$

Answerlist
--------
* Calculate a $z$-score for the `r f[1]` sample mean. (Round to the nearest hundredth.)
* What proportion of `r f[1]` samples of size $n=`r ns[1]`$ have a smaller mean than the `r f[1]` sample?
* Calculate a $z$-score for the `r f[2]` sample mean. (Round to the nearest hundredth.)
* What proportion of `r f[2]` samples of size $n=`r ns[2]`$ have a smaller mean than the `r f[2]` sample?
* Calculate a $z$-score for the `r f[3]` sample mean. (Round to the nearest hundredth.)
* What proportion of `r f[3]` samples of size $n=`r ns[3]`$ have a smaller mean than the `r f[3]` sample?
* Calculate a $z$-score for the `r f[4]` sample mean. (Round to the nearest hundredth.)
* What proportion of `r f[4]` samples of size $n=`r ns[4]`$ have a smaller mean than the `r f[4]` sample?
* The `r f[1]` sample mean is most unusually large.
* The `r f[2]` sample mean is most unusually large.
* The `r f[3]` sample mean is most unusually large.
* The `r f[4]` sample mean is most unusually large.
* The `r f[1]` sample mean is most unusually small.
* The `r f[2]` sample mean is most unusually small.
* The `r f[3]` sample mean is most unusually small.
* The `r f[4]` sample mean is most unusually small.
* The `r f[1]` sample mean is most typically sized.
* The `r f[2]` sample mean is most typically sized.
* The `r f[3]` sample mean is most typically sized.
* The `r f[4]` sample mean is most typically sized.
* The `r f[1]` sample mean is most unusually sized.
* The `r f[2]` sample mean is most unusually sized.
* The `r f[3]` sample mean is most unusually sized.
* The `r f[4]` sample mean is most unusually sized.

Solution
========

[video](https://vimeo.com/531945861){target="_blank"}

The formula to determine the $z$-score of a **sample mean** is the ratio with numerator the difference between measurement and population mean and denominator the standard error of the mean.

$$z = \frac{\bar{x}-\mu}{\sigma/\sqrt{n}} $$

The highest $z$-score (furthest right on [number line](https://en.wikipedia.org/wiki/Number_line){target="blank"}) corresponds to the most unusually large sample mean. 

The smallest $z$-score (furthest left on [number line](https://en.wikipedia.org/wiki/Number_line){target="blank"}) corresponds to the most unusually small sample mean.

The smallest [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$-score corresponds to the most usually sized sample mean.

The largest [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$-score corresponds to the most unusually sized sample mean.


### Spreadsheet:

By using the formulas, you should get a spreadsheet like the one displayed here:

```{r fig.width=10,fig.height=5,echo=F}
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

data2 = data.frame(f,mus,sigs,ns,xbars,signif(ses,4),zs,ps,abs(zs))
colnames(data2) = c("fruit",
                   "mu",
                   "sigma",
                   "n",
                   "xbar",
                   "SEM",
                   "z",
                   "P(Z<z)",
                   "abs(z)")

draw_ss(data2)
```
\

### To do this in R:

```{r echo=TRUE}
data = read.csv("fruit.csv", as.is=TRUE)
fruit = data$fruit
mu = data$mu
sigma = data$sigma
n = data$n
xbar = data$xbar

SEM = sigma/sqrt(n)
z = round( (xbar-mu)/SEM ,2)
cum_prob = round( pnorm(z) ,4)
data.frame(data,SEM,z,cum_prob,abs(z))

# Most unusually large
fruit[z==max(z)]

# Most unusually small
fruit[z==min(z)]

# Most typically sized
fruit[abs(z)==min(abs(z))]

# Most unusually sized
fruit[abs(z)==max(abs(z))]
```


Answerlist
--------
* We use the formula for calculating the `r f[1]` sample-mean $z$ score.
$$z = \frac{`r xbars[1]`-`r mus[1]`}{`r sigs[1]`/\sqrt{`r ns[1]`}} = `r zs[1]` $$
* Use the $z$ table to find the cumulative probability.
$$P(Z<`r zs[1]`) = `r ps[1]` $$
* We use the formula for calculating the `r f[2]` sample-mean $z$ score.
$$z = \frac{`r xbars[2]`-`r mus[2]`}{`r sigs[2]`/\sqrt{`r ns[2]`}} = `r zs[2]` $$
* Use the $z$ table to find the cumulative probability.
$$P(Z<`r zs[2]`) = `r ps[2]` $$
* We use the formula for calculating the `r f[3]` sample-mean $z$ score.
$$z = \frac{`r xbars[3]`-`r mus[3]`}{`r sigs[3]`/\sqrt{`r ns[3]`}} = `r zs[3]` $$
* Use the $z$ table to find the cumulative probability.
$$P(Z<`r zs[3]`) = `r ps[3]` $$
* We use the formula for calculating the `r f[4]` sample-mean $z$ score.
$$z = \frac{`r xbars[4]`-`r mus[4]`}{`r sigs[4]`/\sqrt{`r ns[4]`}} = `r zs[4]` $$
* Use the $z$ table to find the cumulative probability.
$$P(Z<`r zs[4]`) = `r ps[4]` $$
* We determine the maximum $z$ score is $`r max(zs)`$, which belongs to the `r f[zs==max(zs)]`.
* .
* .
* .
* We determine the minimum $z$ score is $`r min(zs)`$, which belongs to the `r f[zs==min(zs)]`.
* .
* .
* .
* We determine the minimum [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$ score is $`r min(abs(zs))`$, which belongs to the `r f[abs(zs)==min(abs(zs))]`. 
* .
* .
* .
* We determine the maximum [**absolute**](https://en.wikipedia.org/wiki/Absolute_value){target="blank"} $z$ score is $`r max(abs(zs))`$, which belongs to the `r f[abs(zs)==max(abs(zs))]`.
* .
* .
* .


Meta-information
============
extype: cloze
exclozetype: num|num|num|num|num|num|num|num|schoice|schoice|schoice|schoice
exsolution: `r zs[1]`|`r ps[1]`|`r zs[2]`|`r ps[2]`|`r zs[3]`|`r ps[3]`|`r zs[4]`|`r ps[4]`|`r scs1`|`r scs2`|`r scs3`|`r scs4`
exname: compare zs
extol: 0.011|0.01|0.011|0.01|0.011|0.01|0.011|0.01|0|0|0|0
