```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)

w = read.csv("yob2019_w_header.txt",as.is=T)
n = sample(10:20,1)
names = sort(sample(w$name,n,replace=F,prob=w$freq))

speeds = runif(n,3,13) #mph
times = round(1/speeds*60*60)

sigdel = runif(1,20,50)
SEdel = sigdel/sqrt(n)
MEdel = 2*SEdel
mudel = sample(c(-MEdel,MEdel),1)
del = rnorm(n,mudel,sigdel)
times2 = round(times+del)

x = data.frame(names,times,times2)
colnames(x) = c("name","x1","x2")
```

Question
========

```{r CSS, echo = FALSE, results = "asis"}
writeLines(readLines("bordered_table_style.html"))
```

A study asked participants to time a mile run (in seconds). After a month, the same participants timed another mile run. You are asked to perform a paired-data $t$ test to investigate whether fitness changed.

```{r tab1, results="asis"}
knitr::kable(x,format="html",table.attr='class="bordered-table"',
             align = "lcc",row.names=F)
```

To do this, it can be helpful to first determine the list of differences.
$$d_i = x_{2,i}-x_{1,i}$$

Then, determine the mean ($\bar{d}$) and standard deviation ($s_d$) of the differences. The $t$ score is then calculated to determine the $p$-value.

$$p\text{-value} = P\left(|T|>\frac{\overline{d}-0}{s_d\sqrt{n}}\right) $$


Solution
========



Meta-information
================
extype: cloze
exsolution: 10
exclozetype: num
exname: hypothesis test paired
extol: 0
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE
