```{r,data,generation,echo=FALSE,results="hide"}
n = 10 #must be bigger than 8, or change xcom to replacement sampling with T
f = sample(1:9,1)*10
while(T){
  x = c(59,53,55,58,52,58,58,58,58,58)
  
  b1 = c(50.5,58.5)  #boundaries
  c1 = mean(b1) #center of boundaries
  d1 = b1[2]-c1 #distance of boundaries from center
  nb1 = sum(x>b1[1] & x<b1[2]) #number between
  no1 = sum(x<b1[1] | x>b1[2]) #number outside
  
  b2 = c(51.5,54.5)  #boundaries
  c2 = mean(b2) #center of boundaries
  d2 = b2[2]-c2 #distance of boundaries from center
  nb2 = sum(x>b2[1] & x<b2[2]) #number between
  no2 = sum(x<b2[1] | x>b2[2]) #number outside
  
  b3 = c(55.5,59.5)  #boundaries
  c3 = mean(b3) #center of boundaries
  d3 = b3[2]-c3 #distance of boundaries from center
  nb3 = sum(x>b3[1] & x<b3[2]) #number between
  no3 = sum(x<b3[1] | x>b3[2]) #number outside
  
  b4 = c(51.5,59.5)  #boundaries
  c4 = mean(b4) #center of boundaries
  d4 = b4[2]-c4 #distance of boundaries from center
  nb4 = sum(x>b4[1] & x<b4[2]) #number between
  no4 = sum(x<b4[1] | x>b4[2]) #number outside

  b5 = c(56.5,58.5) #boundaries
  c5 = mean(b5) #center of boundaries
  d5 = b5[2]-c5 #distance of boundaries from center
  nb5 = sum(x>b5[1] & x<b5[2]) #number between
  no5 = sum(x<b5[1] | x>b5[2]) #number outside
  nb5l = sum(x>(b5[1]+1) & x<(b5[2]-1))
  nb5h = sum(x>(b5[1]-1) & x<(b5[2]+1))
  
  b6 = c(52.5,57.5)  #boundaries
  c6 = mean(b6) #center of boundaries
  d6 = b6[2]-c6 #distance of boundaries from center
  nb6 = sum(x>b6[1] & x<b6[2]) #number between
  no6 = sum(x<b6[1] | x>b6[2]) #number outside
  nb6l = sum(x>(b6[1]+1) & x<(b6[2]-1))
  nb6h = sum(x>(b6[1]-1) & x<(b6[2]+1))
  
  if(nb5 != nb5l & nb5 != nb5h & 2*d5==round(2*d5) & nb6 != nb6l & nb6 != nb6h & 2*d6==round(2*d6)){
    if(d5==round(d5)){
      wrdint5 = "an integer"
    } else {
      wrdint5 = "a [half-integer](https://en.wikipedia.org/wiki/Half-integer){target='blank'}"
    }
    if(d6==round(d6)){
      wrdint6 = "an integer"
    } else {
      wrdint6 = "a [half-integer](https://en.wikipedia.org/wiki/Half-integer){target='blank'}"
    }
    break
  }
}

if(d3 == 1){
  wrd3 = "unit"
} else {
  wrd3 = "units"
}

if(d4 == 1){
  wrd4 = "unit"
} else {
  wrd4 = "units"
}

i = 1:n
```


Question
========

A random procedure generated the following sample (sequence of measurements):

$$\mathbf{x} = `r x` $$

```{r echo=F}
fn = paste0("list_counting_between_data.csv")
write.csv(data.frame(i,x), file = fn, quote = FALSE, row.names = FALSE)
```

You can download the data as a CSV file (for importing into a spreadsheet or R): 
[`r fn`](`r fn`)

Answerlist
--------
* How many of the measurements are inside `r b1[1]` to `r b1[2]`? In other words, determine $\#[\mathbf{x}>`r b1[1]`\texttt{ AND } \mathbf{x}<`r b1[2]`]$. In other words, determine $\#[`r b1[1]`<\mathbf{x}<`r b1[2]`]$.
* How many of the measurements are outside `r b2[1]` to `r b2[2]`? In other words, determine $\#[\mathbf{x}<`r b2[1]`\texttt{ OR } \mathbf{x}>`r b2[2]`]$. In other words, determine $\#[\texttt{NOT}(`r b2[1]`<\mathbf{x}<`r b2[2]`)]$.
* How many of the measurements are closer than `r d3` `r wrd3` from `r c3`? In other words, determine $\#\left[\big|\mathbf{x}-`r c3`\big|<`r d3`\right]$.
* How many of the measurements are farther than `r d4` `r wrd4` from `r c4`? In other words, determine $\#\left[\big|\mathbf{x}-`r c4`\big|>`r d4`\right]$.
* Determine `r wrdint5` radius $r$ of an interval with center `r c5` such that $\#\left[\big|\mathbf{x}-`r c5`\big|<r\right] ~=~ `r nb5`$.
* Determine `r wrdint6` radius $r$ of an interval with center `r c6` such that $\#\left[\big|\mathbf{x}-`r c6`\big|>r\right] ~=~ `r no6`$.


Solution
========

First, it helps to sort the data.
$$\text{sort}(\mathbf{x}) = `r sort(x)`$$

### R

You could use R to sort the data:

```{r echo=FALSE,comment=NA}
str1 = paste('sort(c(',paste(x,collapse=","),'))',collapse="")
str2 = gsub(" ", "", str1, fixed = TRUE)
cat(str2)
```

```{r echo=FALSE}
sort(x)
```

[Video of a similar problem solved with R](https://vimeo.com/531555871){target="_blank"}


### Spreadsheet

You could also use a spreadsheet. Import the table into a spreadsheet. Then, highlight the column of measurements and use ``Sort`` function.

[Video of a similar problem solved with a  spreadsheet.](https://vimeo.com/531555892){target="_blank"}


Answerlist
--------
* (ANS=`r nb1`) There are `r nb1` measurements between `r b1[1]` and `r b1[2]`. $$\#[`r b1[1]`<\mathbf{x}<`r b1[2]`]=`r nb1`$$
* (ANS=`r no2`) There are `r no2` measurements outside `r b2[1]` and `r b2[2]`. $$\#[\mathbf{x}<`r b2[1]`\texttt{ OR } \mathbf{x}>`r b2[2]`]=`r no2`$$
* (ANS=`r nb3`) There are `r nb3` measurements closer than `r d3` units from `r c3`. It might help to rephrase the problem as finding how many measurements are between `r b3[1]` and `r b3[2]`. These boundaries are found by subtracting `r d3` from `r c3` and adding `r d3` to `r c3`. $$\#\left[\big|\mathbf{x}-`r c3`\big|<`r d3`\right]=`r nb3`$$
* (ANS=`r no4`) There are `r no4` measurements farther than `r d4` units from `r c4`. It might help to rephrase the problem as finding how many measurements are outside `r b4[1]` and `r b4[2]`. These boundaries are found by subtracting `r d4` from `r c4` and adding `r d4` to `r c4`. $$\#\left[\big|\mathbf{x}-`r c4`\big|>`r d4`\right]=`r no4`$$
* (ANS=`r d5`) You can guess and check until you find `r wrdint5` that satisfies the equation. It turns out that $r = `r d5`$ satisfies the equation, because $\#\left[\big|\mathbf{x}-`r c5`\big|<`r d5`\right] ~=~ `r nb5`$. Notice the boundaries are `r b5[1]` and `r b5[2]`, which are found by adding/subtracting `r d5` to/from `r c5`. $$\#\left[\big|\mathbf{x}-`r c5`\big|<`r d5`\right] ~=~ `r nb5`$$
* (ANS=`r d6`) You can guess and check until you find `r wrdint6` that satisfies the equation. It turns out that $r = `r d6`$ satisfies the equation, because $\#\left[\big|\mathbf{x}-`r c6`\big|>`r d6`\right] ~=~ `r no6`$. Notice the boundaries are `r b6[1]` and `r b6[2]`, which are found by adding/subtracting `r d6` to/from `r c6`. $$\#\left[\big|\mathbf{x}-`r c6`\big|>`r d6`\right] ~=~ `r no6`$$


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num
exsolution: `r nb1`|`r no2`|`r nb3`|`r no4`|`r d5`|`r d6`
exname: list_counting_between
extol: 0|0|0|0|0|0
exextra[numwidth,numeric]: 8
