```{r,data,generation,echo=FALSE,results="hide"}
# f = sample(1:9,1)*10
x = c(45.89, 49.26, 47.54, 40.1, 49.44, 40.9, 40.77, 42.22, 49.12, 49.57, 42.88, 42.82, 46.94, 48.71, 45.22, 45.77, 44.6, 48.23, 41.94, 42.24, 47.24, 45.37, 46.04, 42.72, 49.84, 48.57, 47.19, 42.48, 40.9, 49.73, 42.11, 40.1, 44.58, 48.82, 48.36, 45.14, 43.03, 42.48, 41.58, 42.3, 49.18, 46.6, 48.09, 44.63, 45.65, 41.65, 41.63, 42.87, 40.19, 47.84, 47.33, 44.02, 47.26, 47.77, 40.1, 42.25, 42.3, 45.01, 49.49, 49.6, 46.33, 45.12, 42.74, 47.69, 47.55, 44.99, 44.06, 43.66, 44.96, 42.52, 42.74, 49.06, 45.6, 41.58, 45.2, 43.9, 49.43, 49.15, 44.76, 42.4, 49.57, 40.9, 42.39, 46.53, 46.97, 47.07, 40.02, 40.44, 43.77, 46.02, 45.79, 44.64, 40.79, 45.34, 42.89, 49.6, 48.16, 40.14, 46.98, 41.1, 40.81, 40.92, 45.87, 43.75, 44.14, 44.69, 48.02, 41.56, 44.13, 49.23, 47.5, 46.75, 48.58, 49.58, 45.76, 41.04, 46.67, 45.08, 41.56, 40.07, 41.89, 49.06, 41.71, 49.18, 48.12, 46.63, 43.76, 45.3, 48.14, 44.66, 40.43, 47.38, 41.66, 48.09, 45.96, 46.9, 43.52, 45.35, 49.06, 42.69, 47.25, 44.29, 45.07, 49.35, 46.18, 45.18, 49.09, 40.41, 48.66, 44.62)
n = length(x)
i = 1:n

while(T){
  b2 = 40.6
  b3 = 47.5
  # b4 = round(sort(runif(2,min(x),max(x))),1)
  r4 = 1.5
  c4 = 43.3
  b4 = c(c4-r4,c4+r4)
  # b5 = round(sort(runif(2,min(x),max(x))),1)
  r5 = 1
  c5 = 42.2
  b5 = c(c5-r5,c5+r5)
  
  if(abs(b4[2]-b4[1])>1 & abs(b5[2]-b5[1])>1 & r4==round(r4,1) & r5==round(r5,1) & c4==round(c4,1) & c5==round(c5,1)){
    break
  }
}
n2 = sum(x<b2)
n3 = sum(x>b3)
n4 = sum(abs(x-c4)<r4)
n5 = sum(abs(x-c5)>r5)

if(r4 == 1){
  wrd4 = "unit"
} else {
  wrd4 = "units"
}

if(r5 == 1){
  wrd5 = "unit"
} else {
  wrd5 = "units"
}
```

```{r echo=F}
fn = paste0("list_counting_large_data.csv")
write.csv(data.frame(i,x), file = fn, quote = FALSE, row.names = FALSE)
```

Question
========

A random procedure generated many measurements: [`r fn`](`r fn`)

```{r echo=F, comment=NA, results="asis"}
cat(paste0(sapply(x,function(a){sprintf("%.2f",a)}),collapse=", "))
```

Answerlist
--------
* How many measurements are in the sample? In other words, determine $\#[\mathbf{x}]$. In the future, we will use $n$ to denote the number of measurements in the sample.
* How many of the measurements are less than `r b2`? In other words, determine $\#[\mathbf{x}<`r b2`]$.
* How many of the measurements are greater than `r b3`? In other words, determine $\#[\mathbf{x}>`r b3`]$.
* How many of the measurements are closer than `r r4` `r wrd4` from `r c4`? In other words, determine $\#\left[\big|\mathbf{x}-`r c4`\big|<`r r4`\right]$.
* How many of the measurements are farther than `r r5` `r wrd5` from `r c5`? In other words, determine $\#\left[\big|\mathbf{x}-`r c5`\big|>`r r5`\right]$.

Solution
========

You will want to use a computer to answer these questions. 

```{r echo=F,fig.width=12,fig.height=4}
mydf = data.frame(c(i,""),c(x,""))
mydf = cbind(mydf,c(paste0('=IF(B',i+1,'<',b2,',1,0)'),paste0('=SUM(C2:C',n+1,')')))
mydf = cbind(mydf,c(paste0('=IF(B',i+1,'>',b3,',1,0)'),paste0('=SUM(D2:D',n+1,')')))
mydf = cbind(mydf,c(paste0('=IF(ABS(B',i+1,'-',c4,')<',r4,',1,0)'),paste0('=SUM(E2:E',n+1,')')))
mydf = cbind(mydf,c(paste0('=IF(ABS(B',i+1,'-',c5,')>',r5,',1,0)'),paste0('=SUM(F2:F',n+1,')')))
names(mydf) = c("i","x",paste0("x<",b2),paste0("x>",b3),paste0("abs(x-",c4,")<",r4),paste0("abs(x-",c5,")>",r5))
fn2 = paste0("list_counting_large_solution.csv") 
write.csv(mydf, file = fn2, row.names = FALSE)
```
\


### Spreadsheet

[video](https://vimeo.com/531555960){target="_blank"}

You need to learn about references and formulas: [https://wiki.openoffice.org/wiki/Documentation/OOo3_User_Guides/Calc_Guide/Creating_formulas](https://wiki.openoffice.org/wiki/Documentation/OOo3_User_Guides/Calc_Guide/Creating_formulas){target="_blank"}

You need to learn about the `IF` function:
[https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_IF_function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_IF_function){target="_blank"}

You need to learn about the `SUM` function:
[https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_SUM_function](https://wiki.openoffice.org/wiki/Documentation/How_Tos/Calc:_SUM_function){target="_blank"}


If you used a spreadsheet, you should end up with [this solution csv](`r fn2`). You can also see the solution spreadsheet here (with the middle rows removed).


<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
    color: #88888888;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td:first-child{
    color: #888888;
}
</style>

```{r tab1, results="asis", echo=F}
mydf2 = mydf[c(1:4,(length(x)-2):(length(x)+1)),]
mydf2 = type.convert(mydf2, as.is = TRUE)
mydf2 = rbind(names(mydf2),mydf2)
mydf2 = cbind(c(1:5,(length(x)-1):(length(x)+2)),mydf2)
colnames(mydf2) = c(" ",LETTERS[1:6])
mydf2[length(mydf2[,1]),2] = " "
mydf2[length(mydf2[,1]),3] = " "
# mydf2 = cbind(c(1:4,(length(x)-2):(length(x)+1)),mydf2)
knitr::kable(mydf2,format="html",table.attr='class="bordered-table"', align = "ccccccc",row.names=F)
```

If you have the solution spreadsheet open in a spreadsheet program, you can display the functions by simultaneously pressing `ctrl` and `~` and then return to the outputs by pressing `ctrl` and `~` again. If you see ###, you can expand the cell-width until there is enough room to display the functions.



### R

[video](https://vimeo.com/531555935){target="_blank"}

To use R, the following commands would answer the questions.
```{r echo=F, comment=NA}
cat(paste0("x = c(",paste0(x,collapse=","),")"))
```
```{r}
length(x)
```

```{r echo=F, comment=NA}
cat(paste0("sum(x<",b2,")"))
```
```{r echo=F}
sum(x<b2)
```

```{r echo=F, comment=NA}
cat(paste0("sum(x>",b3,")"))
```
```{r echo=F}
sum(x>b3)
```
```{r echo=F, comment=NA}
cat(paste0("sum(abs(x-",c4,")<",r4,")"))
```
```{r echo=F}
sum(abs(x-c4)<r4)
```
```{r echo=F, comment=NA}
cat(paste0("sum(abs(x-",c5,")>",r5,")"))
```

```{r echo=F}
sum(abs(x-c5)>r5)
```

Notice that the standard output in R starts with "##" and then [1]. For now, just ignore these. If the [1] really bothers you, you could use the [`cat`](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cat){target="_blank"} function.



Answerlist
--------
* (ANS=`r n`) There are `r n` measurements.
* (ANS=`r n2`) There are `r n2` measurements less than `r b2`.
* (ANS=`r n3`) There are `r n3` measurements greater than `r b3`.
* (ANS=`r n4`) There are `r n4` measurements closer than `r r4` units from `r c4`. You could also say there are `r n4` measurements inside `r b4[1]` to `r b4[2]`.
* (ANS=`r n5`) There are `r n5` measurements farther than `r r5` units from `r c5`. You could also say there are `r n5` measurements outside `r b5[1]` to `r b5[2]`.


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num
exsolution: `r n`|`r n2`|`r n3`|`r n4`|`r n5`
exname: list_counting_large
extol: 0
exextra[numwidth,numeric]: 8
