```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
n = 100
i = 1:n
type = 2 #sample(1:2,1)
ns = c(6,8)[type]
val = 1:8
fre = c(12,9,10,14,16,10,18,11)
x = rep(val,fre) #sample(1:ns,n,T)
brks = 1:(ns+1)-0.5
myhist = hist(x,breaks=brks)
counts = myhist$counts
strdie = c("six","eight")[type]

b1 = 2.5 #sample(brks[3:(length(brks)-1)],1)
n1 = sum(x<b1)
b2 = 3.5 #sample(brks[2:(length(brks)-2)],1)
n2 = sum(x>b2)
b3 = c(3.5,6.5) #sort(sample(brks[2:(length(brks)-1)],2))
c3 = mean(b3)
r3 = (b3[2]-b3[1])/2
n3 = sum(abs(x-c3)<r3)
b4 = c(2.5,4.5) #sort(sample(brks[2:(length(brks)-1)],2))
c4 = mean(b4)
r4 = (b4[2]-b4[1])/2
n4 = sum(abs(x-c4)>r4)

b5 = 7.5 #sample(brks[3:(length(brks)-1)],1)
n5 = sum(x<b5)
b6 = 4.5 #sample(brks[2:(length(brks)-2)],1)
n6 = sum(x>b6)
b7 = c(5.5,7.5) #sort(sample(brks[2:(length(brks)-1)],2))
c7 = mean(b7)
r7 = (b7[2]-b7[1])/2
if(r7==round(r7)){
  wrd7 = "integer"
} else {
  wrd7 = "half-integer"
}
n7 = sum(abs(x-c7)<r7)
b8 = c(2.5,6.5) #sort(sample(brks[2:(length(brks)-1)],2))
c8 = mean(b8)
r8 = (b8[2]-b8[1])/2
if(r8==round(r8)){
  wrd8 = "integer"
} else {
  wrd8 = "half-integer"
}
n8 = sum(abs(x-c8)>r8)
```


Question
========

A standard `r strdie`-sided [die](https://en.wikipedia.org/wiki/Dice) was rolled `r n` times, and the results were organized into the [pie chart](https://en.wikipedia.org/wiki/Pie_chart){target="blank"} shown below.

```{r echo=F,fig.width=4,fig.height=4}
size = 1.2
ps = counts/100
cp = cumsum(c(0,ps,1))
mids = (cp[2:(ns+1)]+cp[1:ns])/2
par(mar=c(0,0,0,0))
plot(0,0,type="n",xlim=c(-size,size),ylim=c(-size,size))
th = seq(0,2*pi,length.out=1000)
lines(sin(th),cos(th))
for(a in cp){
  lines(c(0,sin(a*2*pi)),c(0,cos(a*2*pi)))
}
for(a in seq(0.1,0.9,0.1)){
  lines(c(sin(a*2*pi),1.13*sin(a*2*pi)),c(cos(a*2*pi),1.13*cos(a*2*pi)))
  text(1.19*sin(a*2*pi),1.19*cos(a*2*pi),a*100)
}
for(a in seq(0.05,0.95,0.1)){
  lines(c(sin(a*2*pi),1.10*sin(a*2*pi)),c(cos(a*2*pi),1.10*cos(a*2*pi)))
  text(1.16*sin(a*2*pi),1.16*cos(a*2*pi),a*100)
}
for(a in seq(0.01,0.99,0.01)){
  lines(c(sin(a*2*pi),1.04*sin(a*2*pi)),c(cos(a*2*pi),1.04*cos(a*2*pi)))
}
for(i in 1:ns){
  a = mids[i]
  text(0.7*sin(a*2*pi),0.7*cos(a*2*pi),(1:n)[i],cex=2)
}
text(0.04,1.15,0,srt=90)
text(-0.04,1.15,100,srt=90)
lines(c(0,0),c(1,size),lwd=3)
```
\


Answerlist
--------
* Evaluate $\#[\mathbf{x}<`r b1`]$.
* Evaluate $\#[\mathbf{x}>`r b2`]$.
* Evaluate $\#[|\mathbf{x}-`r c3`|<`r r3`]$.
* Evaluate $\#[|\mathbf{x}-`r c4`|>`r r4`]$.
* Evaluate half-integer boundary $b$ such that $\#[\mathbf{x}<b] = `r n5`$.
* Evaluate half-integer boundary $b$ such that $\#[\mathbf{x}>b] = `r n6`$.
* Evaluate `r wrd7` radius $r$ such that $\#[|\mathbf{x}-`r c7`|<r] = `r n7`$.
* Evaluate `r wrd8` radius $r$ such that $\#[|\mathbf{x}-`r c8`|>r] = `r n8`$.


Solution
========

[video](https://vimeo.com/531899870){target="blank"}

The first 4 questions involve adding up frequencies of the indicated intervals. The last 4 questions can be done by guessing and checking until something works.



Answerlist
--------
* (ANS=`r n1`) There are `r n1` measurements less than `r b1`. So, $\#[\mathbf{x}<`r b1`]=`r n1`$.
* (ANS=`r n2`) There are `r n2` measurements more than `r b2`. So, $\#[\mathbf{x}>`r b2`]=`r n2`$.
* (ANS=`r n3`) There are `r n3` measurements closer than `r r3` units from `r c3`. So, $\#[|\mathbf{x}-`r c3`|<`r r3`]=`r n3`$.
* (ANS=`r n4`) There are `r n4` measurements further than `r r4` units from `r c4`. So, $\#[|\mathbf{x}-`r c4`|>`r r4`]=`r n4`$.
* (ANS=`r b5`) A boundary at $b=`r b5`$ has `r n5` measurements less than it. So, $\#[\mathbf{x}<`r b5`]=`r n5`$.
* (ANS=`r b6`) A boundary at $b=`r b6`$ has `r n6` measurements more than it. So, $\#[\mathbf{x}>`r b6`]=`r n6`$.
* (ANS=`r r7`) An interval with radius $r=`r r7`$ around center `r c7` contains `r n7` measurements. So, $\#[|\mathbf{x}-`r c7`|<`r r7`]=`r n7`$.
* (ANS=`r r8`) An interval with radius $r=`r r8`$ around center `r c8` excludes `r n8` measurements.  So, $\#[|\mathbf{x}-`r c8`|>`r r8`]=`r n8`$.


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num
exsolution: `r n1`|`r n2`|`r n3`|`r n4`|`r b5`|`r b6`|`r r7`|`r r8`
exname: read_circle_freq
extol: 0
exextra[numwidth]: 999
