```{r data generation, echo = FALSE, results = "hide"}
library(knitr)
opts_chunk$set(echo = FALSE)
options(scipen=100)

gamma = sample(c(0.8,0.9,0.95,0.95,0.96,0.98,0.99),1)
gamma2=gamma
n = round(runif(1,7,20)^2)
p = runif(1,0.20,0.80)
x = (runif(n)<p)*1
x2 = x
phat = mean(x)
zstar = round(qnorm(0.5+gamma/2),2)
phatr = round(phat,3)
zstarr = round(zstar,2)
LB = phat-zstar*sqrt(phat*(1-phat)/n)
UB = phat+zstar*sqrt(phat*(1-phat)/n)
LBr = round(LB,3)
UBr = round(UB,3)

write.csv(data.frame(x),"freethrows.csv",row.names=F)

default_source_hook <- knit_hooks$get('source')
knit_hooks$set(source = function(x, options) {
  x <- gsub('gamma2',paste0(gamma),x)
  x <- gsub('x2',paste0(x2,collapse=", "),x)
  default_source_hook(x, options)
})
```

Question
========

A basketball player has decided to estimate her probability to score a [freethrow](https://en.wikipedia.org/wiki/Free_throw){target="blank"}. To do this, she shoots freethrows. If she scores, she records a "1". If she misses, she records a "0".

The results are shown below and can be [downloaded as a csv](freethrows.csv){target="blank"}.

```{r, comment=NA}
cat(paste0(x,collapse=", "))
```

You are asked to determine a [confidence interval](https://en.wikipedia.org/wiki/Binomial_proportion_confidence_interval){target="blank"} using a given confidence level: $\gamma=`r gamma`$. To do this, you [determine](https://en.wikipedia.org/wiki/Confidence_interval#Basic_steps){target="blank"} $z^\star$ such that $P(|Z|<z^\star)=`r gamma`$. You also calculate the sample size and sample proportion. Then, you use the following formulas:
$$\text{LB} = \hat{p}-z^\star \sqrt{\frac{\hat{p}(1-\hat{p})}{n}} $$
$$\text{UB} = \hat{p}+z^\star \sqrt{\frac{\hat{p}(1-\hat{p})}{n}} $$


Answerlist
--------
* Determine the lower boundary of the confidence interval
* Determine the upper boundary of the confidence interval



Solution
========

You should be able to determine that $P(|Z|<`r zstar`)=`r gamma`$, so $z^\star=`r zstar`$. Using a spreadsheet or R, you should determine that $n=`r n`$ and $\hat{p} = \frac{`r sum(x)`}{`r n`} = `r round(phat,3)`$. Then, use the given formulas.

### R

```{r, echo=T}
gamma = gamma2
zstar = qnorm(0.5+gamma/2)
x = c(x2)
n = length(x)
phat = mean(x)
LB = phat-zstar*sqrt(phat*(1-phat)/n)
UB = phat+zstar*sqrt(phat*(1-phat)/n)
print(data.frame(LB,UB,row.names=""))
```

### Spreadsheet

```{r}
spacer = rep("",n)
variables = c("","gamma","cumulative","zstar","","n","phat","LB","UB",rep("",n-9))
formulas = c("",paste(gamma),"=0.5+D2/2",
             "=NORM.INV(D3,0,1)","",
             "=COUNT(A2:A1000)",
             "=SUM(A2:A1000)/D6",
             "=D7-D4*SQRT(D7*(1-D7)/D6)",
             "=D7+D4*SQRT(D7*(1-D7)/D6)",
             rep("",n-9))
sol = data.frame(x,spacer,variables,formulas)
colnames(sol)=c("x","","variable","formula")

write.csv(sol,"freethrowsol.csv",row.names=F)
```

You can [download a solution spreadsheet](freethrowsol.csv){target="blank"}.

The first 10 lines are shown here:

```{r, fig.width=6,fig.height=4}
draw_ss = function(df,widths=rep(1,length(df[1,])+1)){
  xline = c(0,cumsum(widths)/sum(widths))
  nrow = length(df[,1])
  ncol = length(df[1,])
  xtex = (xline[2:(ncol+2)]+xline[1:(ncol+1)])/2
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  #xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=xline)
  abline(v=xline[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}

draw_ss(sol[1:9,],c(1,1,0.3,2,4))
```
\

Answerlist
--------
* The lower boundary: 
$$\begin{align}
\text{LB} &= \hat{p}-z^\star\sqrt{\frac{\hat{p}(1-\hat{p})}{n}} \\
&= (`r phatr`)-(`r zstarr`)\sqrt{\frac{(`r phatr`)(1-`r phatr`)}{`r n`}}\\
&= `r LBr`
\end{align}
$$
* The upper boundary: 
$$\begin{align}
\text{UB} &= \hat{p}+z^\star\sqrt{\frac{\hat{p}(1-\hat{p})}{n}} \\
&= (`r phatr`)+(`r zstarr`)\sqrt{\frac{(`r phatr`)(1-`r phatr`)}{`r n`}}\\
&= `r UBr`
\end{align}
$$

Meta-information
================
extype: cloze
exclozetype: num|num
exsolution: `r LB`|`r UB`
exname: confidence interval prop raw
extol: 0.01
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE