```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)
n = sample((8:20)^2,1)
la = runif(1,0.01,0.99)
mag = sample(-1:2,1)
N = sample(c(4,6,8,10,12,20),1)
mu = (N+1)/2
sig = sqrt((N^2-1)/12)
mut = mu*n
sigt = sig*sqrt(n)
zapp = qnorm(la)
xt = ceiling(mut+zapp*sigt)-0.5
z = (xt-mut)/sigt
zlo = floor(z*100)/100
zhi = ceiling(z*100)/100
la = pnorm(z)
lalo = pnorm(zlo)
lahi = pnorm(zhi)

tol = max(abs(lahi-lalo)*1.1,0.002)

```

Question
========

A fair `r N`-sided die (with sides numbered 1 through `r N`) will be rolled `r n` times, and the sum (total) of those rolls will be recorded. What is the probability that the **sum** is less than `r xt`?
$$P\left(\sum X < `r xt`\right) ~=~ ? $$

To help you along, I will calculate the mean and standard deviation of **single** rolls. The formulas for an $N$-sided die can be derived from the formulas for [discrete uniform distribution](https://en.wikipedia.org/wiki/Discrete_uniform_distribution).
$$\mu_{ X} ~=~ \frac{N+1}{2} ~=~ \frac{`r N`+1}{2} ~=~ `r mu` $$
$$\sigma_{ X} ~=~ \sqrt{\frac{N^2-1}{12}} ~=~ \sqrt{\frac{`r N`^2-1}{12}} ~=~ `r sig`$$

Please use a normal approximation based on the central limit theorem. 


Solution
========

[video](https://vimeo.com/531941885){target="_blank"}

We calculate the mean and standard deviation of the **random sum** using the formulas from the central limit theorem.
$$\begin{aligned}
\mu_{ \sum X} &= n\mu_{ X} \\
&= (`r n`)(`r mu`) \\
&= `r mut`
\end{aligned}$$

$$\begin{aligned}
\sigma_{ \sum X} &= \sigma_{ X} \sqrt{n} \\
&= (`r sig`)\sqrt{`r n`} \\
&= `r sigt`
\end{aligned}$$

The central limit theorem tells us that the random sum is approximately normal with the parameters calculated above.

$$\sum X \sim N(`r mut`,\,`r sigt`) $$

In other words, we can approximate the summing of `r n` rolls of `r N`-sided dice with a single spin of the following spinner.


```{r, fig.width=8, fig.height=8}
library(latex2exp)
par(mar=c(0,0,0,0))
plot(0,0,type="n",xlim=c(-1.2,1.2),ylim=c(-1.2,1.2),ann=F)
theta = seq(0,2*pi,0.01)
polygon(cos(theta),sin(theta))
z1 = round(seq(-3,3,1),3)
z2 = round(seq(-3,3,0.5),2)
z3 = round(seq(-3,3,0.1),2)
z4 = round(seq(-2,2,0.05),2)
z5 = round(seq(-2,2,0.01),2)
markz = function(z,L,lwd=1){
  p = pnorm(z)
  for(pp in p){
    lines(c((1-L)*cos(2*pi*pp),cos(2*pi*pp)),c((1-L)*sin(2*pi*pp),sin(2*pi*pp)),lwd=lwd)
  }
}
markz(z1,0.3,3)
markz(z2,0.18)
markz(z3,0.1)
markz(z4,0.05)
markz(z5,0.025)
for(z0 in z1){
  p = pnorm(z0)
  pc = min(max(p,0.005),0.995)
  text((1-0.37)*cos(2*pi*pc),(1-0.37)*sin(2*pi*pc),round(signif(mut+sigt*qnorm(p),4),6),cex=0.9)
}
for(z0 in z2[!(z2 %in% z1)]){
  p = pnorm(z0)
  pc = min(max(p,0.005),0.995)
  text((1-0.26)*cos(2*pi*pc),(1-0.26)*sin(2*pi*pc),signif(mut+sigt*qnorm(p),4),cex=0.9)
}
for(z0 in z3[!(z3 %in% z2) & abs(z3)<2]){
  p = pnorm(z0)
  pc = min(max(p,0.005),0.995)
  text((1-0.16)*cos(2*pi*pc),(1-0.16)*sin(2*pi*pc),signif(mut+sigt*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.9)
}
for(z0 in z4[!(z4 %in% z3) & abs(z4)<1.5]){
  p = pnorm(z0)
  pc = min(max(p,0.005),0.995)
  text((1-0.1)*cos(2*pi*pc),(1-0.1)*sin(2*pi*pc),signif(mut+sigt*qnorm(p),4),srt=(90+p*360)%%180-90,cex=0.6)
}
p1 = round(seq(0,1,0.1),2)
p2 = round(seq(0,1,0.05),2)
p3 = round(seq(0,1,0.01),2)
lines(c(1,1.2),c(0,0),lwd=3)
for(p in p1){
  lines(c(cos(p*2*pi),1.15*cos(p*2*pi)),c(sin(p*2*pi),1.15*sin(p*2*pi)))
  pc = max(min(p,0.996),0.004)
  text(1.2*cos(pc*2*pi),1.2*sin(pc*2*pi),round(p,2),cex=1)
}
for(p in p2[!(p2 %in% p1)]){
  lines(c(cos(p*2*pi),1.1*cos(p*2*pi)),c(sin(p*2*pi),1.1*sin(p*2*pi)))
  pc = max(min(p,0.996),0.004)
  text(1.16*cos(pc*2*pi),1.16*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}
for(p in p3[!(p3 %in% p2)]){
  lines(c(cos(p*2*pi),1.05*cos(p*2*pi)),c(sin(p*2*pi),1.05*sin(p*2*pi)))
  pc = max(min(p,0.996),0.004)
  text(1.12*cos(pc*2*pi),1.12*sin(pc*2*pi),round(p,2),srt=(90+p*360)%%180-90,cex=0.8)
}

text(1,1,"U\nStandard Uniform\nDistribution")
text(0,0.2,TeX("$\\sum X$"))
text(0,0,sprintf("A Normal Distribution\nwith\nmean = %.4g \n and \nstdev = %.4g",mut,sigt))
```
\

Find the appropriate $z$ score.

$$z ~=~ \frac{(\sum x)-\mu_{ \sum X}}{\sigma_{ \sum X}} ~=~ \frac{`r xt`-`r mut`}{`r sigt`} ~=~ `r z`$$

You can round this to either `r zlo` or `r zhi`. I will continue with the unrounded $z$ score. We now rephrase the question as a standard normal probability.

$$P\left(\sum X < `r xt`\right) ~=~ P(Z<`r z`) $$
$$P(Z<`r z`) ~=~ \, ?$$

You can use the $z$ table to find the probability.

$$P(Z<`r z`) ~=~ `r la`$$

$$P(Z<`r zlo`) ~=~ `r round(lalo,4)`$$
$$P(Z<`r zhi`) ~=~ `r round(lahi,4)`$$

We can sketch the density curve and shade the appropriate region.

```{r, fig.width=6, fig.height=3}
par(mar=c(7,3,0,0))
library(latex2exp)
Z = seq(-3,3,0.001)
plot(Z,dnorm(Z),type="l",xlim=c(-3,3),ylim=c(0,0.4),xlab="Z",ylab="density",axes=F,ann=F,lwd=3)
axis(1,at=round(seq(-3,3,1),1),line=0)
mtext("Z axis",1,0,at=-3.5)
axis(1,at=round(seq(-3,3,1),1),labels=signif(seq(mut-3*sigt,mut+3*sigt,length.out=7),3),line=3)
mtext(TeX("$\\sum X$ axis"),1,3,at=-3.5)
abline(h=0,lwd=3)
abline(h=seq(0,0.4,0.05),lwd=0.5)
abline(v=seq(-3,3,0.2),lwd=0.5)

polygon(c(Z[Z<=z],z,-3),c(dnorm(Z[Z<=z]),0,0),col=rgb(1,0.5,0.5,0.5))

text(-2.01,0.32,TeX(sprintf("$P(%s<%s) = %.4f$","\\sum X",xt,round(pnorm(z),4))))
arrows(-1.6,0.3,(2*z-3)/3,max(dnorm((2*z-3)/3)/2,0.03),lwd=2)
```
\


Meta-information
================
extype: num
exsolution: `r la`
exname: prob tot less
extol: `r tol`
