```{r,data,generation,echo=FALSE,results="hide"}
n = 10 #must be bigger than 8, or change xcom to replacement sampling with T
f = sample(1:9,1)*10
x = sort(sample(0:9,n,TRUE))+f
xcom = sample(x,8,FALSE)
x1 = xcom[1]
x2 = xcom[2]
x3 = xcom[3]
x4 = xcom[4]
i1 = sum(x<xcom[1])
i2 = sum(x<=xcom[2])
i3 = sum(x>xcom[3])
i4 = sum(x>=xcom[4])
p1 = i1/n
p2 = i2/n
p3 = i3/n
p4 = i4/n
x5 = xcom[5]
x6 = xcom[6]
x7 = xcom[7]
x8 = xcom[8]
i5 = sum(x<xcom[5])
i6 = sum(x<=xcom[6])
i7 = sum(x>xcom[7])
i8 = sum(x>=xcom[8])
p5 = i5/n
p6 = i6/n
p7 = i7/n
p8 = i8/n
xrand = sample(x)
x = xrand
i = 1:n
```


Question
========

A [random procedure](https://en.wikipedia.org/wiki/Random_number_generation){target="blank"} generated the following sample (sequence of measurements):

$$\mathbf{x} = `r x` $$

```{r echo=F}
fn = paste0("list_proportion_basic_data.csv")
write.csv(data.frame(i,x), file = fn, quote = FALSE, row.names = FALSE)
```

You can download the data as a CSV file (for importing into a spreadsheet or R): 
[`r fn`](`r fn`)

Answerlist
--------
* What proportion of the measurements are less than `r x1`? In other words, determine $\text{prop}[\mathbf{x}<`r x1`]$.
* What proportion of the measurements are less than or equal to `r x2`? In other words, determine $\text{prop}[\mathbf{x}\le`r x2`]$.
* What proportion of the measurements are greater than `r x3`? In other words, determine $\text{prop}[\mathbf{x}>`r x3`]$.
* What proportion of the measurements are greater than or equal to `r x4`? In other words, determine $\text{prop}[\mathbf{x}\ge`r x4`]$.
* If the proportion of measurements less than $x_{\text{c}}$ is `r p5`, and $x_{\text{c}}$ is a measurement from sample $\mathbf{x}$, then what is $x_{\text{c}}$? In other words, determine a measurement $x_{\text{c}}$ from sequence $\mathbf{x}$ such that $\text{prop}[\mathbf{x}<x_{\text{c}}] = `r p5`$.
* If the proportion of measurements less than or equal to $x_{\text{c}}$ is `r p6`, and $x_{\text{c}}$ is a measurement from sample $\mathbf{x}$, then what is $x_{\text{c}}$? In other words, determine a measurement $x_{\text{c}}$ from sequence $\mathbf{x}$ such that $\text{prop}[\mathbf{x}\le x_{\text{c}}] = `r p6`$.
* If the proportion of measurements greater than $x_{\text{c}}$ is `r p7`, and $x_{\text{c}}$ is a measurement from sample $\mathbf{x}$, then what is $x_{\text{c}}$? In other words, determine a measurement $x_{\text{c}}$ from sequence $\mathbf{x}$ such that $\text{prop}[\mathbf{x}>x_{\text{c}}] = `r p7`$.
* If the proportion of measurements greater than or equal to $x_{\text{c}}$ is `r p8`, and $x_{\text{c}}$ is a measurement from sample $\mathbf{x}$, then what is $x_{\text{c}}$? In other words, determine a measurement $x_{\text{c}}$ from sequence $\mathbf{x}$ such that $\text{prop}[\mathbf{x}\ge x_{\text{c}}] = `r p8`$.


Solution
========

First, it helps to sort the data.
$$`r sort(x)`$$

You could use R to sort the data:

```{r echo=FALSE,comment=NA}
str1 = paste('sort(c(',paste(x,collapse=","),'))',collapse="")
str2 = gsub(" ", "", str1, fixed = TRUE)
cat(str2)
```

```{r echo=FALSE}
sort(x)
```

You could also use a spreadsheet. Import the table into a spreadsheet. Then, highlight the column of measurements and use ``Sort`` function.

Answerlist
--------
* (ANS=`r p1`) The proportion of measurements less than `r x1` is `r p1`. $$\text{prop}[\mathbf{x}<`r x1`]=`r p1`$$
* (ANS=`r p2`) The proportion of measurements less than or equal to `r x2` is `r p2`. $$\text{prop}[\mathbf{x}\le `r x2`]=`r p2`$$
* (ANS=`r p3`) The proportion of measurements greater than `r x3` is `r p3`. $$\text{prop}[\mathbf{x}> `r x3`]=`r p3`$$
* (ANS=`r p4`) The proportion of measurements greater than or equal to `r x4` is `r p4`. $$\text{prop}[\mathbf{x}\ge `r x4`]=`r p4`$$
* (ANS=`r x5`) The proportion of measurements less than `r x5` is `r p5`. $$\text{prop}[\mathbf{x}<`r x5`]=`r p5`$$
* (ANS=`r x6`) The proportion of measurements less than or equal to `r x6` is `r p6`. $$\text{prop}[\mathbf{x}\le `r x6`]=`r p6`$$
* (ANS=`r x7`) The proportion of measurements greater than `r x7` is `r p7`. $$\text{prop}[\mathbf{x}> `r x7`]=`r p7`$$
* (ANS=`r x8`) The proportion of measurements greater than or equal to `r x8` is `r p8`. $$\text{prop}[\mathbf{x}\ge `r x8`]=`r p8`$$


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num
exsolution: `r p1`|`r p2`|`r p3`|`r p4`|`r x5`|`r x6`|`r x7`|`r x8`
exname: list_proportion_basic
extol: 0
exextra[numwidth,character]: 999.999