```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
n = sample(c(20,50,100,200,500,1000),1)
i = 1:n
type = sample(1:2,1)
ns = c(6,8)[type]
x = sample(1:ns,n,T)
brks = 1:(ns+1)-0.5
myhist = hist(x,breaks=brks)
counts = myhist$counts
strdie = c("six","eight")[type]

b1 = sample(brks[3:(length(brks)-1)],1)
n1 = sum(x<b1)
b2 = sample(brks[2:(length(brks)-2)],1)
n2 = sum(x>b2)
b3 = sort(sample(brks[2:(length(brks)-1)],2))
c3 = mean(b3)
r3 = (b3[2]-b3[1])/2
n3 = sum(abs(x-c3)<r3)
b4 = sort(sample(brks[2:(length(brks)-1)],2))
c4 = mean(b4)
r4 = (b4[2]-b4[1])/2
n4 = sum(abs(x-c4)>r4)

b5 = sample(brks[3:(length(brks)-1)],1)
n5 = sum(x<b5)
b6 = sample(brks[2:(length(brks)-2)],1)
n6 = sum(x>b6)
b7 = sort(sample(brks[2:(length(brks)-1)],2))
c7 = mean(b7)
r7 = (b7[2]-b7[1])/2
if(r7==round(r7)){
  wrd7 = "integer"
} else {
  wrd7 = "half-integer"
}
n7 = sum(abs(x-c7)<r7)
b8 = sort(sample(brks[2:(length(brks)-1)],2))
c8 = mean(b8)
r8 = (b8[2]-b8[1])/2
if(r8==round(r8)){
  wrd8 = "integer"
} else {
  wrd8 = "half-integer"
}
n8 = sum(abs(x-c8)>r8)
p1 = n1/n
p2 = n2/n
p3=n3/n
p4=n4/n
p5=n5/n
p6=n6/n
p7=n7/n
p8=n8/n
```


Question
========

A standard `r strdie`-sided [die](https://en.wikipedia.org/wiki/Dice) was rolled `r n` times, and the results were organized into the [histogram](https://en.wikipedia.org/wiki/Histogram){target="blank"} shown below. 
<!-- In [dice notation](https://en.wikipedia.org/wiki/Dice_notation), we could say the results of `r n`d`r ns` were plotted as a histogram. (Pedantic sidenote: it is common to interpret `r n`d`r ns` as the SUM of `r n` rolls, but we will interpret `r n`d`r ns` as the LIST of `r n` rolls and use $\sum `r n`\text{d}`r ns`$ as the sum of `r n` rolls.) -->

```{r echo=F, fig.width=6, fig.height=4}
ylim = max(counts/n)*1.1
hist(x,breaks=brks,labels = TRUE,ylim=c(0,ylim),xlim=c(0,ns+1),freq=F,ylab="Relative Frequency",xaxt='n')
axis(1,1:ns)
```
\


Answerlist
--------
* Evaluate $\text{prop}[\mathbf{x}<`r b1`]$.
* Evaluate $\text{prop}[\mathbf{x}>`r b2`]$.
* Evaluate $\text{prop}[|\mathbf{x}-`r c3`|<`r r3`]$.
* Evaluate $\text{prop}[|\mathbf{x}-`r c4`|>`r r4`]$.
* Evaluate half-integer boundary $b$ such that $\text{prop}[\mathbf{x}<b] = `r n5/n`$.
* Evaluate half-integer boundary $b$ such that $\text{prop}[\mathbf{x}>b] = `r n6/n`$.
* Evaluate `r wrd7` radius $r$ such that $\text{prop}[|\mathbf{x}-`r c7`|<r] = `r n7/n`$.
* Evaluate `r wrd8` radius $r$ such that $\text{prop}[|\mathbf{x}-`r c8`|>r] = `r n8/n`$.


Solution
========

The first 4 questions involve adding up relative frequencies of the indicated intervals. The last 4 questions can be done by guessing and checking until something works.

You may also find it is helpful to determine frequencies (counts). You do this by multiplying each relative frequency by the total number of measurements.

```{r echo=F, fig.width=6, fig.height=4}
ylim = max(counts)*1.1
hist(x,breaks=brks,labels = TRUE,ylim=c(0,ylim),xlim=c(0,ns+1),xaxt='n')
axis(1,1:ns)
```
\




Answerlist
--------
* There are `r n1` measurements less than `r b1`. So, $\text{prop}[\mathbf{x}<`r b1`]=\frac{`r n1`}{`r n`}=`r n1/n`$.
* There are `r n2` measurements more than `r b2`. So, $\text{prop}[\mathbf{x}>`r b2`]=\frac{`r n2`}{`r n`}=`r n2/n`$.
* There are `r n3` measurements closer than `r r3` units from `r c3`. So, $\text{prop}[|\mathbf{x}-`r c3`|<`r r3`]=\frac{`r n3`}{`r n`}=`r n3/n`$.
* There are `r n4` measurements further than `r r4` units from `r c4`. So, $\text{prop}[|\mathbf{x}-`r c4`|>`r r4`]=\frac{`r n4`}{`r n`}=`r n4/n`$.
* First, convert the proportion into a count. $`r p5`\cdot`r n`=`r n5`$. A boundary at $b=`r b5`$ has `r n5` measurements less than it. So, $\text{prop}[\mathbf{x}<`r b5`]=`r n5/n`$.
* First, convert the proportion into a count. $`r p6`\cdot`r n`=`r n6`$. A boundary at $b=`r b6`$ has `r n6` measurements more than it. So, $\text{prop}[\mathbf{x}>`r b6`]=`r n6/n`$.
* An interval with radius $r=`r r7`$ around center `r c7` contains `r n7` measurements. So, $\text{prop}[|\mathbf{x}-`r c7`|<`r r7`]=`r n7/n`$.
* An interval with radius $r=`r r8`$ around center `r c8` excludes `r n8` measurements.  So, $\text{prop}[|\mathbf{x}-`r c8`|>`r r8`]=`r n8/n`$.


Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num
exsolution: `r n1/n`|`r n2/n`|`r n3/n`|`r n4/n`|`r b5`|`r b6`|`r r7`|`r r8`
exname: read_histogram_discrete_prop
extol: 0.001
exextra[numwidth]: 999.99
