---
title: "Intervals, counting"
author: "Chad Worley"
date: "9/15/2021"
output:
  slidy_presentation:
    css: basic.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Class discussion

- To simplify class, I want to focus on a single tool during lecture.
  - spreadsheets or Rstudio
- I believe it will be easier for everyone if we pick Rstudio.

## [Intervals](https://en.wikipedia.org/wiki/Interval_(mathematics)){target="_blank"} (review)

- A half-bounded interval has 1 boundary.
  - $x<10$ <p>
```{r open1,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(-55,55,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in seq(-50,50,10)){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(-55,10),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=8)
points(10,0,cex=3,pch=20,col=rgb(1,1,1))
points(10,0,cex=2.5,col=rgb(1,0.2,0.2),lwd=5)
```  
  - $x\ge20$ <p>
```{r greater,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(-55,55,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in seq(-50,50,10)){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(20,55),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=8)
points(20,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
```
- A bounded interval has 2 boundaries.
  - $-25\le x ~~~\mathtt{AND}~~~ x<40$ <p>
```{r boundedinterval,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(-55,55,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in seq(-50,50,10)){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(-25,40),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=8)
points(-25,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
points(40,0,cex=3,pch=20,col=rgb(1,1,1))
points(40,0,cex=2.5,col=rgb(1,0.2,0.2),lwd=5)
```
- In stats, we are also interested in two-tail "intervals".
  - $x<15 ~~~\mathtt{OR}~~~ x>35$ <p>
```{r twotail,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(-55,55,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in seq(-50,50,10)){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(-55,15),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=8)
lines(c(35,55),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=8)
points(15,0,cex=3,pch=20,col=rgb(1,1,1))
points(15,0,cex=2.5,col=rgb(1,0.2,0.2),lwd=5)
points(35,0,cex=3,pch=20,col=rgb(1,1,1))
points(35,0,cex=2.5,col=rgb(1,0.2,0.2),lwd=5)
```

## Bounded intervals have a middle (center) and radius

- In Statistics, we care about distance from center (like in archery, accuracy, error).
- $3.5\le x ~~~\mathtt{AND}~~~ x\le 8.5$ <p>
```{r boundedinterval2,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(0.5,10.5,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in 1:10){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(3.5,8.5),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=10)
points(3.5,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
points(8.5,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
arrows(6,-0.6,6,-0.1,col=rgb(0.2,0.2,0.9),lwd=6)
text(6,-0.8,"middle",col=rgb(0.2,0.2,0.9),cex=1.5)
lines(c(6,6,6,8.5,8.5,8.5),c(0.9,0.7,0.8,0.8,0.7,0.9),lwd=6,col=rgb(0,0.5,0.5))
text((6+8.5)/2,0.95,"radius",col=rgb(0,0.5,0.5),cex=1.5)
```
- Let's say the boundaries are $b_1$ and $b_2$, where $b_2>b_1$.
$$m = \frac{b_1+b_2}{2}$$
$$r=\frac{b_2-b_1}{2}$$
- So, in this example...
$$m = \frac{3.5+8.5}{2} = \frac{12}{2} = 6$$
$$r = \frac{8.5-3.5}{2} = \frac{5}{2} = 2.5$$

## It is helpful to define intervals using middle and radius.

- $|x-6|\le 2.5$ <p>
```{r ttint,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(0.5,10.5,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in 1:10){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(3.5,8.5),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=10)
points(3.5,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
points(8.5,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
arrows(6,-0.6,6,-0.1,col=rgb(0.2,0.2,0.9),lwd=6)
text(6,-0.8,"middle",col=rgb(0.2,0.2,0.9),cex=1.5)
lines(c(6,6,6,8.5,8.5,8.5),c(0.9,0.7,0.8,0.8,0.7,0.9),lwd=6,col=rgb(0,0.5,0.5))
text((6+8.5)/2,0.95,"radius",col=rgb(0,0.5,0.5),cex=1.5)
```
- $|x-6|\ge 2.5$ <p>
```{r boundedinterval3,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
x = seq(0.5,10.5,1)
plot(x,rep(0,length(x)),ann=F,axes=F,type="l",lwd=3,ylim=c(-1,1))
for(x in 1:10){
  lines(c(x,x),c(-0.3,0.3))
  text(x,0.5,x)
}
lines(c(0.5,3.5),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=10)
lines(c(8.5,10.5),c(0,0),col=rgb(1,0.2,0.2,0.6),lwd=10)
points(3.5,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
points(8.5,0,cex=4,col=rgb(1,0.2,0.2),pch=20)
arrows(6,-0.6,6,-0.1,col=rgb(0.2,0.2,0.9),lwd=6)
text(6,-0.8,"middle",col=rgb(0.2,0.2,0.9),cex=1.5)
lines(c(6,6,6,8.5,8.5,8.5),c(0.9,0.7,0.8,0.8,0.7,0.9),lwd=6,col=rgb(0,0.5,0.5))
text((6+8.5)/2,0.95,"radius",col=rgb(0,0.5,0.5),cex=1.5)
```

```{r absolutediffs,echo=T}
x = 1:10
ineq1 = abs(x-6)<=2.5
ineq2 = abs(x-6)>=2.5
data.frame(x,ineq1,ineq2)
```

## If you know the middle and radius, you can calculate the bounds

- Doubly bounded interval: $$|x-m|<r$$
- Two-tailed interval: $$|x-m|>r$$
- Absolute difference is 1-dimensional distance.
$$|x-m|=\text{distance between }x\text{ and }m $$
- If you know the middle and radius, you can determine the boundaries.
$$b_{1} = m-r $$
$$b_{2} = m+r $$

## Counting {.incremental}

<style>
.tiny{
  font-size:60%;
}
</style>

- An archer shoots 72 arrows each day. Each arrow scores between 0 and 10 points, depending on its location. The archer records her total score each day.
```{r}
# set.seed(1)
# x = round(rnorm(50,72*8.5,17))
x = c(601, 615, 598, 639, 618, 598, 620, 625, 622, 607, 638, 619, 601, 574, 631, 611, 612, 628, 626, 622, 628, 625, 613, 578, 623, 611, 609, 587, 604, 619, 635, 610, 619, 611, 589, 605, 605, 611, 631, 625, 609, 608, 624, 621, 600, 600, 618, 625, 610, 627)
write.csv(x,file="data.csv")
```
<div class="tiny">
- `r x`
</div>
- The [data can be downloaded](data.csv){target="_blank"}.

```{r,fig.width=8,fig.height=2}
par(mar=c(0,0,0,0))
a = seq(560,660,10)
plot(a,rep(0,length(a)),ann=F,axes=F,type="l",lwd=3,ylim=c(-0.6,0.6))
for(a in 560:660){
  if(a%%10==0){
    lines(c(a,a),c(-0.1,0),lwd=3)
    text(a,-0.2,a)
  } else if(a%%5==0){
    lines(c(a,a),c(-0.07,0),lwd=2)
    # text(a,0.4,a,cex=0.8)
  } else {
    lines(c(a,a),c(-0.03,0))
  }
}
mytab = table(x)
for(i in 1:length(mytab)){
  name = names(mytab)[i]
  freq = as.vector(mytab)[i]
  val = as.integer(name)
  for(i in 1:freq){
    points(val,(i-1)*0.05+0.03,pch=20,cex=1.3,col=rgb(0,0,0))
    points(val,(i-1)*0.05+0.03,pch=20,cex=0.8,col=rgb(0.8,0,0))
  }
}
```

- Determine $n$, the sample size.

```{r,echo=T}
x = c(601, 615, 598, 639, 618, 598, 620, 625, 622, 607, 638, 619, 601, 574, 631, 611, 612, 628, 626, 622, 628, 625, 613, 578, 623, 611, 609, 587, 604, 619, 635, 610, 619, 611, 589, 605, 605, 611, 631, 625, 609, 608, 624, 621, 600, 600, 618, 625, 610, 627)
n = length(x)
print(n)
```
- Determine $\#(x=600)$
```{r,echo=T}
sum(x==600)
```
- Determine $\#(x<600)$
```{r,echo=T}
sum(x<600)
```
- Determine $\#(x>620)$
```{r,echo=T}
sum(x>620)
```
- Determine $\#(|x-610|<10)$
```{r,echo=T}
sum(abs(x-610)<10)
```
- Determine $\#(|x-610|>10)$
```{r,echo=T}
sum(abs(x-610)>10)
```
