```{r data generation, echo = FALSE, results = "hide"}
while(T){
  n = sample(8:25,1)
  x = sample(3:6,1)
  ans = factorial(n)/factorial(n-x)/factorial(x)
  if(ans < 1e8){
    break
  }
}

q = paste("A pizza shop has ",n," different toppings available. You will choose ",x," different toppings for your pizza. How many possibilities exist?",collapse="")

sss = ''
sss = paste(sss,n)
for(i in (n-1):(n-x+1)){
  sss = paste(sss,i,sep="\\cdot")
}

ss2 = ''
ss2 = paste(ss2,x)
for(i in (x-1):(1)){
  ss2 = paste(ss2,i,sep="\\cdot")
}

options(scipen = 100)

```

Question
========

`r q`

Solution
========

This scenario describes a combinations problem (order of selection does not matter). We are considering the subsets of size `r x` from a set of size `r n`.

$$\begin{aligned}
{_nC_r} &= \frac{n!}{(n-r)! \cdot r!} \\\\
n &= `r n` \\\\
r &= `r x` \\\\
{_{`r n`}C_{`r x`}} &= \frac{`r n`!}{(`r n`-`r x`)!\cdot `r x`! } \\\\
&= \frac{`r n`!}{`r n-x`! \cdot `r x`!} \\\\
&= \frac{`r sss`}{`r ss2`} \\\\
&= `r ans`
\end{aligned}$$

Remember, we care about combinations because they represent all the ways we can select $r$ 1s and $n-r$ 0s. So, in this case, ${_{`r n`}C_{`r x`}} = `r ans`$ tells us there are `r ans` ways of selecting `r n-x` 0s and `r x` 1s. (Think of the 1s as the toppings that are selected and 0s as the toppings NOT selected.)

```{r, echo=F}
samples <- vector(mode = "list", length = 10)
ss <- vector(mode = "list", length = 10)
for(i in 1:10){
  while(T){
    sam = sort(sample(1:n,x))
    if(!(all(sam %in% samples))){
      s = ''
      samples[[i]] = list(sam)
      for(j in 1:n){
        if(j %in% sam){
          s = paste(s,'1',collapse='')
        } else {
          s = paste(s,'0',collapse='')
        }
      }
      ss[[i]] = s
      break
    }
  }
}
```

So, if you had a lot of time, you could list out all possibilities:

| Count | Possibility |
|:----:|:-----:|
| 1 | `r ss[[1]]` |
| 2 | `r ss[[2]]` |
| 3 | `r ss[[3]]` |
| 4 | `r ss[[4]]` |
| 5 | `r ss[[5]]` |
| $\vdots$ | $\vdots$ |
| `r ans-4` | `r ss[[6]]` |
| `r ans-3` | `r ss[[7]]` |
| `r ans-2` | `r ss[[8]]` |
| `r ans-1` | `r ss[[9]]` |
| `r ans` | `r ss[[10]]` |

Of course, you'd want to be more systematic than that...


Meta-information
============
extype: num
exsolution: `r ans`
exname: combinations
extol: 0.01
