```{r data generation, echo = FALSE, results = "hide"}
ell = runif(1,0.03,0.97)
z = round(qnorm(ell),2)
ell = round(pnorm(z),4)
```

Question
========
Determine the probability that the standard normal variable is less than `r z`. In other words, evaluate $P(Z < `r z`)$.


Solution
========

[video](https://vimeo.com/531931017){target="_blank"}

The numbers that satisfy $Z<`r z`$ are on the left side of a number line (toward $-\infty$). The probability equals a left area under the density curve.

```{r, echo=F, fig.height=4, fig.width=6}
par(mar=c(7,0,0,0))
zs = seq(-3,3,0.01)
plot(zs,dnorm(zs),xlim=c(-3,3),type='l',ylab='density',xlab='Z',axes=F,ann=F)
polygon(c(zs[zs<=z],z,-3),c(dnorm(zs[zs<=z]),0,0),col=rgb(0.7,0.7,0.7,0.7))
lines(c(-3,3),c(0,0))
axis(1,at=c(-3,z,3),line = 0)
axis(1,at=seq(-3,3,1),line=2)
mtext('Z',1,5)
text(mean(c(-3,z)),max(dnorm(mean(c(-3,z)))/2,0.02),"?")
```
\


By using the z-table, we find the appropriate probability.
$$P(Z<`r z`)=`r ell`$$

It might help to visualize with a spinner:

```{r echo=F,fig.width=6,fig.height=6}
par(mar=c(0,0,0,0))
theta=seq(0,2*pi,length.out=1000)
plot(sin(theta),cos(theta),type="l",ann=F,axes=F,xlim=c(-1.3,1.3),ylim=c(-1.3,1.3))
zs = -3:3
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.8,1),cos(p*2*pi)*c(0.8,1),lwd=2)
  pt = min(max(p,0.007),0.993)
  if(abs(zs[i])<3){
    text(0.75*sin(pt*2*pi),0.75*cos(pt*2*pi),zs[i])
  }
}
zs = seq(-2.5,2.5,1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.85,1),cos(p*2*pi)*c(0.85,1),lwd=2)
  if(abs(zs[i])<3){
    text(sin(p*2*pi)*0.8,cos(p*2*pi)*0.8,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.9,1),cos(p*2*pi)*c(0.9,1),lwd=1)
}
zs = seq(-2,2,0.1)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  if(zs[i]%%0.5 != 0){
    text(sin(p*2*pi)*0.85,cos(p*2*pi)*0.85,cex=0.5,round(zs[i],2),srt=(180-360*p)%%180-90)
  }
}
zs = seq(-3,3,0.05)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.95,1),cos(p*2*pi)*c(0.95,1))
}
zs = seq(-3,3,0.01)
for(i in 1:length(zs)){
  p = pnorm(zs[i])
  lines(sin(p*2*pi)*c(0.98,1),cos(p*2*pi)*c(0.98,1))
}

for(p in seq(0.1,0.9,0.1)){
  lines(sin(p*2*pi)*c(1,1.05),cos(p*2*pi)*c(1,1.05))
  text(sin(p*2*pi)*1.1,cos(p*2*pi)*1.1,p,cex=0.8,srt=-(p*360-90)%%180-90)
}
for(p in seq(0.01,0.99,0.01)){
  lines(sin(p*2*pi)*c(1,1.03),cos(p*2*pi)*c(1,1.03))
}
for(p in seq(0.002,0.998,0.002)){
  lines(sin(p*2*pi)*c(1,1.01),cos(p*2*pi)*c(1,1.01))
}
lines(c(0,0),c(1,1.05),lwd=2)
text(0,1.09,"1 or 0",cex=0.8)
text(0,0,cex=1.5,"z")

pp = pnorm(z)
ps = seq(0,pp,0.001)
polygon(c(0,sin(ps*2*pi),0),c(0,cos(ps*2*pi),0),col=rgb(0.9,0.3,0.3,0.4),border=rgb(0.9,0.3,0.3))
pe = c(0,pp)
pm = mean(pe)

lines(sin(ps*2*pi)*1.2,cos(ps*2*pi)*1.2,lwd=2)
lines(sin(pe[1]*2*pi)*c(1.15,1.2),cos(pe[1]*2*pi)*c(1.15,1.2),lwd=2)
lines(sin(pe[2]*2*pi)*c(1.15,1.2),cos(pe[2]*2*pi)*c(1.15,1.2),lwd=2)
text(0.5*sin(2*pi*pe[1]),0.5*cos(2*pi*pe[1]),"negative infinity",srt=(180-360*pe[1])%%180-90)
text(0.5*sin(2*pi*pe[2]),0.5*cos(2*pi*pe[2]),z,srt=(180-360*pe[2])%%180-90)
text(1.25*sin(2*pi*pe[1]),1.25*cos(2*pi*pe[1]),round(pe[1],4),srt=-(pe[1]*360-90)%%180-90,cex=0.8)
text(1.25*sin(2*pi*pe[2]),1.25*cos(2*pi*pe[2]),round(pe[2],4),srt=-(pe[2]*360-90)%%180-90,cex=0.8)
text(sin(pm*2*pi)*1.25,
     cos(pm*2*pi)*1.25,
     sprintf("%.4f",round(pe[2],4)-round(pe[1],4)),
     srt=-(pm*360-90)%%180-90)
```
\


Using a spreadsheet:
```
=NORM.DIST(`r z`,0,1,TRUE)
```

Using R:
```
pnorm(`r z`)
```


Meta-information
================
extype: num
exsolution: `r ell`
exname: Get leftarea from Z
extol: 0.002
