```{r,data,generation,echo=FALSE,results="hide",fig.show='hide'}
while(T){
  type = sample(1:5,1) #uniform, bell, bimodal, skewright, skewleft
  n = sample(c(60,75,80,90,100,110,120,125,130,150),1)
  beta1 = c(1,10,0.2,0.5,2)[type]
  beta2 = c(1,10,0.2,2,0.5)[type]
  bottom = sample(seq(20,70,5),1)
  width = sample(seq(5,30,5),1)
  i = 1:n
  x = rbeta(n,beta1,beta2)*width+bottom
  myhist = hist(x,breaks=5)
  brks = myhist$breaks
  counts = myhist$counts
  relf = counts/n
  dens = relf/(brks[2]-brks[1])
  if(brks[2]-brks[1] != 1){
    break
  }
}

sol = signif(c(rbind(relf,dens)),6)
soln = paste0(sol,collapse="|")
tol = paste0(sol/10,collapse="|")

```


Question
========

```{r echo=F}
write.csv(data.frame(i,x),file="make_freq_dist.csv",row.names = FALSE)
```

A random procedure generated $n=`r n`$ measurements: [download data](make_freq_dist.csv)

I've already determined the frequencies. Please determine the relative frequencies and the densities. A brief description of relative frequency and density can be found [here](https://en.wikipedia.org/wiki/Histogram){target="blank"}.

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>

```{r anstab, results="asis", echo=F}
NN = length(counts)
Interval = paste(brks[1:NN],"to",brks[2:(NN+1)])
Frequency = counts
Relative_frequency = paste("##ANSWER",seq(1,NN*2-1,2),"##",sep="")
Density = paste("##ANSWER",seq(2,2*NN,2),"##",sep="")
anstab = data.frame(Interval,Frequency,Relative_frequency,Density)
knitr::kable(anstab,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```


Solution
========

[video](https://vimeo.com/531903568){target="_blank"}

To determine the relative frequencies, just divide each frequency by `r n` (because $n=`r n`$). To determine the densities, divide the relative frequencies by the width of the interval, which in this case is the same for each interval ($`r brks[2]`-`r brks[1]`=`r brks[2]-brks[1]`$).

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>


```{r soltab, results="asis", echo=F}
NN = length(counts)
Interval = paste(brks[1:NN],"to",brks[2:(NN+1)])
Frequency = counts
Relative_frequency = signif(counts/n,4)
Density = signif(counts/n/(brks[2]-brks[1]),4)
soltab = data.frame(Interval,Frequency,Relative_frequency,Density)
knitr::kable(soltab,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```




Meta-information
================
extype: cloze
exclozetype: `r paste0(rep("num",length(counts)*2),collapse="|")`
exsolution: `r soln`
exname: relfreq_and_dens
extol: `r tol`
exextra[numwidth,character]: 9999
