```{r data generation, echo = FALSE, results = "hide"}
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)

repeat{
  n1 = sample(40:80,1)
  p1 = runif(1,0.2,0.8)
  n2 = sample(40:80,1)
  p2 = runif(1,0.2,0.8)
  x1 = (runif(n1)<p1)*1
  x2 = (runif(n2)<p2)*1
  ns1 = sum(x1)
  ns2 = sum(x2)
  nf1 = n1-ns1
  nf2 = n1-ns1
  fttab = matrix(c(ns1,nf1,ns2,nf2),nrow=2)
  pval1 = prop.test(c(ns1,ns2),c(n1,n2),correct=T)$p.value
  pval2 = prop.test(c(ns1,ns2),c(n1,n2),correct=F)$p.value
  
  pval3 = fisher.test(matrix(c(ns1,n1-ns1,ns2,n2-ns2),
                             nrow=2))$p.value
  
  if( (pval1<0.05)==(pval2<0.05) & (pval1<0.05)==(pval3<0.05) &
      abs(pval1-0.05)>0.01 &
      abs(pval2-0.05)>0.01 &
      abs(pval3-0.05)>0.01 &
      pval1<0.2 & pval1>0.005 & 
      pval2<0.2 & pval2>0.005 & 
      pval3<0.2 & pval3>0.005){break}
}

phat1 = ns1/n1
phat2 = ns2/n2
phat = (ns1+ns2)/(n1+n2)

z = abs(phat2-phat1)/sqrt(phat*(1-phat)*(1/n1+1/n2))

ans = ":NUMERICAL:=%s:0.005~%%100%%%s:0.005~%%100%%%s:0.005"
ans = sprintf(ans,pval3,pval2,pval1)


if(pval1<0.05){
  ans2="10"
  s2="The $p$-value is less than 0.05, so the result is significant."
} else {
  ans2="01"
  s2="The $p$-value is more than 0.05, so the result is NOT significant."
}

x1rep = x1
x2rep = x2
default_source_hook <- knitr::knit_hooks$get('source')
knitr::knit_hooks$set(source = function(xx, options) {
  xx <- gsub('x1rep',paste(x1,collapse=","),xx)
  xx <- gsub('x2rep',paste(x2,collapse=","),xx)
  default_source_hook(xx, options)
})
```

Question
========

A doctor runs a [controlled experiment](https://en.wikipedia.org/wiki/Randomized_controlled_trial){target="blank"}. The sick patients are randomly assigned to two groups: control and treatment. The patients in the control group are given a [placebo](https://en.wikipedia.org/wiki/Placebo-controlled_study){target="blank"}. The patients in the treatment group are given a drug.

```{r}
len = max(n1,n2)
if(n1>n2){
  x1d = x1
  x2d = c(x2,rep(NA,n1-n2))
} else {
  x2d = x2
  x1d = c(x1,rep(NA,n2-n1))
}
data = data.frame(x1d,x2d)
colnames(data) = c("x1","x2")
write.csv(data,file="recovery.csv",row.names=F,na = "")
```

After a month, each patient was checked for whether they recovered from the sickness. A "0" means no recovery while a "1" means recovery. This data can also be [downloaded as a csv](recovery.csv){target="blank"}.

```{r}
cat(
sprintf("Control: %s\n\nTreatment: %s",
        paste0(x1,collapse = ", "),
        paste0(x2,collapse = ", ")
        )
)
```

You are asked to perform a two-tailed [two-proportion test](https://en.wikipedia.org/wiki/Test_statistic#Common_test_statistics){target="blank"}. You can do a two-proportions $z$ test (which is equivalent to a 2x2 $\chi^2$ test [[chi squared test](https://en.wikipedia.org/wiki/Chi-squared_test){target="blank"}]). You will get credit whether or not you apply a continuity correction, but please pool data for the standard error estimation.

Just for completeness, you can also get credit for using [Fisher's exact test](https://en.wikipedia.org/wiki/Fisher%27s_exact_test){target="blank"}.

##ANSWER1##

Is the difference of means significant (using a significance level of 0.05)?

##ANSWER2##

Answerlist
--------
* Determine the $p$-value.
* Yes, the drug causes a difference in recovery.
* No, we don't know whether the drug causes a difference.

Solution
========

There are many ways to do this problem. I will show the following:

* "By hand"
* Use a spreadsheet's `chisq.test` function on 2x2 [contingency table](https://en.wikipedia.org/wiki/Contingency_table){target="blank"}
* Use `R`'s `prop.test` and `fisher.test` functions.

### "By hand"

First, somehow you need to determine the sample sizes ($n_1$ and $n_2$) and the sample totals (numbers of recovery), $\sum x_1$ and $\sum x_2$. You could count... but a computer is probably helpful.

$$n_1 = `r n1`$$
$$n_2 = `r n2`$$
$$\sum x_1 = `r ns1`$$
$$\sum x_2 = `r ns2`$$

It can be helpful to organize these summary statistics into a contingency table.

<style type="text/css" rel="stylesheet">
.bordered-table {
    border-collapse: collapse;
    border-spacing: 0;
    border:2px solid #000000;
}
.bordered-table th {
    border:2px solid #000000;
    padding: 4px;
}
.bordered-table td{
    border:2px solid #000000;
    padding: 4px;
}
</style>

```{r tab1, results="asis"}
group = c("Control","Treatment","TOTAL")
Recover = c(ns1,ns2,ns1+ns2)
Not_recover = c(nf1,nf2,nf1+nf2)
TOTAL = c(n1,n2,n1+n2)
ct = data.frame(group,Recover,Not_recover,TOTAL)
colnames(ct) = c("","Recover","Not_recover","TOTAL")
knitr::kable(ct,format="html",table.attr='class="bordered-table"', align = "cccc",row.names=F)
```

Calculate the proportions.
$$\hat{p}_1 = \frac{`r ns1`}{`r n1`} = `r phat1` $$
$$\hat{p}_2 = \frac{`r ns2`}{`r n2`} = `r phat2` $$
$$\hat{p} = \frac{`r ns1`+`r ns2`}{`r n1`+`r n2`} = `r phat` $$

Determine the absolute $z$ score.

$$|z| = \frac{\left|\hat{p}_2-\hat{p}_1\right|}{\sqrt{\hat{p}(1-\hat{p})\left(\frac{1}{n_1}+\frac{1}{n_2}\right)}} =`r z`$$

Using a $z$ table or an online standard-normal probability tool, determine the appropriate probabilities.

$$P(Z<`r round(z,2)`) = `r pnorm(z)`$$
$$P(|Z|>`r round(z,2)`) = (2)(1-`r round(pnorm(z),4)`) = `r 2*(1-pnorm(z))` $$
$$p\text{-value} = `r pval2` $$
This $p$-value did not use the continuity correction.

`r s2`


### Spreadsheet

```{r ss}
spacer = rep("",length(x1d))
D = c("OBSERVED","","control","treatment",
      "TOTAL","","phat pooled =","",
      "EXPECTED","","control","treatment",
      "TOTAL","","p-value =")
E = c("","recover","=SUM(A2:A1000)","=SUM(B2:B1000)",
      "=E4+E5","","=E6/G6","","","recover",
      "=E8*G12","=E8*G13","=E8*G14","","=CHISQ.TEST(E4:F5,E12:F13)")
FF = c("","not_recover","=G4-E4","=G5-E5","=F4+F5","",
       "","","","not_recover","=G12-E12","=G13-E13",
       "=G14-E14","","")
G = c("","TOTAL","=COUNT(A2:A1000)","=COUNT(B2:B1000)",
      "=G4+G5","","","","","TOTAL","=G4","=G5","=G6","","")
D = c(D,rep("",length(x1d)-length(D)))
E = c(E,rep("",length(x1d)-length(E)))
FF = c(FF,rep("",length(x1d)-length(FF)))
G = c(G,rep("",length(x1d)-length(G)))

sol = data.frame(x1d,x2d,spacer,D,E,FF,G)
colnames(sol) = c("x1","x2","","","","","")
write.csv(sol,"recovery_solution.csv",row.names=F)
```

Solution [download](recovery_solution.csv){target="blank"}.

```{r fig.height=8,fig.width=10}
draw_ss = function(df,widths=rep(1,length(df[1,])+1)){
  xline = c(0,cumsum(widths)/sum(widths))
  nrow = length(df[,1])
  ncol = length(df[1,])
  xtex = (xline[2:(ncol+2)]+xline[1:(ncol+1)])/2
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+2)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  #xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=xline)
  abline(v=xline[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
    text(xtex[i],ytex[2],colnames(df)[i-1])
  }    
  for(i in 2:(nrow+2)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+2],df[i,j])
    }
  }
}
draw_ss(sol[1:16,],c(1,1,1,0.5,2,4,2,3))
```
\

If you wanted to make Yates' continuity correction, you would need to calculate the EXPECTED table the same, but then make another table, where the observed values are all 0.5 closer to the expected values. Then, you would use `=CHISQ.TEST` on these new values (as the "observed") and the expected values.

It looks challenging to do a Fisher exact test in a spreadsheet.

### R

```{r echo=T}
x1 = c(x1rep)
x2 = c(x2rep)
n1 = length(x1)
n2 = length(x2)
ns1 = sum(x1) #number of successes in sample 1
ns2 = sum(x2)
prop.test(c(ns1,ns2),c(n1,n2))
prop.test(c(ns1,ns2),c(n1,n2),correct=F)
nf1 = n1-ns1
nf2 = n2-ns2
conttab = matrix(c(ns1,nf1,ns2,nf2),nrow=2)
fisher.test(conttab)
```


Meta-information
================
extype: cloze
exsolution: `r ans`|`r ans2`
exclozetype: verbatim|schoice
exname: hypothesis test 2 proportions
extol: 0.005|0
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE
