```{r data generation, echo = FALSE, results = "hide"}
library(knitr)
opts_chunk$set(echo = FALSE)
options(scipen=100)

while(T){
    p0 = round(runif(1,0.2,0.8),2)
    n = round(runif(1,6,12)^2)
    sigma = sqrt(p0*(1-p0))
    se = sigma/sqrt(n)
    gamma = 0.95
    zstar = qnorm(0.5+gamma/2)
    me = zstar*se
    p = p0+sample(c(-1,1),1)*me
    x = (runif(n)<p)*1
    ns = sum(x)
    zt1 = prop.test(ns,n,p0,correct=F)
    zt2 = prop.test(ns,n,p0,correct=T)
    bt = binom.test(ns,n,p0)
    pval1 = zt1$p.value
    pval2 = zt2$p.value
    pval3 = bt$p.value
    if(pval1<0.30 & pval1>0.005 & abs(pval1-0.05)>0.01 
       & abs(pval2-0.05)>0.005 & abs(pval3-0.05)>0.005
       & (pval1<0.05)==(pval2<0.05) &  (pval1<0.05)==(pval3<0.05)){break}
}

phat = mean(x)
phatr = round(phat,4)

z1 = (phat-p0)/sqrt(p0*(1-p0)/n)
pval4 = 2*(1-pnorm(abs(z1)))

CI1 = zt1$conf.int
CI2 = zt2$conf.int
CI3 = bt$conf.int
CI4 = c(phat-1.96*sqrt(phat*(1-phat)/n),
        phat+1.96*sqrt(phat*(1-phat)/n))

ans1 = ":NUMERICAL:=%s:0.005~%%100%%%s:0.005~%%100%%%s:0.005~%%100%%%s:0.005"
ans1 = sprintf(ans1,CI1[1],CI2[1],CI3[1],CI4[1])
ans2 = ":NUMERICAL:=%s:0.005~%%100%%%s:0.005~%%100%%%s:0.005"
ans2 = sprintf(ans2,CI1[2],CI2[2],CI3[2],CI4[2])
ans3 = ":NUMERICAL:=%s:0.005~%%100%%%s:0.005~%%100%%%s:0.005~%%100%%%s:0.005"
ans3 = sprintf(ans3,pval1,pval2,pval3,pval4)

write.csv(data.frame(x),"data_one_prop.csv",row.names=F)

if(pval1<0.05){
  ans4="10"
  s4="The p-value is less than 0.05, so the difference is significant."
} else {
  ans4="01"
  s4="The p-value is more than 0.05, so the difference is NOT significant."
}

xrep = x
default_source_hook <- knit_hooks$get('source')
knit_hooks$set(source = function(y, options) {
  y <- gsub('xrep',paste0(x,collapse=", "),y)
  default_source_hook(y, options)
})
```

Question
========

A scratch-off lottery has a stated chance of `r p0` to win. You sampled some tickets, marking a win as "1" and a loss as "0".


```{r datadump,comment=NA}
rlen = 20
xst = lapply(x,function(y){sprintf("%d",y)})
nfilledrows = length(xst)%/%rlen
nlast = length(xst)%%rlen
s = ""
if(nfilledrows>0){
  for(i in 1:nfilledrows){
    s = paste0(s,paste0(xst[((i-1)*rlen+1):(i*rlen)],collapse=", "),",\n")
  }
}
if(nlast>0){
  s = paste0(s,
             paste0(xst[(nfilledrows*rlen+1):length(xst)],collapse=", "))
}
cat(s)
```

[Download data.](data_one_prop.csv){target="blank"}

Please determine a 95% confidence interval, calculate an appropriate $p$-value, and state whether the sample proportion is significantly different than the stated chance (using a significance level of 0.05).

For the confidence interval, you can use a normal approximation interval, a Wilson score interval, a Wilson score interval with continuity correction, or the exact Clopper-Pearson interval ([see descriptions here](https://en.wikipedia.org/wiki/Binomial_proportion_confidence_interval){target="blank"}).

For the $p$-value, you can use similarly use a $z$ test, a $\chi^2$ test (with or without continuity correction), or an exact test.

##ANSWER1##

##ANSWER2##

##ANSWER3##

Using a significance level of 0.05, is the sample proportion significantly different from the stated chance?

##ANSWER4##

Answerlist
--------
* Determine the lower boundary of the confidence interval.
* Determine the upper boundary of the confidence interval.
* Determine the $p$-value.
* Yes
* No


Solution
========


By hand, this is easiest to do with a normal approximation without the continuity correction. 

$$n = `r n`$$
$$\hat{p} = `r phatr`$$
$$z^\star = 1.96 $$
$$\text{CI}_\text{boundaries} = \hat{p} \pm z^\star\sqrt{\frac{\hat{p}(1-\hat{p})}{n}} $$
$$\text{CI}_\text{boundaries} = (`r CI4[1]`,`r CI4[2]`)$$
$$z = \frac{\hat{p}-p_0}{\sqrt{p_0(1-p_0)/n}} = `r z1`$$

$$p\text{-value} = P\left(|Z|>`r abs(z1)`\right) = `r pval1`$$

Answerlist
--------
* Normal approx: $`r CI4[1]`$. Wilson: $`r CI1[1]`$. Wilson with CC: $`r CI2[1]`$. Exact Clopper-Pearson: $`r CI3[1]`$.
* Normal approx: $`r CI4[2]`$. Wilson: $`r CI1[2]`$. Wilson with CC: $`r CI2[2]`$. Exact Clopper-Pearson: $`r CI3[2]`$.
* Normal approx: $`r pval4`$. Chi squared: $`r pval1`$. Chi squared with CC: $`r pval2`$. Exact: $`r pval3`$.
* `r s4`
* 



Meta-information
================
extype: cloze
exsolution: `r ans1`|`r ans2`|`r ans3`|`r ans4`
exclozetype: verbatim|verbatim|verbatim|schoice
exname: one prop
extol: 0.005|0.005|0.005|0
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE