```{r data generation, echo = FALSE, results = "hide"}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
options(scipen=100)

xbar = round(runif(1,10,100),1)
s = round(runif(1,xbar/6,xbar/3),1)
n = sample(5:50,1)
gamma = sample(c(0.8,0.9,0.95,0.96,0.98,0.99),1)
tstar = qt(gamma+(1-gamma)/2,n-1)
ME = tstar*s/sqrt(n)
LB = round(xbar-tstar*s/sqrt(n),1)
UB = round(xbar+tstar*s/sqrt(n),1)
tol = max(0.02*s/sqrt(n),0.1)

xbar2 = xbar
s2 = s
n2 = n
gamma2 = gamma
default_source_hook <- knit_hooks$get('source')
knit_hooks$set(source = function(x, options) {
  x <- gsub('xbar2',paste0(xbar),x)
  x <- gsub('s2',paste0(s),x)
  x <- gsub('n2',paste0(n),x)
  x <- gsub('gamma2',paste0(gamma),x)
  default_source_hook(x, options)
})
```

Question
========

A population's mean ($\mu$) and standard deviation ($\sigma$) are unknown, but the population is approximately normal. A sample of size $n=`r n`$ is taken, and the sample mean is calculated: $\bar{x}=`r xbar`$. The sample standard deviation is also calculated: $s=`r s`$. You are tasked with determining a confidence interval using a given confidence level: $\gamma=`r gamma`$.

To do this, you need to first determine $t^\star$ such that $P(|T|<t^\star)=`r gamma`$. Then, the boundaries are determined by evaluating $\bar{x}\pm \frac{t^\star s}{\sqrt{n}}$.

(You can round $t^\star$ to the nearest hundredth and the boundaries to the nearest tenth.)

Answerlist
--------
* Determine the lower boundary of the confidence interval by using $\bar{x}-\frac{t^\star s}{\sqrt{n}}$.
* Determine the upper boundary of the confidence interval by using $\bar{x}+\frac{t^\star s}{\sqrt{n}}$.


Solution
========

[video](https://vimeo.com/531949124){target="_blank"}

First, determine $t^\star$ such that $P(|T|<t^\star)=`r gamma`$. It can help to draw a sketch.

```{r middle,fig.width=5,fig.height=3}
library(latex2exp)
tmax = ceiling(qt(0.998,n-1))
ts = seq(-tmax,tmax,0.001)
par(mar=c(2,0,0,0))
plot(ts,dt(ts,n-1),type="l",lwd=2,axes=F,ann=F)
abline(h=0)
polygon(c(-tstar,ts[abs(ts)<=tstar],tstar),
        c(0,dt(ts[abs(ts)<=tstar],n-1),0),
        col=rgb(1,0.5,0.5,0.5))
axis(1,at=c(-tmax,-tstar,0,tstar,tmax),labels=c(paste(-tmax),TeX("$$-t^{*}$$"),"0",TeX("$$t^{*}$$"),paste(tmax)))
text(0,0.15,gamma,cex=2,col=rgb(0.5,0.2,0.2,1))
```
\

The entire area under the density curve is 1. Thus, we can determine the areas of the tails (using symmetry).
$$\text{area of each tail} = \frac{1-\gamma}{2} = \frac{1-`r gamma`}{2} = `r (1-gamma)/2`$$

```{r allthree,fig.width=5,fig.height=3}
library(latex2exp)
tmax = ceiling(qt(0.998,n-1))
ts = seq(-tmax,tmax,0.001)
par(mar=c(2,0,0,0))
plot(ts,dt(ts,n-1),type="l",lwd=2,axes=F,ann=F)
abline(h=0)
polygon(c(-tstar,ts[abs(ts)<=tstar],tstar),
        c(0,dt(ts[abs(ts)<=tstar],n-1),0),
        col=rgb(1,0.5,0.5,0.5))
axis(1,at=c(-tmax,-tstar,0,tstar,tmax),labels=c(paste(-tmax),TeX("$$-t^{*}$$"),"0",TeX("$$t^{*}$$"),paste(tmax)))
text(0,0.15,gamma,cex=2,col=rgb(0.5,0.2,0.2,1))
polygon(c(-tmax,ts[ts<=-tstar],-tstar),
        c(0,dt(ts[ts<=-tstar],n-1),0),
        col=rgb(0.5,1,0.5,0.5))
text(-tmax+0.8,0.15,(1-gamma)/2,cex=2,col=rgb(0.2,0.5,0.2,1))
polygon(c(tstar,ts[ts>=tstar],tmax),
        c(0,dt(ts[ts>=tstar],n-1),0),
        col=rgb(0.5,0.5,1,0.5))
text(tmax-0.8,0.15,(1-gamma)/2,cex=2,col=rgb(0.2,0.2,0.5,1))
```
\

Thus, we determine a left area.
$$P(T<t^\star)=\gamma + \frac{1-\gamma}{2}= `r gamma`+`r (1-gamma)/2`=`r gamma+(1-gamma)/2`$$
It should be mentioned the expression can be simplified.
$$P(T<t^\star)= \frac{1}{2}+\frac{\gamma}{2}= 0.5+`r gamma/2`=`r gamma+(1-gamma)/2`$$


```{r left,fig.width=5,fig.height=3}
par(mar=c(2,0,0,0))
plot(ts,dt(ts,n-1),type="l",lwd=2,axes=F,ann=F)
abline(h=0)
polygon(c(-tmax,ts[ts<=tstar],tstar),
        c(0,dt(ts[ts<=tstar],n-1),0),
        col=rgb(1,1,0.5,0.5))
axis(1,at=c(-tmax,0,tstar,tmax),labels=c(paste(-tmax),"0",TeX("$$t^{*}$$"),paste(tmax)))
text(0,0.15,gamma+(1-gamma)/2,cex=2,col=rgb(0.5,0.5,0.2,1))
```
\

We rephrase the puzzle. We wish to determine $t^\star$ such that $P(T<t^\star)=`r gamma+(1-gamma)/2`$. This is easy to determine using a $t$-table, `qt` in R, `t.inv` in spreadsheets, or other methods for evaluating the [quantile function](https://en.wikipedia.org/wiki/Quantile_function){target="blank"}.

$$P(T<`r round(tstar,2)`)=`r gamma+(1-gamma)/2`$$
In the original phrasing,

$$P(|T|<`r round(tstar,2)`)=`r gamma`$$

So, $t^\star = `r round(tstar,2)`$. Now, use the given expressions.

$$\text{Lower bound} = `r xbar`-(`r round(tstar,2)`)(`r s`)/\sqrt{`r n`} ~=~ `r round(LB,1)`$$
$$\text{Upper bound} = `r xbar`+(`r round(tstar,2)`)(`r s`)/\sqrt{`r n`} ~=~ `r round(UB,1)`$$


### Spreadsheet

A spreadsheet can use the [`confidence.t`](https://support.google.com/docs/answer/9365672){target="blank"} function. It takes three arguments:

* Alpha ($\alpha$) is the two-tail area, $\alpha=1-\gamma=1-`r gamma`=`r 1-gamma`$
* sample standard deviation, $s=`r s`$
* Sample size, $n=`r n`$

The function returns the [margin of error](https://en.wikipedia.org/wiki/Margin_of_error){target="blank"}.
$$\text{margin of error} = \frac{t^\star s}{\sqrt{n}}$$

So, in this case, if in a spreadsheet you typed ```=confidence.t(1-`r gamma`,`r s`,`r n`)```, the result would be the margin of error, `r ME`. You then add/subtract the margin of error to/from the sample mean.
$$\bar{x}-\text{ME} = `r xbar`-`r round(ME,3)` = `r round(xbar-ME,3)`$$
$$\bar{x}+\text{ME} = `r xbar`+`r round(ME,3)` = `r round(xbar+ME,3)`$$

```{r echo=F}
labels = c("Conf_int when sigma is unknown",
          "",
          "xbar",
          "s",
          "n",
          "gamma",
          "",
          "ME",
          "LB",
          "UB")
formulas = c("","",paste0(xbar),
          paste0(s),
          paste0(n),
          paste0(gamma),"",
          "=CONFIDENCE.T(1-B6,B4,B5)",
          "=B3-B8",
          "=B3+B8")
sol = data.frame(labels,formulas)
write.table(sol,file="conf_int_t.csv", row.names=F, col.names=F)
```

You can do this with [a spreadsheet](conf_int_t.csv).

```{r fig.height=5,fig.width=9,echo=F}
draw_ss = function(df){
  nrow = length(df[,1])
  ncol = length(df[1,])
  par(mar=c(0,0,0,0))
  plot(0,0,type="n",xlim=c(0,1),ylim=c(0,1),axes=F,ann=F)
  hi = 1:((nrow+2)*2+1)
  vi = 1:((ncol+1)*2+1)
  hpos = seq(1,0,length.out=(nrow+1)*2+1)
  vpos = seq(0,1,length.out=(ncol+1)*2+1)
  hlin = hpos[hi%%2==1]
  ytex = hpos[hi%%2==0]
  vlin = vpos[vi%%2==1]
  xtex = vpos[vi%%2==0]
  abline(h=hlin)
  abline(h=hlin[2],lwd=3)
  abline(v=vlin)
  abline(v=vlin[2],lwd=3)
  for(i in 2:(ncol+1)){
    text(xtex[i],ytex[1],LETTERS[i-1],col=rgb(0.5,0.5,0.5))
  }    
  for(i in 2:(nrow+1)){
    text(xtex[1],ytex[i],i-1,col=rgb(0.5,0.5,0.5))
  }
  for(i in 1:nrow){
    for(j in 1:ncol){
      text(xtex[j+1],ytex[i+1],df[i,j])
    }
  }
}
draw_ss(sol)
```
\

### R

```{r echo=T}
xbar = xbar2
s = s2
n = n2
gamma = gamma2
df = n-1   # df is the degrees of freedom
tstar = qt(0.5+gamma/2, df)
ME = tstar*s/sqrt(n)
LB = xbar-ME
UB = xbar+ME
cat(sprintf("The lower bound: %.4f\nThe upper bound: %.4f",LB,UB))
```


Answerlist
--------
* Lower bound = $~`r xbar`-(`r round(tstar,2)`)(`r s`)/\sqrt{`r n`} ~=~ `r round(LB,1)`$
* Upper bound = $~`r xbar`+(`r round(tstar,2)`)(`r s`)/\sqrt{`r n`} ~=~ `r round(UB,1)`$


Meta-information
================
extype: cloze
exsolution: `r LB`|`r UB`
exclozetype: num|num
exname: confidence interval T
extol: `r tol`
exextra[numwidth,numeric]: 6

