```{r data generation, echo = FALSE, results = "hide"}
library(knitr)
opts_chunk$set(echo = FALSE)
options(scipen=100)

makebullseye = function(scale=1,axes=T){
  x = cos(2*seq(0,pi,length.out=1000)*pi)*scale
  y = sin(2*seq(0,pi,length.out=1000)*pi)*scale
  plot(x,y,type="l",lwd=3,xlim=c(-scale,scale)*1.1,ylim=c(-scale,scale)*1.1,ann=F,axes=axes)
  lines(x=c(-1,1,1,-1,-1)*scale*1.1,y=c(1,1,-1,-1,1)*scale*1.1,lwd=4)
  colors = c("yellow","yellow","red","red","blue","blue","grey30","grey30","white","white")
  colors2 = lapply(colors, col2rgb)
  linecolors = c(rep("black",6),"white",rep("black",3))
  textcolors = c(rep("black",6),rep("white",2),rep("black",2))
  for(i in 10:1){
    x2 = x*i/10
    y2 = y*i/10
    polygon(x2[1:(length(x2)-1)],y2[1:(length(x2)-1)],col=colors[i],border=F)
    lines(x2,y2,col=linecolors[i])
  }
  points(0,0,pch=20,cex=0.5)
}

# 600 mm diameter target (Indoor recurve)

n = sample(6*1:6,1)
sigx = round(runif(1,25,100))
sigy = round(sigx*runif(1,0.8,1.2))
gamma = sample(c(0.8,0.9,0.95,0.96,0.98,0.99),1)
zstar = qnorm(0.5+gamma/2)
mex = zstar*sigx/sqrt(n)
mey = zstar*sigy/sqrt(n)
mux = runif(1,-2*mex,2*mex)
muy = runif(1,-2*mey,2*mey)

X = round(rnorm(n,mux,sigx))
Y = round(rnorm(n,muy,sigy))

xbar = mean(X)
ybar = mean(Y)


# default_source_hook <- knit_hooks$get('source')
# knit_hooks$set(source = function(x, options) {
#   x <- gsub('gamma2',paste0(gamma),x)
#   x <- gsub('xlist',paste0(x2,collapse=", "),x)
#   default_source_hook(x, options)
# })
```

Question
========

When Archie shoots archery, she records the horizontal position ($x$) and vertical position ($y$) of every arrow (in millimeters), using the bullseye as the origin. 

From years of shooting, Archie has determined that $X$ and $Y$ are roughly bell-shaped with population standard deviations $\sigma_x = `r sigx`$ and $\sigma_y=`r sigy`$ mm. However, Archie has a new [sight](https://en.wikipedia.org/wiki/Sight_(device)){target="blank"}, so her current population means ($\mu_x$ and $\mu_y$) are unknown. (She hopes both are zero.)

Archie wants to get confidence intervals for $\mu_x$ and $\mu_y$, using a sample size $n=`r n`$ and a confidence level $\gamma=`r gamma`$. Whether or not a confidence interval straddles 0 will determine whether Archie adjusts that aspect of her sight.

### Critical value

 ##ANSWER1##

### Margins of error.
The margin of error represents how much variation we expect in sample means.
$$\text{ME}_x = \frac{z^\star \sigma_x}{\sqrt{n}}$$
$$\text{ME}_y = \frac{z^\star \sigma_y}{\sqrt{n}}$$
 ##ANSWER2##

 ##ANSWER3##

### Sample

Archie shoots $n=`r n`$ arrows.

```{r, fig.width=5,fig.height=5}
par(mar=c(3,3,0,0))
makebullseye(300,T)
mtext("horizontal position, x (mm)",1,line=2)
mtext("vertical position, y (mm)",2,line=2)
points(X,Y,pch=20)
abline(v=seq(-300,300,25),lwd=1,lty=2)
abline(h=seq(-300,300,25),lwd=1,lty=2)
```
\

```{r}
write.csv(data.frame(X,Y),"arrows.csv",row.names=F)
```

The exact positions can [be downloaded as a csv](arrows.csv){target="blank"}.

### Sample means

 ##ANSWER4##
 
 ##ANSWER5##

### Confidence intervals

Calculate the `r gamma*100`% confidence interval of $\mu_x$ by using $\bar{x}\pm\text{ME}_x$.

 ##ANSWER6##

 ##ANSWER7##

Calculate the `r gamma*100`% confidence interval of $\mu_y$ by using $\bar{y}\pm\text{ME}_y$.

 ##ANSWER8##

 ##ANSWER9##

### Inference

If a confidence interval straddles 0, Archie will leave that aspect alone. If a confidence interval does not straddle 0, Archie will adjust that aspect on the sight.

Does Archie adjust the horizontal aspect of her sight? ##ANSWER10##

Does Archie adjust the vertical aspect of her sight? ##ANSWER11##



Answerlist
--------
* Determine the boundary $z^\star$ such that $P(|Z|<z^\star)=`r gamma`$:
* Calculate $\text{ME}_x$, the margin of error when sampling $\bar{x}$:
* Calculate $\text{ME}_y$, the margin of error for sampling $\bar{y}$:
* Calculate $\bar{x}$, the horizontal sample mean:
* Calculate $\bar{y}$, the vertical sample mean:
* Calculate the lower boundary:
* Calculate the upper boundary:
* Calculate the lower boundary:
* Calculate the upper boundary:
* Yes
* No
* Yes
* No



Solution
========

[video](https://vimeo.com/531949449){target="_blank"}

```{r}
if(abs(xbar)>mex){
  ans10 = "10"
} else {
  ans10 = "01"
}
if(abs(ybar)>mey){
  ans11 = "10"
} else {
  ans11 = "01"
}
```



Answerlist
--------
* `r round(zstar,2)`
* `r round(mex,1)`
* `r round(mey,1)`
* `r round(xbar,1)`
* `r round(ybar,1)`
* `r round(xbar-mex,1)`
* `r round(xbar+mex,1)`
* `r round(ybar-mey,1)`
* `r round(ybar+mey,1)`
* The confidence interval of $\mu_x$ does `r if(mex<abs(xbar)){"not"}` contain 0. So Archie does `r if(mex>=abs(xbar)){"not"}` adjust the horizontal aspect of her sight.
* .
* The confidence interval of $\mu_y$ does `r if(mey<abs(ybar)){"not"}` contain 0. So Archie does `r if(mey>=abs(ybar)){"not"}` adjust the vertical aspect of her sight.
* .



Meta-information
================
extype: cloze
exclozetype: num|num|num|num|num|num|num|num|num|schoice|schoice
exsolution: `r zstar`|`r mex`|`r mey`|`r xbar`|`r ybar`|`r xbar-mex`|`r xbar+mex`|`r ybar-mey`|`r ybar+mey`|`r ans10`|`r ans11`
exname: confidence interval Z raw
extol: 0.01|0.1|0.1|1|1|1|1|1|1|0|0
exextra[numwidth,logical]: TRUE
exextra[stringwidth,logical]: TRUE
